---
title: "notebook20181212"
author: "Mark Hagemann"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Continuing with new A0-only likliehood. Current TODO:

- get MLE for A0 on Pepsi 1 cases
- calculate observed information matrix for new likelihood on Pepsi 1
- Show eigenvalues, first and last eigenvectors
- Connect to accuracy of MLE A0 estimate.
- Document math in reports/ rmarkdown document

### MLE for A0 on Pepsi 1 cases

Try a single case.

```{r}
case1 <- sscase
case1_ll <- case1 %>% 
  swot_bamdata(Qhat = 100) %>% 
  A0_llfun_vector(negative = TRUE, gradient = TRUE)

case1_A0hat <- case1 %>% 
  swot_bamdata(Qhat = 100) %>% 
  bamr::bam_priors() %>% 
  `[[`("logA0_hat") %>% 
  exp()


case1_inits <- c(case1_A0hat, 0.15)

case1_ll(case1_inits)

case1_mle <- nlm(f = case1_ll, p = c(case1_A0hat, 0.15), iterlim = 1000, hessian = TRUE)

case1_mle
```

Doesn't work so well! Inspect the Hessian.

```{r}
case1_hess <- hessian(case1_ll, case1_mle$estimate)

eigs1 <- eigen(case1_mle$hessian)
eigs1 <- eigen(case1_hess)
eigs1$values

ns <- length(eigs1$values) - 1

plot(eigs1$vectors[, 1])
plot(eigs1$vectors[, length(eigs1$values)])

# Make a unit vector
unitize <- function(x) {
  lenx <- as.vector(sqrt(x %*% x))
  x / lenx
}

unitize(c(rep(1, ns), 0)) %*% case1_hess %*% unitize(c(rep(1, ns), 0))
```

Can I now try for the whole pepsi 1 dataset?

```{r}
pepcases <- reachdata %>% 
  map(swot_bamdata)

pepnlls <- pepcases %>% 
  map(A0_llfun_vector, negative = TRUE, gradient = TRUE)

peppriors <- pepcases %>% 
  map(bamr::bam_priors)

pepinits <- peppriors %>% 
  map(~.$logA0_hat) %>% 
  map(~c(exp(.), 0.15))

pepmles <- map2(.x = pepnlls, .y = pepinits, 
                .f = possibly(~nlm(f = .x, p = .y, hessian = TRUE, iterlim = 1000),
                              otherwise = NA))

pepcodes <- map_int(pepmles, ~.$code)
pepcodes
pepests <- map(pepmles, ~.$estimate)
pephess <- map(pepmles, ~.$hessian)

pepeigs <- map(pephess, eigen)
```

Compare to truth

```{r}
getTruth <- function(swotlist) {
  trueA0 <- apply(swotlist$A, 1, median)
  truesigma <- mcman_sigma(swotlist)
  out <- c(trueA0, truesigma)
  out
}

getTruth(reachdata$Connecticut)  
peptruth <- map(reachdata, getTruth)
pepests$Connecticut
pepeigs$Connecticut$values

mineigs <- map_dbl(pepeigs, ~min(.$values))
mineigs %>% sort(decreasing = TRUE)

plot(pepeigs$Connecticut$vectors[, 4])

```

```{r}
compplot <- function(case, sigma = FALSE) {
  truth <- getTruth(reachdata[[case]])
  est <- pepests[[case]]
  if (!sigma) {
    truth <- truth[-length(truth)]
    est <- est[-length(est)]
  }
  plot(truth, est); abline(0, 1)
}
compplot("SacramentoDownstream")
```

How well is sigma estimated?

```{r}
peptruth_sig <- peptruth %>% 
  map_dbl(~.[length(.)])

pepests_sig <- pepests %>% 
  map_dbl(~.[length(.)])

plot(peptruth_sig, pepests_sig, log = "xy"); abline(0, 1)
sort(peptruth_sig, decreasing = TRUE)
```

