---
title: "notebook20180312"
author: "Mark Hagemann"
date: "March 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Today I am combing back through the A0 analysis from last week. Goal is to be methodical and document the steps. 

Recall: we have a nice linear model for estimating A0 using $\delta A$, S, and W. Only it doesn't always work, and it is not immediately clear why not. In my last notebook I showed: 

- A0 can be estimated using this method for most Pepsi cases (some failed)
- Relative error on A0 varied widely, with some in excess of 200%. 
- Linear model $R^2$ was not a reliable predictor of A0 inference accuracy. 
- Spot checks of linear models showed some collinearity of predictors, but nothing catastrophic. 
- Termplots of linear models showed good behavior in lots of cases, but not always great *fit* (again, not strongly related to A0 inference performance)
- Mass conservation and steady-state assumption do not appear to affect A0 inference performance. 
    - I should come up with a better way to quantify steady-state assumption validity.
- Manning $R^2$ appears to be very good even in cases where A0 inference was poor
    - ***Need to check this!*** Can Connecticut really have a minimum $R^2$ of 0.993?
- Empirical Manning-type regressions (of logQ on logA, logW, logS) differed widely from Manning exponents. 
    - Presumably collinearity was an issue, although I did not inventory this. 
    - Nearness of empirical log-scale regression coefficients to Manning exponents did not appear to indicate goodness of A0 inference. 
    - Substitution of emprical coefficients for Manning exponents in Connecticut A0 linear model did not improve A0 inference
        - TODO: Replicate this. 


Main action items are: 

1. Perform Manning regressions on every reach in every Pepsi case. 
    - Show termplots
    - Aggregate coefficient values, fit statistics into table
        - Verify results from before: Did Connecticut really have great fit?
    - PCA on predictors
2. Redo A0 lm for all cases
    - Diagnose, fix ones that failed before
    - Show termplots
    - PCA on predictors
3. Summarize results, diagnose inferrability in terms of e.g. data quality, river hydraulics 


## 0. Data prep

Copied from 0308.

```{r}
library(swotData)

pep1cases <- names(Pepsi_v2)
pep1ncs <- sprintf("../../swotData/data/NC_files/%s.nc", pep1cases)

nclists <- map(pep1ncs, nc_list) %>% 
  setNames(pep1cases)

nclists$Platte$River_Info.gdrch <- 1:12 # get rid of 13-14.

gdrchs <- map(nclists, ~.$River_Info.gdrch)

keepvars <- c("W", "S", "A", "Q")
reachparts <- paste0("Reach_Timeseries.", keepvars)

zeroToNA <- function(mat) {
  zeros <- mat <= 0
  mat[zeros] <- NA
  mat
}

reachdata0 <- nclists %>% 
  map(~.[reachparts]) %>% 
  map(setNames, keepvars) %>% 
  map2(gdrchs, function(x, y) map(x, function(z) z[y, ])) %>% 
  map(~map(., zeroToNA))

reachncols <- map(reachdata0, ~ncol(.$A))

reachA0 <- reachdata0 %>% 
  map(~.$A) %>% 
  map(apply, 1, median) %>% 
  map2(reachncols, function(x, y) matrix(rep(x, y), ncol = y, byrow = FALSE))

reachdA <- reachdata0 %>% 
  map(~.$A) %>% 
  map2(reachA0, function(x, y) x - y)

reachdata <- map2(reachdata0, reachdA, function(x, y) {x$dA <- y; x})


Qmats <- nclists %>% 
  map(~.$Reach_Timeseries.Q[.$River_Info.gdrch, ])
```


## 1. Manning regressions

```{r}
manlmfun <- function(datalist, qmat) {
  allist <- c(datalist, list(Q = qmat))
  reachlsts <- map(allist, log) %>% 
    map(split, f = 1:nrow(qmat))
  reachdfs <- reachdfs <- map(1:nrow(qmat), ~map(reachlsts, function(x) x[[.]])) %>% 
    map(as.data.frame)
  # browser()
  mandfs <- map(reachdfs, ~lm(Q ~ W + A + S, data = ., y = TRUE))
  mandfs
}

manlms <- map2(reachdata, Qmats, possibly(manlmfun, otherwise = NA))

which(is.na(manlms))

manlms$Tanana <- NA # This one doesn't have usable A data either. 
manlms <- manlms[!is.na(manlms)]
```

Try to fix failed cases. 

- Fixed several by adding zeroToNA step in data prep. 
- StLawrence are a lost cause, since no A data exists. 

```{r}
plot_DAWG(reachdata$Tanana$A)
plot_DAWG(reachdata$StLawrenceUpstream$A)

reachdata$StLawrenceDownstream <- NULL
reachdata$StLawrenceUpstream <- NULL
reachdata$Tanana <- NULL
```

Now make termplots and aggregate coefs, fit stats

```{r}
ggTermPlot(manlms$Connecticut[[1]], xvar = "A")
ggTermPlot(manlms$Connecticut[[2]], xvar = "A")


# coefs

getcoefs <- function(lmlist) {
  nreach <- length(lmlist)
  nms <- paste0(1:nreach)
  coeftbls <- map(setNames(lmlist, nms), function(x) summary(x)$coefficients) %>% 
    map(~as.data.frame(.)) %>% 
    map(setNames, c("est", "stderr", "t", "p")) %>% 
    map(~mutate(., var = rownames(.))) %>% 
    bind_rows(.id = "reach")
  coeftbls
}

mancoefs <- manlms[!is.na(manlms)] %>% 
  map(~getcoefs(.)) %>% 
  bind_rows(.id = "case")

ggcoefdf <- mancoefs %>% 
  mutate(ymax = est + 2 * stderr, ymin = est - 2 * stderr)

ggcoef_w <- ggcoefdf %>% 
  filter(var == "W") %>% 
  ggplot(aes(x = case, y = est)) +
  geom_pointrange(aes(color = case, ymin = ymin, ymax = ymax)) +
  geom_hline(aes(yintercept = -2/3)) +
  theme(axis.text.x = element_text(angle = 90))

ggcoef_s <- ggcoefdf %>% 
  filter(var == "S") %>% 
  ggplot(aes(x = case, y = est)) +
  geom_pointrange(aes(color = case, ymin = ymin, ymax = ymax)) +
  geom_hline(aes(yintercept = 1/2)) +
  theme(axis.text.x = element_text(angle = 90))

ggcoef_a <- ggcoefdf %>% 
  filter(var == "A") %>% 
  ggplot(aes(x = case, y = est)) +
  geom_pointrange(aes(color = case, ymin = ymin, ymax = ymax)) +
  geom_hline(aes(yintercept = 5/3)) +
  theme(axis.text.x = element_text(angle = 90))

# plotly::ggplotly(ggcoef_w)
# plotly::ggplotly(ggcoef_s)
# plotly::ggplotly(ggcoef_a)

# GoF stats
manR2 <- map(manlms, function(x) data.frame(R2 = map_dbl(x, markstats::Q2), 
                                            reach = as.character(1:length(x)))) %>% 
  bind_rows(.id = "case")

ggcoefdf %>% 
  left_join(manR2, by = c("case", "reach")) %>% 
  filter(var == "A") %>% 
  ggplot(aes(x = R2, y = est)) +
  geom_pointrange(aes(color = case, ymin = ymin, ymax = ymax)) +
  geom_hline(aes(yintercept = 5/3)) +
  theme(axis.text.x = element_text(angle = 90))

# termplots

allTermplots_mod <- function(model) {
  xvars <- names(model$model)[-1]
  out <- map(xvars, ~ggTermPlot(model, xvar = ., data = model$model))
  out
}

allTermplots_case <- function(modlist) {
  map(setNames(modlist, paste0("reach", 1:length(modlist))),
      allTermplots_mod)
}

gglist <- map(manlms, possibly(allTermplots_case, otherwise = NA))

gglist$Platte$reach6[[3]]
```

## 1.5: Canonical Manning exponents--how well does this explain data?

HOW TO QUANTIFY?!

Idea: $(QnA^{-5/3})^{3/5} = (W^{-2/3}S^{1/2})^{3/5}$. Look at $R^2$ of LHS on RHS for each case and reach. 

```{r}
makePartData <- function(lst, medianQ = FALSE) {
  
  Qmat <- lst$Q
  Wmat <- lst$W
  Amat <- lst$A
  Smat <- lst$S
  
  if (medianQ) {
    Q_median <- apply(Qmat, 2, median)
    Qmat_median <- matrix(rep(Q_median, nrow(Wmat)), nrow = nrow(Wmat), byrow = TRUE)
    Qmat <- Qmat_median
  }
  # browser()
  nreach <- nrow(Wmat)
  ntime <- ncol(Wmat)
  
  lhsmat <- t((Qmat * Amat^(-5/3))^(3/5))
  rhsmat <- t((Wmat^(-2/3) * Smat^(1/2))^(3/5))
  
  outdf <- data.frame(lhs = as.vector(lhsmat), rhs = as.vector(rhsmat), 
                      reach = rep(1:nreach, each = ntime))
  outdf
}

partR2df <- map(reachdata, makePartData, medianQ = TRUE) %>% 
  bind_rows(.id = "case")

partR2df %>% 
  # filter(case == "Connecticut") %>% 
  ggplot(aes(x = rhs, y = lhs)) +
  geom_point(aes(color = reach)) +
  facet_wrap(~case, scales = "free")


lm_nointercept <- function(x, y) {
  mod <- lm(y ~ 0 + x, data = data.frame(x = x, y = y), y = TRUE)
  out <- mod
  out
}
R2hack <- function(mod) {
  pred <- predict(mod)
  out <- markstats::R2(x = mod$y, pred)
  out
}

partR2res0 <- na.omit(partR2df)
partR2lm <- partR2res0 %>% 
  split(.$case) %>% 
  map(~lm_nointercept(x = .$rhs, y = .$lhs))

map_dbl(partR2lm, markstats::Q2) %>% sort()
map_dbl(partR2lm, R2hack) %>% sort()
```

I should really be doing this for individual reaches. And keep track of which ones fail the test. 

```{r}

partR2res0 <- na.omit(partR2df) %>% 
  mutate(casereach = paste0(case, reach))
partR2lm <- partR2res0 %>% 
  split(.$casereach) %>% 
  map(~lm_nointercept(x = .$rhs, y = .$lhs))

partR2res1 <- data.frame(casereach = names(partR2lm), 
                        R2 = map_dbl(partR2lm, markstats::Q2))

partR2res2 <- left_join(partR2res0, partR2res1, by = "casereach")


```

Now I can apply a sliding filter for R2 at reach scale to select reaches to use in A0est calculations. 


```{r}
ggTermPlot(partR2lm$SacramentoUpstream1, xvar = "x", 
           data = partR2lm$SacramentoUpstream$model)
foo <- lm(lhs ~ 0 + rhs, data = filter(partR2res0, case == "SacramentoUpstream"))
summary(foo)

partR2res0 %>% 
  filter(case == "SacramentoUpstream") %>% 
  ggplot(aes(x = rhs, y = lhs)) +
  geom_point(aes(color = reach)) +
  stat_smooth(method = "lm", formula = y ~ x + 0)



```


## 2. A0 lm for all cases

```{r}
estA0_list <- function(reachlist, keeprows = "all") {
  if (length(keeprows) == 1 && keeprows == "all")
    keeprows <- 1:nrow(reachlist$W)
  with(reachlist, estA0(wmat = W[keeprows, ], 
                        smat = S[keeprows, ], 
                        damat = dA[keeprows, ]))
}

# Control on which reaches to keep
minR2 <- -Inf
keeprowlist <- partR2res2 %>% 
  split(f = .$case) %>%
  map(~filter(., R2 > minR2)) %>% 
  map(~unique(.$reach))
  
A0lms <- map2(reachdata, keeprowlist, 
              possibly(function(x, y) estA0_list(reachlist = x, keeprows = y), 
                       otherwise = NA))
summary(A0lms$Connecticut)
which(is.na(A0lms))
A0lms <- A0lms[!is.na(A0lms)]
```

What's wrong with Platte?

- Missing A for a couple reaches. Deja vu, man. 
- FIXED. Just put these into gdrch

Now get errors.

```{r}
predA0 <- map(A0lms, coef) %>% 
  map2(keeprowlist[names(A0lms)], function(x, y) setNames(x, y))
trueA0 <- map2(reachdata[names(predA0)], keeprowlist[names(predA0)], 
               function(x, y) (x$A - x$dA)[y,1])


A0evaldf <- map2(trueA0[names(predA0)], predA0, 
                 function(x, y) data.frame(true = x, pred = y, reach = names(y))) %>% 
  bind_rows(.id = "case") %>% 
  group_by(case) %>% 
  mutate(relerr = (pred - true) / true)

glimpse(A0evaldf)
```

Plot it!

```{r}
A0evaldf %>% 
  ggplot(aes(x = case, y = relerr)) +
  geom_point(aes(color = case)) +
  theme(axis.text.x = element_text(angle = 90))

A0evaldf %>% 
  mutate(reach = as.numeric(reach)) %>% 
  left_join(partR2res2, by = c("case", "reach")) %>% 
  ggplot(aes(x = R2, y = relerr, color = case)) +
  geom_point(aes(group = case)) +
  xlim(-1, 1)
```

I now want to develop this further and make show relative error (across all cases) as a function of minimum R2 in partial Manning regression. This will take some data assembly. 

```{r}

calcRelErr_minR2 <- function(minR2) {
  
  keeprowlist <- partR2res2 %>% 
    split(f = .$case) %>%
    map(~filter(., R2 > minR2)) %>% 
    map(~unique(.$reach))
    
  A0lms <- map2(reachdata, keeprowlist, 
                possibly(function(x, y) estA0_list(reachlist = x, keeprows = y), 
                         otherwise = NA))
  A0lms <- A0lms[!is.na(A0lms)]
  
  predA0 <- map(A0lms, coef)
  trueA0 <- map2(reachdata[names(predA0)], keeprowlist[names(predA0)], 
                 function(x, y) (x$A - x$dA)[y,1])
  
  
  out <- map2(trueA0[names(predA0)], predA0, 
                   function(x, y) data.frame(true = x, pred = y)) %>% 
    bind_rows(.id = "case") %>% 
    group_by(case) %>% 
    mutate(relerr = (pred - true) / true)
  
  out
}

foo <- calcRelErr_minR2(0.9)

errByR2 <- map(c(-Inf, 0, 0.25, 0.5, 0.75, 0.9), calcRelErr_minR2) %>% 
  setNames(c("none", as.character(c(0, 0.25, 0.5, 0.75, 0.9)))) %>% 
  bind_rows(.id = "threshold")
errByR2$threshold <- as.factor(errByR2$threshold)


glimpse(errByR2)
```

Now I can make that plot!

```{r}
foo <- errByR2 %>% 
  mutate(absrelerr = abs(relerr),
         threshold = relevel(threshold, ref = "none")) %>% 
  group_by(case, threshold) %>% 
  mutate(medianerr = median(absrelerr)) %>% 
  ggplot(aes(x = threshold, y = absrelerr, color = case)) + 
  geom_point() +
  geom_line(aes(y = medianerr, group = case))
foo
```

That's...perplexing. Only thing I can think of to do next is to impose mass conservation and redo. 


## Old stuff below 

Maybe look at max p value as a predictor

```{r}
ggcoefdf %>% 
  left_join(manR2, by = c("case", "reach")) %>% 
  # filter(var == "A") %>% 
  ggplot(aes(x = R2, y = p)) +
  geom_point(aes(color = var)) +
  # geom_hline(aes(yintercept = 5/3)) +
  theme(axis.text.x = element_text(angle = 90))

maxpgg <- ggcoefdf %>% 
  filter(var != "(Intercept)") %>% 
  group_by(case) %>% 
  summarize(max_p = max(p)) %>% 
  left_join(A0evaldf, by = "case") %>% 
  ggplot(aes(x = max_p, y = relerr)) +
  geom_point(aes(color = case))

plotly::ggplotly(maxpgg)
```

This narrative not fit by:

- MississippiDownstream and SacramentoUpstream (low max P, poor accuracy)
- Ganges, GaronneUpstream (high max P, good accuracy)

```{r}
gglist$MississippiDownstream$reach1[[3]]
```

It may be time to start looking at Durand space. 

```{r}
A0lms$MississippiDownstream$model %>% 
  sample_n(1000) %>% 
  pairs()

plot(A0lms$MississippiDownstream)
plot(A0lms$SacramentoUpstream)

gglist$SacramentoUpstream$reach1[[3]]
```


```{r}
reachdata_df <- reachdata %>% 
  map(~tibble(W = as.vector(t(.$W)), S = as.vector(t(.$S)), 
             A = as.vector(t(.$A)), dA = as.vector(t(.$dA)),
             Q = as.vector(t(.$Q)),
             reach = rep(1:nrow(.$W), each = ncol(.$W)),
             A0 = A - dA)) %>% 
  bind_rows(.id = "case")


```



