---
title: "notebook20180611"
author: "Mark Hagemann"
date: "June 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Now that I've resolved the Jacobian issue and I have a clearer view of how to apply these models, do the following:

- Run Pepsi 2 using new manning_nolatent model
- Rerun with tighter A0 priors if no convergence.

```{r}
library(rstan)
testmod <- stan_model("../../bamr/exec/manning_nolatent.stan")
```

Prep data (copied from 0607 notebook)

```{r}
library(bamr)
swotlists <- reachdata[1:15] # Omit StLaw, Tanana, Wabash
qhats <- Qhats[names(swotlists)] %>% 
  map(~.[1])
bamdatas <- map2(swotlists, qhats, ~bam_data(.x$W, .x$S, .x$dA, .y))
otimes <- map(bamdatas, ~.$omitTimes)

smartomit <- function(qvec, omit) {
  if (length(omit) == 0) {
    return(qvec)
  } else {
    return(qvec[-omit])
  }
}

qobs <- swotlists %>% 
  map(~apply(.$Q, 2, mean)) %>% 
  map2(otimes, smartomit)

baminps <- map(bamdatas, bamr:::compose_bam_inputs)
```

Do those runs.

Start with one of just Manning, with and without meas error.

```{r}
samps1 <- sampling(testmod, data = baminps$Cumberland, 
                   cores = 3, chains = 3)
bam_hydrograph(samps1, qobs$Cumberland)
check_rhat(samps1)
cache("samps1")

samps2 <- sampling(testmod, data = baminps$Po, cores = 3, chains = 3)
cache("samps2")
check_rhat(samps2)
stan_trace(samps2, pars = paste0("A0[", 1:8, "]"))
bam_hydrograph(samps2, qobs$Po)
```

I wonder if logQ_sd prior is affecting this? I could inform this better using W stats. 

```{r}
sdQs <- reachdata %>% 
  map_dbl(~sd(apply(.$Q, 2, function(x) mean(log(x)))))
sdWs <- reachdata %>% 
  map_dbl(~sd(apply(.$W, 2, function(x) mean(log(x)))))

meanAs <- reachdata %>% 
  map(~apply(.$A, 1, function(x) median(log(x))))

logA0hats <- baminps %>% 
  map(~.$logA0_hat)

plot(unlist(meanAs[names(logA0hats)]), unlist(logA0hats)); abline(0, 1)

hist(sdQs)
plot(sdWs, sdQs)



```



```{r}
samplist <- list()
for (i in 1:length(baminps)) {
  sampsi <- sampling(object = testmod, data = baminps[[i]], 
                     cores = 3, chains = 3)
  samplist[[i]] <- sampsi
}
names(samplist) <- names(baminps)
cache("samplist")
```



Now trying a HydroSWOT-informed prior on sigma_logQ.

```{r}
estimate_sigmalogQ <- function(Wobs, Hobs) {
  lwsd <- apply(log(Wobs), 1, sd)
  hsd <- apply(Hobs, 1, sd)
  sigmahatlogQ <- 0.6165 + 0.64 * lwsd + 0.1899 * hsd
  sigmahatlogQ
}

# standard deviation should be sqrt(gcv(qsdlm)) = 0.334.
```

Try it for Po (even though it will take a while)

- Without hydroswot (blind prior) this is not much worse, with sd = 0.41


```{r}
baminps$Po$sigmahat_logQ = 0.943
baminps$Po$sigmasd_logQ = 0.334

samps3 <- sampling(testmod, data = baminps$Po, cores = 3, chains = 3, iter = 1000)

pairs(samps3, pars = c("logn", "sigma_logQ"))

check_rhat(samps3)

bam_hydrograph(samps3, qobs$Po)
get_posterior_mean(samps3, par = "sigma_logQ")
stan_trace(samps3, pars = "sigma_logQ")



stan_trace(samps2, pars = "sigma_logQ")
```

```{r}
testmod <- stan_model("../../bamr/exec/manning_nolatent.stan")

samps4 <- sampling(testmod, data = baminps$Po,
                   cores = 3, chains = 3, iter = 2000)

bam_hydrograph(samps4, qobs$Po) + scale_y_log10()
check_rhat(samps4)
stan_trace(samps4, pars = "logn")
stan_trace(samps4, pars = c("truesigma_man", "sigma_logQ", "logQbar"), inc_warmup = FALSE)

```

What if I tighten the A0 prior?

```{r}
samps5 <- baminps$Po %>% 
  # within(., {sigma_man = sigma_man / 2.5}) %>% 
  sampling(testmod, data = ., cores = 3, chains = 3, iter = 2000)
check_rhat(samps5)
stan_trace(samps5, "logn")
check_hmc_diagnostics(samps5)
get_posterior_mean(samps5, pars = c("logn", "sigma_logQ", "logQbar"))
stan_trace(samps5, pars = c("logn", "sigma_logQ", "logQbar"), inc_warmup = FALSE)
stan_trace(samps5, "A0")
bam_hydrograph(samps5, qobs$Po)

opts5 <- baminps$Po %>% 
  within(., {sigma_man = sigma_man / 2.5}) %>%
  optimizing(testmod, data = ., as_vector = FALSE, tol_grad = 1e-15,
             tol_rel_grad = 1e-13, tol_rel_obj = 1, iter = 10000)

opts5$par$logn
opts5$par$logQ %>% exp() %>% plot(type = "l")
points(qobs$Po)
```

No idea why this stuff isn't working anymore. Can I go back to my pared model? If that works then mine should too.

```{r}
pare_baminps <- function(baminplist) {
  out <- with(baminplist, list(
    ns = nx,
    nt = nt,
    x = 1/2 * log(Sobs) - 2/3 * log(Wobs),
    dA = rezero_dA(dAobs, "minimum"),
    dA_shift = dA_shift,
    sigma_err = median(sigma_man),
    mu_hat = logQ_hat + logn_hat,
    mu_sd = sqrt(logQ_sd^2 + logn_sd^2),
    logA0_hat = logA0_hat,
    logA0_sd = logA0_sd
  ))
  out
}

paremod <- stan_model("../src/manning_pared.stan")

pareopts1 <- baminps$Po %>% 
  pare_baminps() %>% 
  optimizing(paremod, ., as_vector = FALSE)

pareopts1$par$truesigma_err
pareopts1$par$sigma_y

paresamps1 <- baminps$Po %>% 
  pare_baminps() %>% 
  sampling(paremod, data = ., chains = 3, cores = 3, iter = 2000)


get_posterior_mean(paresamps1, pars = "A0")
stan_trace(paresamps1, pars = "A0")
get_posterior_mean(paresamps1, pars = c("truesigma_err", "sigma_y"))
pairs(paresamps1, pars = c("truesigma_err", "sigma_y", "mu", "A0[1]", "A0[2]"))
check_hmc_diagnostics(paresamps1)
check_rhat(paresamps1)

```


Is this just a function of the number of reaches? Do other cases work better?

```{r}
pareopts2 <- baminps$MississippiUpstream %>%
  pare_baminps() %>% 
  optimizing(paremod, ., as_vector = FALSE)

pareopts2$par$truesigma_err
pareopts2$par$sigma_y

paresamps2 <- baminps$MississippiUpstream %>% 
  # within(., {logA0_sd = 0.1}) %>% 
  pare_baminps() %>% 
  sampling(paremod, data = ., chains = 3, cores = 3, iter = 10000, thin = 10)


get_posterior_mean(paresamps2, pars = "A0")
stan_trace(paresamps2, pars = "A0", inc_warmup = TRUE) + xlim(0, 1000)
get_posterior_mean(paresamps2, pars = c("truesigma_err", "sigma_y"))
stan_trace(paresamps2, pars = c("truesigma_err", "sigma_y", "mu")) + xlim(0, 1000)
pairs(paresamps2, pars = c("truesigma_err", "sigma_y", "mu"))
check_hmc_diagnostics(paresamps2)
check_rhat(paresamps2)
```



Can I do any better at guessing sigma_logQ a priori?

```{r}
realsiglogQs <- map_dbl(reachdata, ~mean(apply(.$Q, 1, function(x) sd(log(x)))))

logWSsds <- map(reachdata, function(x) 1/2 * log(x$S) - 2/3 * log(x$W)) %>% 
  map_dbl(., function(x) mean(apply(x, 1, function(y) sd(y))))

logWSsds

plot(logWSsds, realsiglogQs)

sdlogQ_hat <- function(swotlist) {
  # logA0hat <- swot_vec2mat(bamr:::estimate_logA0(swotlist$W), swotlist$dA)
  # logAhat <- log(exp(logA0hat) + rezero_dA(swotlist$dA, "median"))
  A0hat <- swot_vec2mat(estA0(swotlist), swotlist$dA)
  logAhat <- log(A0hat + swotlist$dA)
  logQnhat <- 5 / 3 * logAhat + with(swotlist, 1/2 * log(S) - 2/3 * log(W))
  out <- apply(logQnhat, 1, sd, na.rm = TRUE)
  out
}

sdlogQ_hat(reachdata$Po)

sds1 <- reachdata[1:15] %>% 
  map(~apply(.$Q, 1, function(x) sd(log(x), na.rm = TRUE))) %>% 
  unlist()
sds2 <- unlist(map(reachdata[1:15], sdlogQ_hat))
plot(sds1, sds2); abline(0, 1)
```


I'm now reasonably well satisfied with the manning_nolatent model. Next I need to test its performance with observation error. 

```{r}
# pluck from bam_estimate:
baminputs <- baminps$Po %>% 
  within(., {sigma_man = sigma_man / 25})
logS_sigsq_obs <- bamr:::ln_sigsq(obs = baminputs$Sobs, err_sigma = baminputs$Serr_sd)
logW_sigsq_obs <- bamr:::ln_sigsq(obs = baminputs$Wobs, err_sigma = baminputs$Werr_sd)
baminputs$sigma_man <- sqrt(baminputs$sigma_man^2 + 
                            logS_sigsq_obs * (3/6)^2 + 
                            logW_sigsq_obs * (4/6)^2)
baminputs$sigma_amhg <- sqrt(baminputs$sigma_amhg^2 + 
                             logW_sigsq_obs)
plot_DAWG(baminputs$sigma_man)

```

```{r}
samps6 <- baminputs %>% 
  sampling(testmod, data = ., cores = 3, chains = 3, iter = 2000)
check_rhat(samps6)
stan_trace(samps6, "logn")
check_hmc_diagnostics(samps6)
get_posterior_mean(samps6, pars = c("logn", "sigma_logQ", "logQbar"))
stan_trace(samps6, pars = c("logn", "sigma_logQ", "logQbar"), inc_warmup = FALSE)
stan_trace(samps6, pars = c("sigma_logQ"), inc_warmup = TRUE) + ylim(0, 3)
stan_trace(samps6, "A0")
bam_hydrograph(samps6, qobs$Po)
```

Now try manning.stan.

```{r}
testmod2 <- stan_model("../../bamr/exec/manning.stan")

samps7 <- baminps$Kanawha %>% 
  within(., {sigma_man = sigma_man / 2.5}) %>% 
  sampling(object = testmod2, data = ., cores = 3, chains = 1, iter = 1000)

bam_hydrograph(samps7, qobs$Kanawha)
```

