---
title: "notebook20180423"
author: "Mark Hagemann"
date: "April 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, check out Steve's issue. 

```{r}
stevedata <- readRDS("../data/steve/BAMData.rds")
steveest <- readRDS("../data/steve/BAMestimate.rds")

library(rstan)
```

```{r}
lq6 <- extract(steveest, "logQ[6]")

str(lq6, 1)
plot(lq6[[1]] %>% exp())
```

Look at Q vs. manning params for this case. 

```{r}
plot_DAWG(stevedata$dAobs)

A0samps <- extract(steveest, "A0")
head(A0samps[[1]])

summary(steveest)

Asamps <- A0samps[[1]] %>% 
  split(1:nrow(.)) %>% 
  map(function(x) stevedata$dAobs + swot_vec2mat(x, stevedata$dAobs))

A6samps <- map(Asamps, function(x) x[, 6])

lognsamps <- extract(steveest, "logn")

plot(lognsamps[[1]])

lognsamps_list <- lognsamps[[1]] %>% 
  split(1:nrow(.))

awssamps <- Asamps %>% 
  map(function(A) stevedata$Wobs^(-2/3) * stevedata$Sobs^(1/2) * A^(5/3))
```

Now just look at time 6. 

```{r}
aws6samps <- awssamps %>% 
  map(function(x) x[, 6])

Q6estsamps <- map2(aws6samps, lognsamps_list, function(x, y) x / exp(y))

Q6estsamps[[1]]

hist(unlist(Q6estsamps))
hist(exp(lq6[[1]]))
```



```{r}

awssamps %>% 
  sample(1) %>% 
  `[[`(1) %>% 
  plot_DAWG()

plot_DAWG(awssamps[[1]])
plot_DAWG(awssamps[[2]])

awssamps %>% 
  sample(20) %>% 
  map(~apply(., 2, function(x) exp(mean(log(x))))) %>% 
  bind_rows() %>% 
  t() %>% 
  plot_DAWG
```

Now do the same for the whole thing.

```{r}
qhat_samps <- map2(awssamps, lognsamps_list, function(x, y) x / exp(y))

qhat_samps %>% 
  sample(1) %>% 
  `[[`(1) %>% 
  plot_DAWG()

qhat_samps %>% 
  sample(20) %>% 
  map(~apply(., 2, function(x) exp(mean(log(x))))) %>% 
  bind_rows() %>% 
  t() %>% 
  plot_DAWG

```

OK, responded to Steve. That's off my plate for now.

On to new stan models.

### Back to Steve's stuff (email and gchat correspondence)

```{r}
AirSWOT_Data <- stevedata
AirSWOT_priors <- structure(list(lowerbound_logQ = 3.11576832481911, upperbound_logQ = 9.6493176226168, 
    lowerbound_A0 = 30, upperbound_A0 = 1e+06, lowerbound_logn = -4.6, 
    upperbound_logn = -1.5, lowerbound_logQc = 0, upperbound_logQc = 10, 
    lowerbound_logWc = 1, upperbound_logWc = 8, lowerbound_b = 0.01, 
    upperbound_b = 0.8, sigma_man = structure(c(0.25, 0.25, 0.25, 
    0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 
    0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 0.25, 
    0.25), .Dim = c(4L, 6L)), sigma_amhg = structure(c(0.22, 
    0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 
    0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 0.22, 
    0.22, 0.22, 0.22), .Dim = c(4L, 6L)), logQc_hat = 5.34710753071747, 
    logWc_hat = 4.42537368001871, b_hat = c(0.02161, 0.02161, 
    0.02161, 0.02161), logA0_hat = c(4.53066986280961, 4.69196614238028, 
    4.62937503405083, 4.47473836822688), logn_hat = -3.5, logQ_sd = c(0.832554611157698, 
    0.832554611157698, 0.832554611157698, 0.832554611157698, 
    0.832554611157698, 0.832554611157698), logQc_sd = 0.832554611157698, 
    logWc_sd = 4.71249322990639, b_sd = 0.05, logA0_sd = 0.1, 
    logn_sd = 1, Werr_sd = 5L, Serr_sd = 1.6971e-05, dAerr_sd = 5), .Names = c("lowerbound_logQ", 
"upperbound_logQ", "lowerbound_A0", "upperbound_A0", "lowerbound_logn", 
"upperbound_logn", "lowerbound_logQc", "upperbound_logQc", "lowerbound_logWc", 
"upperbound_logWc", "lowerbound_b", "upperbound_b", "sigma_man", 
"sigma_amhg", "logQc_hat", "logWc_hat", "b_hat", "logA0_hat", 
"logn_hat", "logQ_sd", "logQc_sd", "logWc_sd", "b_sd", "logA0_sd", 
"logn_sd", "Werr_sd", "Serr_sd", "dAerr_sd"), class = "bampriors")


AirSWOT_man_amhg <- bamr:::bam_estimate(bamdata = AirSWOT_Data,variant = "manning",bampriors = AirSWOT_priors, reparam = FALSE, seed = 9124)
```


```{r}
Ssamps <- extract(AirSWOT_man_amhg, "Sact")[1]
Wsamps <- extract(AirSWOT_man_amhg, "Wact")[1]
dAsamps <- extract(AirSWOT_man_amhg, "dAact")[1]

library(reshape2)
logAsamps <- extract(AirSWOT_man_amhg, "logA_man")[[1]] %>% 
  melt()
logWsamps <- extract(AirSWOT_man_amhg, "logW")[[1]] %>% 
  melt()
logSsamps <- extract(AirSWOT_man_amhg, "logS")[[1]] %>% 
  melt()

logRHSsamps <- logAsamps %>% 
  mutate(value = 5/3 * logAsamps$value - 2/3 * logWsamps$value + 1/2 * logSsamps$value,
         value = exp(value)) %>% 
  split(.$iterations) %>% 
  map(~acast(., Var2 ~ Var3))

logRHSsamps[[1]] %>% 
  plot_DAWG()

logRHSsamps %>% 
  sample(20) %>% 
  map(~apply(., 2, function(x) exp(mean(log(x))))) %>% 
  bind_rows() %>% 
  t() %>% 
  plot_DAWG
```


What's that do once n is added in?

```{r}
lognsamps <- extract(AirSWOT_man_amhg, "logn")[[1]] %>% 
  melt()

head(lognsamps)


logmanqsamps <- logAsamps %>% 
  left_join(lognsamps, by = "iterations") %>% 
  mutate(value = 5/3 * logAsamps$value - 2/3 * logWsamps$value + 1/2 * logSsamps$value,
         value = exp(value) / exp(lognsamps$value)) %>% 
  # glimpse() %>% 
  select(-value.x, -value.y) %>% glimpse() %>% 
  split(.$iterations) %>% 
  map(~acast(., Var2 ~ Var3))

head(logmanqsamps)
  
logmanqsamps[[1]] %>% 
  plot_DAWG()

logmanqsamps %>% 
  sample(20) %>% 
  map(~apply(., 2, function(x) exp(mean(log(x))))) %>% 
  bind_rows() %>% 
  t() %>% 
  plot_DAWG

```

Still not agreeing with hydrograph!

```{r}
logQsamps <- extract(AirSWOT_man_amhg, "logQ")[[1]] %>% 
  melt()

head(logQsamps)

ggplot(logQsamps, aes(x = Var2, y = exp(value), group = Var2)) + 
  geom_boxplot()

bam_hydrograph(AirSWOT_man_amhg)

foo <- bam_estimate(bamdata = AirSWOT_Data,variant = "manning",
                    bampriors = AirSWOT_priors, reparam = FALSE, seed = 9124)
bam_hydrograph(foo)
```


## Stan models for A0

Do sampling for known sigma_dgdx and sigma_alpha. 

```{r}
stanmod1 <- stan_model("../stan/A0.stan")

data1 <- bat_data(swotlist = uscase, sigma_dgdx = 1.66e-5, sigma_alpha = 4.7e-3, sigma_err = 1.96e-2)
characterize_closure(uscase)
vb1 <- vb(stanmod1, data = data1, iter = 10000)
summary(vb1) # no good

samp1 <- sampling(stanmod1, data = data1, cores = 3, chains = 3, iter = 10000)

summary(samp1)$summary

stan_trace(samp1, "A0", inc_warmup = TRUE)

stan_trace(samp1, "logQn[1]", inc_warmup = TRUE)
stan_trace(samp1, "logQn[30]", inc_warmup = TRUE)
stan_trace(samp1, "alpha", inc_warmup = TRUE)
stan_trace(samp1, "dgdx[1]", inc_warmup = TRUE)
stan_trace(samp1, "dgdx[1]", inc_warmup = FALSE)
```



```{r}
uscase %>% 
  manning_gamma() %>% 
  decomp_gamma(xmat = uscase$x) %>% 
  `[[`("dgdx") %>% 
  plot()
```

```{r}
foo <- extract(samp1, "dgdx")[[1]] %>% 
  melt()

head(foo)

foo %>% 
  group_by(Var2) %>% 
  summarize(lwr = quantile(value, 0.1), upr = quantile(value, 0.9), mean = mean(value)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Var2, y = mean, ymin = lwr, ymax = upr)) +
  geom_pointrange()
```


```{r}
estA0_lm_ws35()

```



## Po posterior vs prior for MJ

```{r}
library(bamr)
source("../../airSWOT/lib/pripost.R")

load("../../airSWOT/cache/bd_po.RData")
load("../../airSWOT/cache/bp_po.RData")
load("../../airSWOT/cache/est_po.RData")

pripost_A0(bampriors = bp_po, stanfit = est_po)
pripost_alpha(bp_po, bd_po, est_po)
pripost_n(bampriors = bp_po, stanfit = est_po)
pripost_qbar(bampriors = bp_po, bd_po, est_po)

stan_trace(est_po, "logn", inc_warmup = TRUE)
```

What if I initialize logn to be really low?

```{r}
# foo <- bam_estimate(bd_po, "manning", bp_po, init_r = 10, iter = 2000)

stan_trace(foo, "logn", inc_warmup = TRUE)
pripost_n(bampriors = bp_po, stanfit = foo)
```

Doesn't change a thing. Weird. 

```{r}
glimpse(valdf)
pripost_qdot(bp_po, est_po)
```

