---
title: "notebook20180828"
author: "Mark Hagemann"
date: "August 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I finally have a workable gce + docker workflow for comparing results of different bamr branches. Now it's time to do those comparisons so I can push forward on improving bamr. Start by making a utility function for cluster planning.

```{r}
# Based on work in notebook20180823 and online tutorial linked therein.

library(future)
docker_clustfun <- function(ext_ip, docker_image, user = "markh", 
                            keyfile = "~/.ssh/google_compute_engine", ...) {
  cl <- makeClusterPSOCK(
  ext_ip,

  # User name; DigitalOcean droplets use root by default
  user = user,

  # Use private SSH key registered with DigitalOcean
  rshopts = c(
    "-o", "StrictHostKeyChecking=no",
    "-o", "IdentitiesOnly=yes",
    "-i", keyfile
  ),

  # Command to run on each remote machine
  # The script loads the tidyverse Docker image
  # --net=host allows it to communicate back to this computer
  rscript = c("sudo", "docker", "run", "--net=host", 
              docker_image, "Rscript"),

  # These are additional commands that are run on the remote machine. 
  # At minimum, the remote machine needs the future library to work—installing furrr also installs future.
  rscript_args = c(
    # Create directory for package installation
    "-e", shQuote("local({p <- Sys.getenv('R_LIBS_USER'); dir.create(p, recursive = TRUE, showWarnings = FALSE); .libPaths(p)})"),
    # Install furrr and future
    "-e", shQuote("if (!requireNamespace('furrr', quietly = TRUE)) install.packages('furrr')")
  ),

  # Actually run this stuff. Set to TRUE if you don't want it to run remotely.
  dryrun = FALSE,
  ...
)
}

```

Recall: I need to make a snapshot. Alternatively I can just clone the instances from the console. I think I'll do that. 

The workflow will then be:

```{r}
testcases <- reachdata[c("Cumberland", "Ganges", "SacramentoUpstream")]

toeval1 <- expression({
  test_bd <- purrr::map(testcases, ~swot_bamdata)
  test_be <- purrr::map(test_bd, ~bamr::bam_estimate(., variant = "manning", 
                                                     reparam = TRUE, meas_error = TRUE))
})

toeval2 <- expression({
  test_bd <- purrr::map(testcases, ~swot_bamdata)
  test_be <- purrr::map(test_be, ~bamr::bam_estimate(., variant = "manning", 
                                                     reparam = TRUE, meas_error = FALSE))
})

# Start the vm's (do this externally on command line)

clust1 <- docker_clustfun(ext_ip = "35.194.83.35", user = "markh",
                          docker_image = "markwh/bam-docker:master", 
                          # makeNode = myMakenode,
                          keyfile = "C:/Users/markh/.ssh/google_compute_engine", verbose = TRUE)
clust2 <- docker_clustfun(ext_ip = gce_get_external_ip("rocker3", zone = "us-east4-b"),
                          docker_image = "markwh/bam-docker:devel")
clust3 <- docker_clustfun(ext_ip = gce_get_external_ip("rocker4", zone = "us-east4-b"), 
                          docker_image = "markwh/bam-docker:consolidate")
```

