---
title: "notebook20180507"
author: "Mark Hagemann"
date: "May 7, 2018"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

I made a new stan model (A0_v3.stan) and a new math document (probabilisticModel.Rmd). Time to try it out!

```{r}
library(rstan)

# modi <- stan_model("../stan/A0_v3.stan")
# modi <- stan_model("../stan/A0_v3-1.stan")
modi <- stan_model("../stan/A0_v2-1.stan")


casei <- sscase

characterize_closure(casei)



stan_inpsi <- list(nt = ncol(casei$W),
                   nx = nrow(casei$W),
                   dA = t(casei$dA),
                   W = t(casei$W),
                   S = t(casei$S),
                   x = casei$x[, 1], 
                   
                   logA0_hat = mean(bamr::estimate_logA0(casei$W)),
                   # logA0_sd = 1,
                   sigma_logA0 = 1,
                   omega = make_omegas(n = nrow(casei$W)),
                   
                   sigma_dA = 1e-7,
                   sigma_dgdx = 4.5e-9,
                   sigma_nu = 1e-4,
                   sigma_err = 1.4e-3)


esti <- optimizing(modi, stan_inpsi, 
                   # init = list(A0 = rep(100, 3), 
                   #             dgdx = rep(0, 77),
                   #             nu = rep(0, 77), 
                   #             epsilon = rep(0, 77)),
                   sample_file = "optimlog.txt")
```

Finally something works! (v2-1) See what vb does.

```{r}
vbi <- vb(modi, stan_inpsi)
summary(vbi)
plot(vbi)

stan_inpsi$logA0_hat <- 4 # Something ridiculous. 

sampi <- sampling(modi, stan_inpsi, cores = 3, chains = 3)
summary(sampi)

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
```

Now put this into v2 and modify. 

v2-2 now includes closure noise as independent error. 

```{r}
modi_2 <- stan_model("../stan/A0_v2-2.stan")

# optimizing(modi_2, stan_inpsi)

# vbi <- vb(modi_2, stan_inpsi)
# summary(vbi)

sampi <- sampling(modi_2, stan_inpsi, pars = "A0", cores = 3, 
                  control = list(max_treedepth = 25),
                  chains = 2, iter = 10000)
summary(sampi)

pairs(sampi, pars = "A0")

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
```

This isn't working. I have a lot of redundant parameters. Need a new version with a refigured model block. Try jacking up number of iterations. Didn't work. Now a slightly modded version with uncertain sigma_epsilon and prior thereon

```{r}
modi_3 <- stan_model("../stan/A0_v2-3.stan")

# optimizing(modi_2, stan_inpsi)

# vbi <- vb(modi_2, stan_inpsi)
# summary(vbi)

sampi <- sampling(modi_3, stan_inpsi, pars = "A0", cores = 3, 
                  chains = 2, iter = 10000)
summary(sampi)

pairs(sampi, pars = "A0")

stan_trace(sampi, pars = c("A0"), inc_warmup = TRUE)
```

OK, that's not going anywhere. Moving on to v4, which gets rid of matrix-multiplication and instead estimates Qn. 

```{r}
modi_4 <- stan_model("../stan/A0_v4.stan")
newinps <- c(list(logQ_hat = 5, sigma_logQ = 1, logn_hat = -3.5, sigma_logn = 0.5),
             within(stan_inpsi, {
               dA = t(dA)
               W = t(W)
               S = t(S)
             }))

# optimizing(modi_4, newinps)

sampi <- sampling(modi_4, newinps, cores = 3, chains = 2)
summary(sampi, pars = "Qn")

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
stan_trace(sampi, pars = "Qn", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dA", inc_warmup = TRUE)
summary(sampi, pars = "sigma_dA")
```

Now this works! (It didn't before) I guess the difference must be in how I specify sigma_dA. Best not to do it as input data. 

Save this as A0_v4-1.stan. Then test it using a different dataset. 

```{r}
pocase <- reachdata$Po
poinps <- with(pocase, 
               list(
  nx = nrow(W),
  nt = ncol(W),
  W = W,
  S = S,
  dA = dA,
  logA0_hat = mean(bamr::estimate_logA0(W)),
  sigma_logA0 = 1,
  logQ_hat = 6,
  sigma_logQ = 2,
  logn_hat = -3.5,
  sigma_logn = 0.5
))

# optimizing(modi_4, newinps)

sampi <- sampling(modi_4, poinps, cores = 3, chains = 2)
summary(sampi, pars = "Qn")

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
stan_trace(sampi, pars = "Qn", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dA", inc_warmup = TRUE)
summary(sampi, pars = "sigma_dA")
```

That appears to be really good. Compare it against estA0 method. 

```{r}
lmests <- estA0(swotlist = within(pocase, {dA = rezero_dA(dA, zero = "min")}))
stanests <- get_posterior_mean(sampi, "A0")[, "mean-all chains"]

plot(lmests, stanests); abline(0, 1)

plot(lmests, apply(pocase$A, 1, min)); abline(0, 1)
plot(stanests, apply(pocase$A, 1, min)); abline(0, 1)

```

Cool! TODO: comparison of methods across all cases. 

First add in random closure error. 

```{r}
modi_4_2 <- stan_model("../stan/A0_v4-2.stan")
sampi <- sampling(modi_4_2, newinps, cores = 3, chains = 2)
summary(sampi, pars = "Qn")

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
# stan_trace(sampi, pars = "Qn", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dA", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_err", inc_warmup = TRUE)
summary(sampi, pars = "sigma_dA")
```

Nice! Now Po!

```{r}
sampi <- sampling(modi_4_2, poinps, cores = 3, chains = 2)
summary(sampi, pars = "Qn")

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
stan_trace(sampi, pars = "Qn", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dA", inc_warmup = TRUE)
summary(sampi, pars = "sigma_dA")
```

OK, that took 37 minutes. 

```{r}
stanests <- get_posterior_mean(sampi, "A0")[, "mean-all chains"]

plot(lmests, stanests); abline(0, 1)

plot(lmests, apply(pocase$A, 1, min)); abline(0, 1)
plot(stanests, apply(pocase$A, 1, min)); abline(0, 1)

summary(sampi, "A0")
```

That's not good. Also slow (38 minutes for Po). May be affected by improperly specified model (gotta put in that Jacobian). Also errors are not independent, so there's that. Make a model 4-3 with non-independent errors. 

Done. Try it!

```{r}
modi_4_3 <- stan_model("../stan/A0_v4-3.stan")
sampi <- sampling(modi_4_3, newinps, cores = 3, chains = 2)
summary(sampi, pars = "Qn")

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
# stan_trace(sampi, pars = "Qn", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dA", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dgdx", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_alpha", inc_warmup = TRUE)
summary(sampi, pars = "sigma_dA")
```

Try with uscase?

```{r}
usinps <- with(uscase, list(
  logQ_hat = 5,
  sigma_logQ = 1,
  logn_hat = -3.5,
  sigma_logn = 0.5,
  nx = nrow(W),
  nt = ncol(W),
  
  W = W,
  S = S, 
  dA = dA,
  x = x[, 1],
  logA0_hat = log(50),
  sigma_logA0 = 1
))

sampi <- sampling(modi_4_3, usinps, cores = 3, chains = 2)

stan_trace(sampi, pars = "A0", inc_warmup = TRUE)
# stan_trace(sampi, pars = "Qn", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dA", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_dgdx", inc_warmup = TRUE)
stan_trace(sampi, pars = "dgdx[10]", inc_warmup = TRUE)
stan_trace(sampi, pars = "alpha", inc_warmup = TRUE)
stan_trace(sampi, pars = "sigma_alpha", inc_warmup = TRUE)
```

