---
title: "notebook20180327"
author: "Mark Hagemann"
date: "March 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Already did some work at the end of yesterday's notebook. In short, my joint estimation of ndot and A0 is blowing up. Now I need to work on lisflood stuff in prep for meeting with Mike.

Last time I did lisflood was 3/23 (Friday).

- I now have a function to read in lisflood output as reaches, and convert to dawg (untidy) format.
- A0 not good on unadjusted data
- improves after adjusting via ccf.
- Sensitive to time and reach subsetting
- spatial variability also goes down after ccf lag
- R2 improves after ccf lag

I don't think I'll have time to rerun for longer steady-state conditions before the meeting, but:

- Already did some work on this in 0321 notebook--pick up analysis here
- In future, write function for generating lisflood boundary condition files


Main things:

- flow imbalance quantified
    - Still need to contextualize with spatial and temporal sampling frequency
- Inference pipeline outlined
- Diagnostics and validations enumerated, coded up
- Theoretically possible to jointly estimate ndot and A0 (iteratively), but it's blowing up--needs regularization

Other stuff:

- Deterministic omegas are the new thing
    - Fix inconsistencies depending on where dA is zeroed. 
- Restructuring all code, to be organized into a suite of R packages


```{r}
map_dbl(reachdata, possibly(swot_dqdx, otherwise = NA)) %>% 
  sort()

plot_DAWG(with(reachdata$Platte, manning_qdot(W, S, A)))
```


Now that the call is over, spend some time organizing funcitons. 


- Put Manning validation in its own script? Give its own prefix?
    - e.g. true_dqdx or val_dqdx vs. manning_dqdx
    - or mv_dqdx vs. mm_dqdx