---
title: "notebook20180605"
author: "Mark Hagemann"
date: "June 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Picking up where a very messy file left off after getting distracted from Pepsi 2 runs with bamr improvements. 

Remaining work to be done:

- troubleshoot divergent transitions, chain convergence DONE.
- Decide whether hierarchical error term is worth it to include
    - Not based on HMC diagnostics
- Put median A0 into other stan models, bam dataprep. DONE.
- Purge extraneous parameters from returned object. DONE.

THEN I can run Pepsi 2. 

I put a bunch of model runs into src/bam_experiments_pepsi1/

```{r}
load("../src/bam_experiments_pepsi1/bamruns4.RData")

pryr::object_size(bamruns4)
```

Since these are so big, best to just load 1 or 2 at a time.

```{r}
map_dbl(bamruns4$ests, max_rhat)
```

So it seems the divergence only comes in with hierarchical error. 

```{r}
load("../src/bam_experiments_pepsi1/bamruns5.RData")
map_dbl(bamruns5$ests, max_rhat)
```

```{r}
check1 <- check_rhat(bamruns5$ests$Connecticut)

ests5.2 <- bam_estimate(bamruns5$data$Connecticut, variant = "manning", 
                        bampriors = bamruns5$priors$Connecticut, 
                        cores = 3, chains = 3, iter = 3000, stanmodel = bamruns5$model)
check_rhat(ests5.2)

stan_trace(ests5.2, pars = "A0_min")
stan_trace(ests5.2, pars = "truesigma_man")

check_rhat(ests5.2) %>% 
  arrange(desc(stat)) %>% 
  head(20)
```

```{r}
check1 <- check_rhat(bamruns5$ests$GaronneUpstream)

ests5.2 <- bam_estimate(bamruns5$data$GaronneUpstream, variant = "manning", 
                        bampriors = bamruns5$priors$GaronneUpstream, 
                        cores = 3, chains = 3, iter = 3000, stanmodel = bamruns5$model)
check_rhat(ests5.2)

stan_trace(ests5.2, pars = "A0_min")
stan_trace(ests5.2, pars = "truesigma_man")
stan_trace(ests5.2, pars = "sigma_logQ")
stan_trace(bamruns5$ests$GaronneUpstream, pars = "sigma_logQ")
stan_trace(bamruns5$ests$GaronneUpstream, pars = "A0_min")


stan_trace(ests5.2, pars = "lp__")

check_rhat(ests5.2) %>% 
  arrange(desc(stat)) %>% 
  head(20)
```



Some further diagnostics.

```{r}
for (i in 1:length(bamruns4$ests)) {
  print(names(bamruns4$ests)[i])
  fiti <- bamruns4$ests[[i]]
  check_hmc_diagnostics(fiti)
  cat("\n")
}

for (i in 1:length(bamruns5$ests)) {
  print(names(bamruns5$ests)[i])
  fiti <- bamruns5$ests[[i]]
  check_hmc_diagnostics(fiti)
  cat("\n")
}
```

