---
title: "notebook20180514"
author: "Mark Hagemann"
date: "May 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today and this week: Clean up analysis and presentation of likelihood-based inference of A0. 

Make a document showing math and likelihood plots for simplified case

- Add on priors

Make swot cases for various contrived, simple cases. Show convergence of A0 parameters compared to more easilty inferrable parameters. 

Show how various error structure (gamma decomposition) affects likelihood plots. 


How to plot likelihood? 

- Contour funciton where other parameters are estimated via conditional maximum likelihood. 

Make a document showing different representations of mass-conserved Manning's equation. 

- BAM one
- simplified BAM
- linear model version
- pairwise equalities version

Also how these represent error structure


### Likelihood plotting

I need a function that will take a likelihood function of arbitrarily many parameters, create a partial function of fewer parameters (the independent variable(s) of my plot), and optimize the other parameters conditional on these independent variable parameters.

```{r}
nll1 <- reachdata$Severn %>% 
  swot_sset(keeplocs = 1:2) %>% 
  make_nllfun()

partmax_ll <- function(nllfun, ...) {
  partfun <- purrr::partial(nllfun, ...)
  partopt <- nlm(partfun, p = p)
  
}
```

I need to write up the different likelihood specifications. Then I can code each into a make_nll function in R.

OK, I have a few specifications. 

```{r}
# This one does SWOT likelihood
nllfun_v1 <-  function(swotlist) {
  W <- swotlist$W
  S <- swotlist$S
  X <- 4 * log(W) - 3 * log(S)
  dA <- swotlist$dA
  
  llfun <- function(A0vec, logQvec, logn, sigma = 0.25) {
    A0mat <- swot_vec2mat(rep(A0vec, length.out = nrow(dA)), dA)
    logQmat <- swot_vec2mat(logQvec, dA)
    A <- A0mat + dA
    
    mat1 <- (X - 10 * log(A) + 6 * logn + 6 * logQmat)^2 / (72 * sigma^2)
    mat2 <- 10 * log(A)
    
    out <- sum(mat1 + mat2)
    
    out
  }
  llfun
}


# Modified version that can (hopefully) be used with nlm
nllfun_v4 <-  function(swotlist) {
  W <- swotlist$W
  stopifnot(nrow(W) == 2)
  S <- swotlist$S
  X <- 2/3 * log(W) - 1/2 * log(S)
  dA <- swotlist$dA
  
  llfun <- function(pars) {
    A01 <- pars[1] 
    A02 <- pars[2] 
    logQn <- pars[3]
    sigma <- pars[4]
    A0mat <- swot_vec2mat(c(A01, A02), dA)
    A <- A0mat + dA
    
    mat1 <- (X - 5/3 * log(A) + logQn)^2 / (2 * sigma^2)
    mat2 <- 10 * log(A)
    
    out <- sum(mat1 + mat2)
    
    out
  }
  llfun
}


foo <- nllfun_v4(swot_sset(reachdata$Po, keeplocs = 1:2))

foo(c(A01 = 1000, A02 = 1000, logQn = 4, sigma = 1))
nlm(foo, p = c(A01 = 1000, A02 = 1000, logQn = 4, sigma = 1))

optim(par = c(A01 = 1000, A02 = 1000, logQn = 4, sigma = 1), foo)
optim(par = c(A01 = 1000, A02 = 1000, logQn = 4, sigma = 1))

opt2 <- sscase %>% 
  swot_sset(keeplocs = 1:2) %>% 
  nllfun_v4 %>% 
  optim(par = c(A01 = 10, A02 = 10, logQn = 2, sigma = 1))

opt2
```

Leave the rest of this as to-do for now. A more fruitful venture might be to contrive a simple case where all of the parameters can be visualized (or all but 1 or 2). 
