---
title: "notebook20180409"
author: "Mark Hagemann"
date: "April 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Recall work last week/weekend:

- Calculated real WS matrix
    - I did a few iterations of this, but settled on a version that matches the time-varying mean of a supplied estimate
    - Resulted in closure term
- Did for-real inversions of 2x2 cases. Still need to document this.
- Took some time coming up with additive error term in linA space. Struggled for a while, but ended up with some working functions. 
- Brushed up on error-in-variable model theory (VCU library)
    - McMan model is very complicated! Errors highly interdependent! 
    - TODO: tease out these dependencies
- Finally, did simulation of effects of error magnitude on A0 performance
    - Need to document the resulting functions and put them into lib/ DONE.
    
Picking up from weekend work. 

- Simulation of error, evaluation
- Repeat for unsteady case
- Repeat for Pepsi cases
- Test hypothesis that standard error of parameter estimates is a good predictor of performance. 
- Document this process of calculating closure, considering McMan as a error-in-variables model

## Error simulation for unsteady case

- Recreate steady-case simulation from weekend.
- See if I can represent in relative terms. 
    - Transform columns to have zero mean? Then I can represent error as ratio of variances. 


Read in case. 


```{r}
uscase <- lis_profiles("../lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20)
```


```{r}
# casei <- uscase
casei <- shiftcase
simsds <- 10^seq(-8, -2, length.out = 20) %>% 
  rep(each = 100)

nx <- nrow(casei$W)
nt <- ncol(casei$W)

simerrmats <- map(simsds, ~rnorm(nx * nt, 0, .)) %>% 
  map(~matrix(., nrow = nx, ncol = nt))
simwsmats <- map(simerrmats, ~(manning_linA_closed(casei) + .))

simA0ests <- map(simwsmats, ~estA0_lm_ws35(., dAmat = casei$dA)) %>% 
  map(~as.data.frame(as.list(coef(.)))) %>% 
  bind_rows(.id = "simno")

trueA0df <- data.frame(loc = c("x1", "x2", "x3"), A0true = c(101.95, 119.92, 99.97))

simoutdf <- simA0ests %>% 
  mutate(sd = simsds) %>% 
  select(-simno) %>% 
  gather(key = "loc", value = "A0est", -sd) %>% 
  left_join(trueA0df, by = c("loc")) %>% 
  mutate(relerr_pct = (A0true - A0est) / A0true * 100)
```

Now plug in observed errors and plot. 

```{r}
valdf <- data.frame(loc = c("x1", "x2", "x3"),
                    sd = apply(manning_linA_closure(casei), 1, sd),
                    A0est = estA0(casei)) %>% 
  left_join(trueA0df, by = "loc") %>% 
  mutate(relerr_pct = (A0true - A0est) / A0true * 100,
         # sd = sqrt(mean(sd^2)),  # Optional, to aggregate sd across locations.
         absrelerr = abs(relerr_pct))

hist(manning_linA_closure(swotlist = casei))
qqnorm(manning_linA_closure(swotlist = casei))

simoutdf %>% glimpse() %>% 
  mutate(absrelerr = abs(relerr_pct)) %>% 
  # ggplot(aes(x = sd, y = relerr_pct)) +
  ggplot(aes(x = sd, y = absrelerr)) +
  geom_point() +
  geom_point(data = valdf, color = "red", size = 5) +
  scale_y_log10() +
  scale_x_log10() +
  theme_bw()
```

So not a perfect simulation of error effects. Is it because of structure within errors?

```{r}
manning_linA_closure(casei) %>% 
  t() %>% 
  pairs()

manning_linA_closure(casei) %>% 
  # t() %>% summary()
  plot_DAWG()

(manning_linA_closure(casei) * casei$A) %>% 
  # t() %>% summary()
  plot_DAWG()

```

Likely so. I can repeat this for shifted case. 

```{r}
shiftcase <- swot_timelag(uscase, lags = c(0, -2, -4))
estA0(shiftcase)
casei <- shiftcase
```

Hmm. It seems I should be treating the sd differently. No, that didn't fix things either. 

So my simulation is not a great portrayal of impact of error magnitude on A0 inversion error. Table this for now. Work on:

- Pepsi cases estimate standard error vs A0 error plot. 
- Document closure, considering McMan as a error-in-variables model

## Pepsi cases coef standard error vs A0 error

```{r}

badinds <- which(names(reachdata) %in% 
                   c("StLawrenceDownstream", "StLawrenceUpstream", "Tanana"))

rd2 <- reachdata[-badinds] %>% 
  map(ccf_lag, Ahat = FALSE) %>% 
  map(function(x) {x$dA <- rezero_dA(x$dA, zero = "median"); x})

closures <- map(rd2, manning_closure)

peplms <- rd2 %>% 
  map(possibly(~estA0_lm(., ndot = 1), otherwise = NA))
  # map2(closures, possibly(~estA0_lm(.x, ndot = .y), otherwise = NA))

pepsmrys <- map(peplms[!is.na(peplms)], summary)
pepcoefdf <- map(pepsmrys, tidy) %>% 
  map(~mutate(as.data.frame(.), loc = paste0("x", 1:nrow(.)))) %>% 
  bind_rows(.id = "case")

trueA0s <- map(rd2, ~apply(.[["A"]], 1, median)) %>% 
  map(~data.frame(loc = paste0("x", 1:length(.)), A0 = .)) %>% 
  bind_rows(.id = "case")

pepcoefdf %>% 
  left_join(trueA0s, by = c("loc", "case")) %>% 
  mutate(relerr_pct = (estimate - A0) / A0 * 100) %>% 
  ggplot(aes(x = std.error, y = relerr_pct, color = case)) +
  geom_point()
  
```

nothing I'd hang my hat on. 


## Notes on CRSC - Measurement Error in Nonlinear Models

### Chapter 1 - Introduction

Classical measurement error model:

$$
W_{ij} = X_{ij} + U_{ij}
$$
- W is measured
- X is true
- U is error
    - $E[U_{ij} | X_{ij}] = 0$ 
    - No assumption on heteroscedasticity

Note: observed has larger variance than truth

- Contrast to **Berkson Measurement Error Model**: 
    - $X_{ij} = W_{ij} + U_{ij}$
    - $E[U_{ij} | W_{ij}] = 0$
    - Here truth is more variable than observed. 
    - Has more statistical power than classical. 

Extensions include adding bias, variance components:

$$
W_{ij} = \gamma_0 + \gamma_1X_{ij} + U_{ij} \\
U_{ij} = r_i + \epsilon_{ij} 
$$


### Chapter 2 - Important concepts

Functional vs structural models

- Functional: X_i's are fixed unknown parameters
    - Or random, but without a specified distribution or distributional parameters
- Structural: X_i's are random variables

Differential vs nondifferential errors

- Nondifferential: [Y|W, X] only depends on W through X. W is a *surrogate*. i.e. Y|X is indep. of W. 
- Differential if otherwise. e.g.:
    - If U is dependent on response errors, then it's differential. 

### Chapter 3 - Linear regression

Check out the following cited references:

- Hasenabeldy, Fuller, and Ware (1988): Indoor Air Pollution and pulmonary performance: investigating errors in exposure assessment. Statistics in Medicine, 8, 1109-1126.
- Hwang, J.T. (1986) Multiplicative errors in variables models with applications to the recent data released by the U.S. Department of Energy. Journal of the American Statistical Association, 81, 680-688.



## Relative errors using Pepsi

What do I want here? Something similar to the above, but x axis should be ratio of (variance of U) to (variance of X). Or I could simply use the magnitude of 1 / ndot. Call this alpha. 

```{r}

badinds <- which(names(reachdata) %in% 
                   c("StLawrenceDownstream", "StLawrenceUpstream", "Tanana"))

rd2 <- reachdata[-badinds] %>% 
  map(ccf_lag, Ahat = FALSE) %>% 
  map(function(x) {x$dA <- rezero_dA(x$dA, zero = "median"); x})

closures <- map(rd2, manning_closure)
alphas <- map(closures, ~(geomMean(.) / .))

peplogsds <- map_dbl(alphas, ~sd(log(.)))

pepsddf <- data.frame(case = names(peplogsds), sdlog = peplogsds)

peplms <- rd2 %>% 
  map(possibly(~estA0_lm(., ndot = 1), otherwise = NA))
  # map2(closures, possibly(~estA0_lm(.x, ndot = .y), otherwise = NA))

pepsmrys <- map(peplms[!is.na(peplms)], summary)
pepcoefdf <- map(pepsmrys, tidy) %>% 
  map(~mutate(as.data.frame(.), loc = paste0("x", 1:nrow(.)))) %>% 
  bind_rows(.id = "case")

trueA0s <- map(rd2, ~apply(.[["A"]], 1, median)) %>% 
  map(~data.frame(loc = paste0("x", 1:length(.)), A0 = .)) %>% 
  bind_rows(.id = "case")

pepcoefdf %>% 
  left_join(trueA0s, by = c("loc", "case")) %>% 
  mutate(relerr_pct = (estimate - A0) / A0 * 100) %>% 
  left_join(pepsddf, by = "case") %>% 
  ggplot(aes(x = sdlog, y = relerr_pct, color = case)) +
  geom_point()
  
```

