---
title: "notebook20180406"
author: "Mark Hagemann"
date: "April 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Reading at VCU library

Books that cover error in variable / measurement error models

- Andersen and Skovgaard: Regression with Linear Predictors 
    - Section in last chapter of book
- Young: Handbook of Regression Methods
    - Most of chapter 11
- Carroll, Ruppert, Strefanski, and Crainiceanu: Measurement Error in Nonlinear Models
    - Most of the book!
    

### Young, ch. 11

$$
\mathbf{W = X + U}
$$

- $\mathbf{W}$ is observed
- $\mathbf{X}$ is true predictors
- $\mathbf{U}$ is errors

Reliability ratio:

$$
\lambda_j = \frac{\sigma^2_{X_j}}{\sigma^2_{X_j} + \sigma^2_{U_j}}
$$

- Simple relationship exists between $\lambda_j$ and true vs estimated parameters (manifests as bias), but only for SLR. 
- In multiple regression this becomes "Attenuation bias" and is apparently more complex. More info available in Carroll et al.


### Carroll et al.

Chapters 1-3 (pp. 1 - 64) appear to be relevant. But this will take some energy and library hours to get through. 


## Simulation and empirical study. 

Meanwhile I'll just take the empirical approach, as I started to do yesterday.

In hindsight I should start with a dimensionally similar system to my sscase. But as I devised, with the following modifications:

- X is sampled from a normal(0, 1) distribution. 
- deltaA is sampled from a normal(0, sig_A) distribution, and adjusted to have median equal to zero
    - Thus A0 will all be zero
- Delta is sampled from a normal(0, sig_D) distribution

With the following experimental design

- sig_A and sig_D vary independently between 0.01 and 2
- 100 repeats for each

Plot and model relationships between A0 errors (bias, variance wrt zero). Also relate to lambda

Then validate for lisflood and Pepsi 1 cases. 

- Need a function to calculate Delta matrix
- Also function to calculate sig_A, sig_D, lambdas. 


Here we go!

### Simulation

Parameters of simulation

```{r}
nt <- 9
nx <- 3

sig_A <- 1.5
sig_D <- 0.2
```

Simulate X, A, Delta

```{r}
sim_x <- matrix(rnorm(nt * nx, 0, 1), ncol = nx)
sim_A <- matrix(rnorm(nt * nx, 0, sig_A), ncol = nx)
sim_dA <- sim_A - swot_vec2mat(apply(sim_A, 2, median), sim_A)

sim_D <- matrix(rnorm(nt * nx, 0, sig_D), ncol = nx)

sim_omegas <- make_omegas(nx)

sim_lhs <- map(sim_omegas, function(x) (sim_x + sim_D) %*% diag(x)) %>% 
  Reduce(rbind, .)

sim_rhs <- map(sim_omegas, function(x) ((sim_x + sim_D) * sim_dA) %*% -x) %>% 
  unlist()

```

Because of potential distributional issues with simulated data, I think it's better to start with modifying real data. 

Start with that closure funciton! To be clear: Xobs = Xreal + closure. 

```{r}
manning_linA_closure <- function(swotlist, ndot = 1) {
  
  if (length(ndot) > 1 && !is.matrix(ndot)) {
    ndot <- swot_vec2mat(ndot, swotlist$W)
  }
  
  geomMean <- function(x) {
    exp(mean(log(x)))
  }
  
  Qnvec <- 
  
  trueN <- manningN_list(swotlist)
  trueNdot <- trueN / geomMean(trueN)
  
  meas <- manning_ws35(swotlist = swotlist, ndot = ndot)
  # Qmat <- swot_vec2mat(apply(swotlist$Q, 2, mean), swotlist$Q)
  # Qmat <- swotlist$Q
  
  true1 <- manning_ws35(swotlist = swotlist, ndot = trueNdot)
  meanrat <- mean(meas) / mean(true1)
  
  # meanrat <- mean(meas / true1)
  
  true2 <- true1 * meanrat
  closr <- true2 - meas
  
  out <- list(obs = meas, real = true2, closure = closr)
  out
}

foo <- manning_linA_closure(uscase)

plot_DAWG(foo$real)

estA0_lm_ws35(foo$obs + foo$closure, dAmat = uscase$dA) %>% summary()

estA0_lm_ws35(foo$real, dAmat = uscase$dA) %>% summary()

```

Now I can do a simple simulation using sscase.

```{r}
casei <- sscase
closi <- manning_linA_closure(casei)
obswsmat <- manning_ws35(casei)
truewsmat <- obswsmat + closi

obsxmat <- estA0_moddf(obswsmat, dAmat = casei$dA)
truexmat <- estA0_moddf(truewsmat, dAmat = casei$dA)

mat1 <- as.matrix(select(obsxmat, -y))
mat2 <- as.matrix(select(truexmat, -y))
deltamat <- mat2 - mat1

plot(deltamat[, 1])
pairs(deltamat)
```

This is some wild stuff. 

```{r}
pairs(t(closi))
# pairs(closi)

eigen(cov(t(closi) * 182.18735))$values

```

OK, closure term is highly correlated across locations. That would indicate a time-varying component of n. 

```{r}
val_anova(casei)
```

Funny, it doesn't show up there. 


```{r}
plot_DAWG(closi)
plot_DAWG(obswsmat)
plot_DAWG(truewsmat)

summary(estA0_lm_ws35(truewsmat, casei$dA))

nmati <- with(casei, manningN(A, W, S, Q))

plot_DAWG(nmati ^ (-3/5) * obswsmat)
plot_DAWG(obswsmat / mean(nmati^(3/5)))

plot_DAWG(nmati ^ (-3/5))


obswsmat_meann <- obswsmat / mean(nmati^(3/5))
truewsmat_truen <- obswsmat * nmati ^ (-3/5)
lm2 <- estA0_lm_ws35(obswsmat_meann, casei$dA)
lm3 <- estA0_lm_ws35(truewsmat_truen, casei$dA)
summary(lm2)
summary(lm3)

diffmat2 <- truewsmat_truen - obswsmat_meann
plot_DAWG(diffmat2)
mean(diffmat2)

```

