---
title: "notebook20180321"
author: "Mark Hagemann"
date: "March 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Working more with Lisflood output--see if I agree with what Mike showed.

```{r}
pnums <- list.files("../../A0_identifiability/lisflood/toy_1/results_constant/", 
                       pattern = "\\.profile$", full.names = FALSE) %>% 
  stringr::str_extract("[0-9]{4}")

pnums

profiles <- list.files("../../A0_identifiability/lisflood/toy_1/results_constant/", 
                       pattern = "\\.profile$", full.names = TRUE) %>% 
  map(read_lisProfile) %>% 
  setNames(pnums) %>% 
  bind_rows(.id = "profile") %>% 
  mutate(time = as.numeric(profile)) %>% 
  arrange(time, ChanX) %>% 
  group_by(time) %>% 
  mutate(Slope = - c(NA, diff(WaterElev)) / c(NA, diff(ChanX))) %>% 
  ungroup()
```

That worked!

```{r}
profiles$time <- as.numeric(profiles$profile)

glimpse(profiles)
summary(profiles)

profiles %>% 
  # filter(ChanX == 100) %>% 
  filter(time > 20) %>% 
  ggplot(aes(x = time, y = Flow, color = ChanX, group = ChanX)) +
  geom_line()

profiles %>% 
  # filter(ChanX == 100) %>% 
  filter(time > 20) %>% 
  group_by(time) %>% 
  arrange(ChanX) %>%
  mutate(Slope  = c(NA, diff(WaterElev)) / c(NA, diff(ChanX))) %>% 
  ungroup() %>% 
  ggplot(aes(x = time, y = Slope, color = ChanX, group = ChanX)) +
  geom_line()

ggsave("slopePlot.png")


profiles %>% 
  # filter(ChanX == 100) %>% 
  filter(time > 20) %>% 
  ggplot(aes(x = time, y = WaterDepth, color = ChanX, group = ChanX)) +
  geom_line()
```

Now split into reaches. Discard final 10 km.

```{r}
reaches <- profiles %>% 
  mutate(reach = as.factor(floor(ChanX / 10000) + 1)) %>% 
  glimpse() %>% 
  group_by(reach, time) %>% 
  summarize(x_start = min(ChanX), x_length = max(ChanX) - min(ChanX),
            Width = median(Width), Slope = median(Slope, na.rm = TRUE), 
            Height = median(WaterElev), Depth = median(WaterDepth),
            Flow = median(Flow)) %>% 
  ungroup()

reaches %>% 
  # filter(reach == 1) %>% 
  # glimpse()
  ggplot(aes(x = time, y = Slope, color = reach, group = reach)) +
  geom_point()
  geom_line()
```


I should only use the times that are actually at steady-state. That should be just before the perturbations. 

See if I can check for steady-state via diff.

```{r}

mydiff <- function(x) c(NA, diff(x))

reaches_ss <- reaches %>% 
  group_by(reach) %>% 
  mutate(ss = mydiff(Width) + mydiff(Slope) + mydiff(Height) + 
           mydiff(Depth) + mydiff(Flow) == 0) %>% 
  # glimpse() %>% 
  ungroup()

reaches_ss %>% 
  filter(ss) %>% 
  ggplot(aes(x = time, y = Slope, color = reach, group = reach)) +
  geom_point()

```

I should rerun with longer intervals, but let's leave that as a todo for now. 

```{r}
ss_times <- reaches_ss %>% 
  select(time, reach, ss) %>% 
  group_by(time) %>% 
  summarize(n_ss = sum(ss, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(n_ss == 3) %>% 
  glimpse()
  
glimpse(ss_times)
```

### Working with Mike's unsteady case output.

```{r}
pnums <- list.files("../../A0_identifiability/lisflood/toy_1/results_simple/", 
                       pattern = "\\.profile$", full.names = FALSE) %>% 
  stringr::str_extract("[0-9]{4}")

pnums

profiles <- list.files("../../A0_identifiability/lisflood/toy_1/results_simple/", 
                       pattern = "\\.profile$", full.names = TRUE) %>% 
  map(read_lisProfile) %>% 
  setNames(pnums) %>% 
  bind_rows(.id = "profile") %>% 
  mutate(time = as.numeric(profile)) %>% 
  arrange(time, ChanX) %>% 
  group_by(time) %>% 
  mutate(Slope = - c(NA, diff(WaterElev)) / c(NA, diff(ChanX))) %>% 
  ungroup()
```

```{r}
glimpse(profiles)
```

```{r}
reaches <- profiles %>% 
  filter(ChanX < 30000) %>% 
  mutate(reach = as.factor(floor(ChanX / 10000) + 1)) %>% 
  glimpse() %>% 
  group_by(reach, time) %>% 
  summarize(x_start = min(ChanX), x_length = max(ChanX) - min(ChanX),
            Width = median(Width), Slope = median(Slope, na.rm = TRUE), 
            Height = median(WaterElev), Depth = median(WaterDepth),
            Flow = median(Flow)) %>% 
  ungroup()

reaches %>% 
  # filter(reach == 1) %>% 
  # glimpse()
  ggplot(aes(x = time, y = Slope, color = reach, group = reach)) +
  geom_point()
  geom_line()
```

I think I should start around time 20. 

```{r}
source("../../manningEval/lib/utils.R")
reaches2 <- reaches %>% 
  filter(time > 20) %>% 
  transmute(loc = as.numeric(reach), time, W = Width, 
            S = Slope, H = Height, Q = Flow) %>% 
  swot_untidy()
reaches2$dA <- calcdA_mat(reaches2$W, reaches2$H)

plot_DAWG(reaches2$dA)

A0lm1 <- with(reaches2, estA0(wmat = W, smat = S, damat = dA))

# A0lm1$model %>% 
#   sample_n(1000) %>% 
#   pairs()

summary(A0lm1)
```

That's puzzling. I should probably start thinking about a full-on package for exploring Manning behavior. Is it any better for smaller reaches? Only in that the estimates are less strongly negative. 

Go back to original 10-km reaches. 

### Continuing 3/23

Now I'm checking on flow imbalance effects using the same dataset.

Need to compute wave velocity. Easiest to just estimate from hydrograph, or I could do some relationship with flow velocity given area. But that would require digging into bathymetry.

```{r}
reaches %>% 
  ggplot(aes(x = time, y = Flow, color = reach, group = reach)) +
  geom_point()

reaches %>% 
  group_by(reach) %>% 
  mutate(tmax = which.max(Flow)) %>% 
  ungroup() %>% 
  mutate(time = time - (tmax - min(tmax))) %>% 
  ggplot(aes(x = time, y = Depth, color = reach, group = reach)) +
  geom_point()

```

Or could it be the fault of outliers?

```{r}
toestlist <- reaches %>% 
  group_by(reach) %>% 
  mutate(tmax = which.max(Flow)) %>% 
  ungroup() %>% 
  transmute(
    # time = time,
    time = time - (tmax - min(tmax)),
            loc = as.numeric(reach), W = Width, S = Slope, H = Height,
            Q = Flow) %>% glimpse() %>% 
  filter(
    !time %in% unique(time[Q < 40]),
    time > 4,
    time < 90) %>%
  swot_untidy()
toestlist$dA <- with(toestlist, calcdA_mat(w = W, h = H))

A0est2 <- estA0_list(toestlist)
summary(A0est2)
```

Doesn't seem to work so well. Give more thought to that R package. Quickly check Manning assumptions, and assumptions of A0 inference. 

- Manning partial residuals
- Mass conservation / steady-state
- A0 estimation
    - WS35 condition number
    - Backwards stepwise regression by condition number, R2
- Estimate deltaT for flow imbalance. 
    - Estimate wave celerity using Qhat, A0 estimate, slope, v bounds, n = 0.03
        - First pass: use average velocity 
        
```{r}


manningV <- function(Amat, Wmat, Smat, n = 0.03) {
  Amat^(2/3) * Wmat^(-2/3) * Smat^(1/2) / n
}

A0hat <- bamr:::estimate_logA0(toestlist$W) %>% exp()
A0hat_mat <- matrix(rep(A0hat, ncol(toestlist$W)), nrow = nrow(toestlist$W))

foo <- manningV(Amat = toestlist$dA + A0hat_mat, Wmat = toestlist$W, 
                Smat = toestlist$S)
plot_DAWG(foo)

# Another way, using Q / A
bar <- toestlist$Q / (toestlist$dA + A0hat_mat)
plot_DAWG(bar) 
```

The two methods are not terribly different--perhaps 60%. 

Now my question is whether I can use this to inform a search for best R2 by varying deltaT.

4 ways to do it:

- Single average wave celerity
- location-dependent average wave celerity
- time-dependent wave celerity
- time and space-dependent wave celerity

Start with the first one. 

Actually, first show how R2 changes as a function of time adjustment. 

- shift reaches 1, 2, 3 as:
- 0, 0, 1
- 0, 1, 1
- 0, 1, 2
- 0, 1, 3
- 0, 2, 3
- 0, 2, 4
- 0, 2, 5
- 0, 3, 5
- 0, 3, 6

```{r}

vsym <- seq(0, 3, by = 1/3) # time per reach!

te_lists <- list()
ae_lists <- list()

for (i in 1:length(vsym)) {
  vsymi <- vsym[i]
  
  tedf <- reaches %>% 
  transmute(
    # time = time,
    reach, 
    tadj = round((as.numeric(reach) - 1) * vsymi),
    time = time - tadj,
            loc = as.numeric(reach), W = Width, S = Slope, H = Height,
            Q = Flow) %>% glimpse() %>% 
  filter(
    !time %in% unique(time[Q < 40]),
    time > 10,
    time < 90)
  
  print(ggplot(tedf, aes(x = time, y = Q, color = reach, group = reach)) +
    geom_point() +
      ggtitle(paste0("v = ", vsymi)))
  
  te_lists[[i]] <- swot_untidy(tedf)
  
  te_lists[[i]]$dA <- with(te_lists[[i]], calcdA_mat(w = W, h = H))
  
  ae_lists[[i]] <- estA0_list(te_lists[[i]])
}
```

Check these out!

```{r}
R2df <- ae_lists %>%
  setNames(paste0(round(vsym, digits = 2))) %>%
  # map(~tidy(summary(.))) %>%
  map(glance) %>%
  bind_rows(.id = "v") %>% 
  transmute(v = as.numeric(v), R2 = adj.r.squared)
  

smrydf <- ae_lists %>%
  setNames(paste0(round(vsym, digits = 2))) %>%
  map(~tidy(summary(.))) %>%
  # map(glance) %>%
  bind_rows(.id = "v") %>% 
  mutate(v = as.numeric(v))

smrydf %>% 
  left_join(R2df, by = "v") %>% glimpse() %>% 
  ggplot(aes(x = R2, y = estimate)) +
  geom_point()

smrydf %>% 
  left_join(R2df, by = "v") %>% glimpse() %>% 
  ggplot(aes(x = v, y = R2)) +
  geom_point()

```

How about I plot them by v.

```{r}
te_lists %>% 
  map(swot_tidy) %>% 
  setNames(vsym) %>% 
  bind_rows(.id = "v") %>% 
  mutate(loc = as.factor(loc)) %>% 
  ggplot(aes(x = time, y = Q, color = loc, group = loc)) +
  geom_point() + 
  facet_wrap(~v, ncol = 2)
```

I'm noticing significantly different results for different runs of estA0. That means theat the random selections of Omega are affecting things. Time to switch to a deterministic generation of omega. 

```{r}
nrch <- 5

make_omega <- function(n, p1, p2) {
  stopifnot(p1 < p2 && p2 <= n && p1 >= 1)
  out <- rep(0L, n)
  out[p1] <- 1L
  out[p2] <- -1L
  out
}

make_omegas <- function(n) {
  p1s <- as.list(1:(n - 1))
  p2s <- map(p1s, function(x) (x + 1):n)
  out <- map2(p1s, p2s, function(x, y) map(y, ~make_omega(n, x, .)))
  unlist(out, recursive = FALSE)
}

make_omegas(6)
```

Cool, that's done. At some point, check whether this improves estimates on the Pepsi cases. Also, the design matrix will have (p^3 - p^2) / 2 rows. 

```{r}

x <- 1:20
y <- (x^2 * (x - 1)) / 2

plot(x, y)
```

That's actually not so bad. Nothing in the Pepsi case is that big. 

### CCF for reaches

There may also be a way to do the search for time lag using ccf.

- What space to do this in? log?
    - Only works for dA (and not A) in 3/5 space. 
    - No wait, it's not conserved in that space.
    - Do iterative a la EM?
    
    - Choose starting A0 estimates
    - find lags that maximize ccf
    - Re-estimate A0
    - Redo lag selection
    - Repeat until convergence
    
I mean, I could code that but no guarantee it would work. Fun exercise though!



```{r}

datalist <- reaches %>% 
  transmute(loc = as.numeric(reach), 
            time, W = Width, S = Slope, H = Height) %>% 
  swot_untidy()
datalist$dA <- with(datalist, calcdA_mat(w = W, h = H, zero = "minimum"))






# initialize dl, the datalist that will be modified
dl <- datalist

load("../../manningEval/cache/reachdata.RData")
dl <- reachdata$MississippiDownstream


for(i in 1:100) {
  # cat(i, "\n")
  ntimes <- ncol(dl$W)
  tomat <- function(x) {
    matrix(rep(x, ntimes), ncol = ntimes)
  }

  W <- dl$W
  S <- dl$S
  dA <- rezero_dA(dl$dA, "minimum")
  
  # initialize A0
  delta <- 10
  A0 <- coef(estA0(W, S, dA + delta))
  cat(A0, "\n")
  
  A0[A0 <= 0] <- min(W[W>0])
  A <- dA + tomat(A0)
  
  logmanlist <- (W^(-2/3) * S^(1/2) * A^(5/3)) %>% 
    t() %>% 
    log() %>% 
    as.data.frame()
   
  ccf(logmanlist$V1, logmanlist$V2)
  ccs <- map(2:length(logmanlist), 
             ~ccf(logmanlist$V1, logmanlist[[.]], plot = FALSE))
  bestlags <- c(0L, map_int(ccs, ~as.integer(.$lag)[which.max(.$acf)]))
  cat(bestlags, "\n")
  
  
  if (sum(!bestlags == 0) == 0)
    break
  
  dl <- swot_timelag(dl, bestlags)
}
```

Well the deterministic A0 estimation fixed the issue with different A0 reference points giving different A0 estimates (it wasn't ever actually an issue at all).

### Spitballing about package or at least functions

Data functions (many I already have in utils.R or in swotData package):

- deal with NAs TODO
- (un)tidy
- filter negatives / zeros



Primary actions:

- lag time to line up
- estimate A0
- Condition number of WS35 matrix
- Stepwise A0 regressions
    - R2
    - condition number
- lisflood conversion to swotdata
- characterize steady-state/ flow imbalance


If I have Q data

- Manning n ANOVA
- Flow imbalance (and how well it's approximated by SWOT data analysis)
- Reach-scale partial Manning regression behavior (i.e. R^2) -- from 0312 notebook




