---
title: "notebook20180823"
author: "Mark Hagemann"
date: "August 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I've made a lot of progress the last few days in getting docker images up, running, and pushed to dockerhub for the different branches of bamr. This will allow me to quickly verify results across branches without the bottleneck of my personal computer. In today's notebook I'll be documenting the workflow for using these in R. 

I still need to figure out how to run R in a remote vm from within R on local machine. Some options are:

- manually spin up images, get ip addresses, use these diretly in `future` call.
- Spin up using **googleComputeEngineR** and use similar to documented [here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html)

I currently have preference for the first, since I'm not terribly impressed with gceR syntax or documentation. However, this will require first:

- spinning up (or starting) a gce instance
- `docker run`ning on the instance with specified image
- exposing a port for ssh to use. 

Give this a shot. 



```{r}
library(future)
plan(remote(workers = "35.199.36.209"))
```

```{r}
library(googleComputeEngineR)
```

Inspect defaults. 

```{r}
projinfo <- gce_get_project()
projinfo$commonInstanceMetadata$items
```

See if I can reference a running instance

```{r}
rocker2 <- gce_get_instance("rocker2", zone = "us-east4-b")
is.gce_instance(rocker2)
gce_list_instances(zone = "us-east4-b")

rocker2 <- gce_vm(name = "rocker2", zone = "us-east4-b")
```

OK, I had to do a patch and pull request, but this works now. 

Now following instructions [here](https://cloudyr.github.io/googleComputeEngineR/articles/docker-ssh-futures.html#docker-commands)

Ugh, and I had to fix my HOME environment variable.

```{r}
gce_ssh_setup(rocker2)
foo <- gce_check_ssh(rocker2)
gce_ssh(rocker2, "echo foo")

docker_cmd(rocker2, "")
docker_cmd(rocker2, "instances", zone = "us-east4-b", capture_text = TRUE)
docker_cmd(rocker2, "ps", "-a", zone = "us-east4-b", capture_text = TRUE)
docker_inspect(rocker2, zone = "us-east4-b")
```


Hmm. Trying some more R-via-docker from the above-linked tutorial.

```{r}
vm <- gce_vm(template = "markwh/bam-docker", name = "rocker3")
vm <- gce_vm_container()
vm <- gce_vm(template = "r-base", name = "rbase")
vm <- gce_ssh_addkeys(username = "markh", instance = "rbase")
```


## Following a different tutorial

The tutorial [here](https://www.andrewheiss.com/blog/2018/07/30/disposable-supercomputer-future/#full-example) seems to be a good one. Follow that. 

The following is modified copy-paste from the above link.

```{r}
library(future)
# Public IP for droplet(s); this can also be a vector of IP addresses
ip <- "35.194.83.35"

# Path to private SSH key that matches key uploaded to DigitalOcean
ssh_private_key_file <- "~/.ssh/google_compute_engine"
docker_image <- "markwh/bam-docker:consolidate"


# Connect and create a cluster
cl <- makeClusterPSOCK(
  ip,

  # User name; DigitalOcean droplets use root by default
  user = "markwh",

  # Use private SSH key registered with DigitalOcean
  rshopts = c(
    "-o", "StrictHostKeyChecking=no",
    "-o", "IdentitiesOnly=yes",
    "-i", ssh_private_key_file
  ),

  # Command to run on each remote machine
  # The script loads the tidyverse Docker image
  # --net=host allows it to communicate back to this computer
  rscript = c("sudo", "docker", "run", "--net=host", 
              docker_image, "Rscript"),

  # These are additional commands that are run on the remote machine. 
  # At minimum, the remote machine needs the future library to work—installing furrr also installs future.
  rscript_args = c(
    # Create directory for package installation
    "-e", shQuote("local({p <- Sys.getenv('R_LIBS_USER'); dir.create(p, recursive = TRUE, showWarnings = FALSE); .libPaths(p)})"),
    # Install furrr and future
    "-e", shQuote("if (!requireNamespace('furrr', quietly = TRUE)) install.packages('furrr')")
  ),

  # Actually run this stuff. Set to TRUE if you don't want it to run remotely.
  dryrun = FALSE
)
```

That may have finally done the trick!

```{r}
plan(cluster, workers = cl)

n_cpus %<-% { parallel::detectCores()}
n_cpus

remotels %<-% {list.files()}
remotels
```

Try a bam estimate!

```{r}
bd1 <- swot_bamdata(reachdata$Ganges)

be_cons %<-% {bamr::bam_estimate(bd1, variant = "manning", meas_err = TRUE, reparam = TRUE)}

bam_hydrograph(be_cons)
```

So nice when something works! Next steps:

- create several vms, each with its own version of bamr
- run a few cases
- compare results (visually, probably). 