---
title: "notebook20171107"
output: html_document
---


Note: this is a post-hoc notebook made from an attempted clean script that got messy quick when I slowly realized I didn't know how to use the mle and mle2 functions. I figured it out, and decided I should preserve the mess. Now cleaning up the script in src/A0_likelihood.R.

# A0 likelihood

```{r}
ao_loglik <- function(dA_vec, A0, sigsq_logA) {
  logquan <- log(dA_vec + A0)
  loglik <- sum(-logquan - 1/2 * log(sigsq_logA) - 
                  1 / (2 * sigsq_logA) * (logquan - log(A0))^2)
  loglik
}
```

# try using mle function in stats4.

```{r}
makeNLL <- function(dA, trace = FALSE) {
  out <- function(A0, sigsq_logA) {
    
    if (A0 <= 0 | sigsq_logA <= 0 | any(dA < -A0)) {
      nll <- 1e23
    } else {
      logquan <- suppressWarnings(log(dA + A0))
      loglik <- sum(-logquan - 1/2 * log(sigsq_logA) - 
                      1 / (2 * sigsq_logA) * (logquan - log(A0))^2)
      nll <- -loglik
      
    }
    
    # if (is.na(nll))
    #   browser()
    
    if (trace) {
      cat(A0, " ", sigsq_logA, " ", nll, "\n")
    }
    
    nll
  }
}



dA_test <- test_full %>% 
  split(.$xs) %>% 
  map(~.[["dA"]])

dA_mle <- dA_test %>% 
  map(makeNLL) %>% 
  map(safely(mle), start = list(A0 = exp(9.1), sigsq_logA = 0.5))

testfun <- makeNLL(dA = dA_test[[1]], trace = TRUE)
test1 <- mle2(testfun, 
              start = list(A0 = -min(dA_test[[1]]) + 1,
                           sigsq_logA = 0.1))


fakeA <- rlnorm(1000, 1.5, 0.3)
fakedA <- fakeA - median(fakeA)

testfun2 <- makeNLL(dA = fakedA, trace = TRUE)
test2 <- mle2(testfun2, start = list(A0 = -min(fakedA) + 1,
                                     sigsq_logA = 0.01))
median(fakeA)

nll2 <- function(A0, sigsq_logA) {
  # if (A0 < 0 | sigsq_logA < 0)
  if (A0 <= 0 | sigsq_logA <= 0 | any(fakedA < -A0)) {
    nll <- 1e23
  } else {
    llik <- sum(dlnorm(fakedA + A0, meanlog = log(A0), sdlog = sqrt(sigsq_logA), 
                       log = TRUE))
    nll <- -llik
  }

  cat(A0, " ", sigsq_logA, " ", nll, "\n")
  
  nll
}

test3 <- mle2(nll2, start = list(A0 = -min(fakedA) + 1,
                                 sigsq_logA = 0.01))

```


Optimization using lognormal isn't working very well. Try using normal instead.

```{r}
nll4 <- function(A0, sigsq_logA) {
  if (A0 <= 0 | sigsq_logA <= 0 | any(fakedA < -A0)) {
    nll <- 1e23
  } else {
    llik <- sum(dnorm(log(fakedA + A0), mean = log(A0), sd = sqrt(sigsq_logA), 
                       log = TRUE))
    llik = llik - sum(log(fakedA + A0)) # jacobian adjustment
    nll <- -llik
  }

# cat(A0, " ", sigsq_logA, " ", nll, "\n")

nll
}

test4 <- mle2(nll4, start = list(A0 = -min(fakedA) + 1,
                                 sigsq_logA = 0.01))

```



So these are *exactly* the same. That's a good check, but not useful to the cause.
Time to make my own gradient function, I guess. 

```{r}
makeGrad <- function(dA) {
  gr <- function(A0, sigsq_logA) {
    aptl <- - sum(-1 / (dA + A0) - 1 / sigsq_logA * (log(dA + A0) - log(A0)) *
                    (1 / (dA + A0) - 1 / A0))
    sigsqptl <- - sum(-1 / (2 * sigsq_logA) + 1 / (2 * sigsq_logA^2) * 
                        (log(dA + A0) - log(A0))^2)
    
    cat("A0: ", A0, "; sigsq: ", sigsq_logA, "; gr: ", aptl, " ", sigsqptl, "\n")
    
    out <- c(aptl, sigsqptl)
    out
  }
}


mygr <- function(A0, sigsq_logA) {
  aptl <- - sum(-1 / (fakedA + A0) - 1 / sigsq_logA * (log(fakedA + A0) - log(A0)) *
                  (1 / (fakedA + A0) - 1 / A0))
  sigsqptl <- - sum(-1 / (2 * sigsq_logA) + 1 / (2 * sigsq_logA^2) * 
                      (log(fakedA + A0) - log(A0))^2)
  cat("A0: ", A0, "; sigsq: ", sigsq_logA, "; gr: ", aptl, " ", sigsqptl, "\n")

  out <- c(aptl, sigsqptl)
  out
}

test5 <- mle2(nll4, start = list(A0  = -min(fakedA) + 1,
                                 sigsq_logA = 0.01),
              gr = mygr)

```



Yay! That works! I'm puzzled by its scoping though.
Why can't it see dA without referencing global env?
Test again using my function.


```{r}
test6 <- mle2(makeNLL(fakedA, trace = TRUE), start = list(A0  = -min(fakedA) + 1,
                                 sigsq_logA = 0.01),
              gr = mygr)
```

That's the same!


## The next day ...

I'm still having trouble implementing this via map. Maybe I should resort to `nlm` with the nll and gradient. 

```{r}
mygr <- function(A0, sigsq_logA) {
  aptl <- - sum(-1 / (fakedA + A0) - 1 / sigsq_logA * (log(fakedA + A0) - log(A0)) *
                  (1 / (fakedA + A0) - 1 / A0))
  sigsqptl <- - sum(-1 / (2 * sigsq_logA) + 1 / (2 * sigsq_logA^2) * 
                      (log(fakedA + A0) - log(A0))^2)
  
  out <- c(aptl, sigsqptl)
  out
}


makeGrad <- function(dA) {
  gr <- function(pars) {
    A0 <- pars[1]; sigsq_logA <- pars[2]
    aptl <- - sum(-1 / (dA + A0) - 1 / sigsq_logA * (log(dA + A0) - log(A0)) *
                    (1 / (dA + A0) - 1 / A0))
    sigsqptl <- - sum(-1 / (2 * sigsq_logA) + 1 / (2 * sigsq_logA^2) * 
                        (log(dA + A0) - log(A0))^2)
    
    out <- c(aptl, sigsqptl)
    out
  }
  gr
}


makeNLL <- function(dA, trace = FALSE, gr_att = TRUE) {
  
  gradfun <- makeGrad(dA)
  
  out <- function(pars) {
    A0 <- pars[1]
    sigsq_logA <- pars[2]
    if (A0 <= 0 | sigsq_logA <= 0 | any(dA < -A0)) {
      nll <- Inf
    } else {
      logquan <- suppressWarnings(log(dA + A0))
      loglik <- sum(-logquan - 1/2 * log(sigsq_logA) - 
                      1 / (2 * sigsq_logA) * (logquan - log(A0))^2)
      nll <- -loglik
      
    }
    if (trace) {
      cat(A0, " ", sigsq_logA, " ", nll, "\n")
    }
    
    if (gr_att) {
      # browser()
      attr(nll, "gradient") <- gradfun(c(A0, sigsq_logA))
    }
    
    nll
  }
}
```

```{r}
dA_test <- test_full %>% 
  split(.$xs) %>% 
  map(~.[["dA"]])

foogr <- makeGrad(dA_test[[1]])
foo <- makeNLL(dA_test[[1]])

bar <- nlm(f = foo, p = c(1000, 0.1), print.level = 2)

plist <- dA_test[1:5] %>% 
  map(~c(-min(.) + 1, 0.1))

testmap <- dA_test[1:5] %>% 
  map(makeNLL) %>% 
  map2(plist, function(...) suppressWarnings(nlm(...))) %>% 
  map(~data.frame(A0 = .[["estimate"]][1], sigsq_logA = .[["estimate"]][2])) %>% 
  bind_rows(.id = "xs")
```

~~OK, something's wrong with the scoping, and it's really bothering me. I guess I'll try using a regular loop.~~

```{r}
testmap <- list()
for (i in 1:5) {
  nlli <- makeNLL(dA_test[[i]])
  testmap[[i]] <- nlm(nlli, p = c(1))
}

fakenll <- makeNLL(fakedA)
fakenll(c(3.811866, 0.01))
nlm(fakenll, p = c(4, 0.1)) # This is same answer as obtained via mle2. So that's good. 
```

Finally, I'm ready to put this in src/. 
