---
title: "notebook20180607"
author: "Mark Hagemann"
date: "June 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Trying to get these bam runs through. I finally managed to update the package, but I had to switch to my laptop. Will likely remain here until I finish the Pepsi 2 runs. First verify and troubleshoot performance using Pepsi 1. These are the runs I will do:

- Manning only
    - no meas error
        - unk_var = TRUE
        - unk_var = FALSE
    - reparam meas error
        - unk_var = TRUE
        - unk_var = FALSE
    - latent meas error
        - unk_var = TRUE
        - unk_var = FALSE
- Manning + AMHG
    - no meas error
        - unk_var = TRUE
        - unk_var = FALSE
    - reparam meas error
        - unk_var = TRUE
        - unk_var = FALSE
    - latent meas error
        - unk_var = TRUE
        - unk_var = FALSE
- AMHG only
    - no meas error
    - reparam meas error
    - latent meas error

That's 15 sets of runs total. That will take up not only my computing capacity, but also my human faculties. Cut this down to 6:

- Manning only
    - no meas error
        - unk_var = TRUE
        - unk_var = FALSE
    - latent meas error
        - unk_var = FALSE
- Manning + AMHG
    - no meas error
        - unk_var = FALSE
    - latent meas error
        - unk_var = FALSE
- AMHG only
    - no meas error
    - latent meas error

That's actually 7. First see if unk_var actually improves things on Manning only.


### Data and priors

I can now use default priors, so that's a plus.

```{r}
library(bamr)
library(rstan)
swotlists <- reachdata[1:15] # Omit StLaw, Tanana, Wabash
qhats <- Qhats[names(swotlists)] %>% 
  map(~.[1])
bamdatas <- map2(swotlists, qhats, ~bam_data(.x$W, .x$S, .x$dA, .y))
otimes <- map(bamdatas, ~.$omitTimes)

smartomit <- function(qvec, omit) {
  if (length(omit) == 0) {
    return(qvec)
  } else {
    return(qvec[-omit])
  }
}

qobs <- swotlists %>% 
  map(~apply(.$Q, 2, mean)) %>% 
  map2(otimes, smartomit)

map_int(qobs, length) == map_int(bamdatas, function(x) ncol(x$Wobs))
```

### Estimate

```{r}
ests1 <- map(bamdatas, ~bam_estimate(., variant = "manning", meas_error = FALSE, unk_var = FALSE))
cache("ests1")
rm(ests1)

```

```{r}
mypars <- c("logQtn", "A0_med", "man_lhs", "logA_man", 
              "logQnbar", "sigma_logQ", "Sact", "Wact", "dAact", 
              "logW", "logS") #hotfix so I don't have to reinstall after finding small bamr bug

ests2 <- map(bamdatas, ~bam_estimate(., variant = "manning", meas_error = FALSE, unk_var = TRUE,
                                     pars = mypars))
cache("ests2")
ests3 <- map(bamdatas, ~bam_estimate(., variant = "manning", meas_error = TRUE, reparam = FALSE,
                                     pars = mypars))
cache("ests3")

ests4 <- map(bamdatas, ~bam_estimate(., variant = "manning_amhg", meas_error = FALSE, 
                                     pars = mypars))
cache("ests4")

ests5 <- map(bamdatas, ~bam_estimate(., variant = "manning_amhg", meas_error = TRUE, reparam = FALSE,
                                     pars = mypars))
cache("ests5")
```

Save these priors and datas too.


Validating on desktop computer. Need to find out why areas are going so low. 

```{r}
testest <- bam_estimate(bamdatas$Kanawha, variant = "manning", meas_error = TRUE, 
                        pars = "abc", reparam = TRUE)

testlogAman <- get_posterior_mean(testest, "logA_man")
head(testlogAman)
dim(bamdatas$Kanawha$Wobs)
postlogA <- testlogAman[, 4] %>% 
  matrix(nrow = 4, byrow = TRUE)

plot_DAWG(postlogA)



```
```{r}
testlhs <- get_posterior_mean(testest, "man_lhs")
head(testlhs)
postlhs <- testlhs[, 4] %>% 
  matrix(nrow = 4, byrow = TRUE)

plot_DAWG(postlhs)
```



```{r}
postA0min <- get_posterior_mean(testest, "A0")
postA0med <- get_posterior_mean(testest, "A0_med")
```


What's the sscase do?

```{r}
ssbamdata <- with(sscase, bam_data(W, S, dA, 200))
plot(ssbamdata)              

ssest <- bam_estimate(ssbamdata, "manning", meas_error = FALSE)

get_posterior_mean(ssest, "A0")
get_posterior_mean(ssest, "logQ") %>% 
  `[`(, 4) %>% 
  exp()

```

```{r}
sscase$Q %>% apply(2, mean)
```

Try with old version of Manning model

```{r}
# oldmod <- stan_model("../src/manning_nolatent_old.stan")
oldmod <- stan_model("../../bamr/exec/manning_nolatent.stan")

oldpriors <- bam_priors(bamdatas$Kanawha, logA0_sd = 0.5)

oldest <- bam_estimate(bamdatas$Kanawha, bampriors = oldpriors, 
                       variant = "manning", meas_error = FALSE,
                       pars = "abc",
                        # pars = c("man_lhs", "man_rhs", "logA_man"),
                       stanmodel = oldmod, iter = 1000)

oldlhs <- get_posterior_mean(oldest, "man_lhs")
# oldrhs <- get_posterior_mean(oldest, "man_rhs")
oldlhsmat <- oldlhs[, 4] %>% 
  matrix(nrow = 4, byrow = TRUE)
# oldrhsmat <- oldrhs[, 4] %>% 
  # matrix(nrow = 4, byrow = TRUE)
# plot_DAWG(oldlhsmat - oldrhsmat)

oldlogA <- get_posterior_mean(oldest, "logA_man")[, 4] %>% 
  matrix(nrow = 4, byrow = TRUE)

plot_DAWG(oldlogA)

plot_DAWG(manning_qdot(reachdata$Kanawha, log = TRUE))
plot_DAWG(oldlhsmat)
plot_DAWG(manning_closure(reachdata$Kanawha, log = TRUE))

oldinps <- bamr:::compose_bam_inputs(bamdatas$Kanawha)
oldsigman <- oldinps$sigma_man
oldlhs_real <- with(reachdata$Kanawha %>% 
                      swot_sset(keeptimes = -bamdatas$Kanawha$omitTimes), 
                    (((5 / 3 * log(A)) - 
                      (2 / 3 * log(W)) + 
                      (1 / 2 * log(S))) / oldsigman))
olr2 <- oldlhs_real - swot_vec2mat(apply(oldlhs_real, 2, mean), oldlhs_real)

plot_DAWG(oldlhsmat)
plot_DAWG(olr2)

sum(oldlhsmat^2)
sum(olr2^2)

pripost_q(bampriors = bam_priors(bamdatas$Kanawha), stanfit = oldest, 
          true_Q = qobs$Kanawha)
```

Checking against batman_log. That gives really good estimates for A0, at least for Kanawha. Observe:

```{r}
foocase <- reachdata$Kanawha %>% 
  swot_purge_nas("times")
foocase$dA <- rezero_dA(foocase$dA, "median")
batman_A0 <- batman_log(foocase)$A0
batman_logQn <- batman_log(foocase)$logQn

foocase$Ahat <- foocase$dA + swot_vec2mat(batman_A0, foocase$dA)

batman_lhs <- manning_qdot(foocase, Avar = "Ahat", log = TRUE)

plot_DAWG(batman_lhs)
plot_DAWG(batman_lhs - swot_vec2mat(batman_logQn, foocase$dA))

```

```{r}
oldinps <- bamr:::compose_bam_inputs(bamdatas$Kanawha)
oldopts <- optimizing(oldmod, data = oldinps, iter = 10000,
                      tol_rel_grad = 1e-10,
                      tol_rel_obj = 1e-10)
oldopts$par[paste0("A0[", 1:4, "]")]

testopts <- optimizing(bamr:::stanmodels$manning_nolatent, data = oldinps)
testopts$par[paste0("A0[", 1:4, "]")]

get_posterior_mean(oldest, par = "A0")
stan_trace(oldest, "A0")
```

