---
title: "notebook20180403"
author: "Mark Hagemann"
date: "April 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Deciding to add some bamr functions into lib, for improving functionality

- now for A0 estimation. Make sure I can do this for width-invariant datasets.

```{r}
bamr::estimate_logA0(hcl2$W) %>% exp()
load("../../hydroSWOT/cache/alm.RData")
summary(alm)
markstats::ggTermPlot(alm, xvar = "lwbar")
markstats::ggTermPlot(alm, xvar = "lwsd")

```

Come to think of it, I should also fix ggtermplot. Ugh, that's been on my list for a while. 

OK, I now think the width-invarint case is low priority. sd term doesn't count for much. 

### A0-ndot algorithm. 

I have an idea to try out on the iterative algorithm. What if I adjust ndot in 3/5 space? That way I'm not messing with means in that space. How would that work? Maybe it wouldn't...Wait, yeah it should! Just let Qn be mean of rhs in 3/5 space. 

OK, I'm now doing all of the means in linear A space. Maybe that will help the blowup?

```{r}
foo <- estA0_ndot(hcl2 %>% swot_sset(keeplocs = 1:16), trace = TRUE, iters = 10)

```

Test it on a Pepsi case. Preferably one with substantial n variability. Like severn?

```{r}
testcase <- reachdata$Severn
# val_nd_plot(testcase)

foo <- estA0_ndot(testcase, count = TRUE, iters = 1000)

plot(unlist(foo$Qbarn))
fooA0 <- Reduce(rbind, foo$A0hats)
foondot <- Reduce(rbind, foo$ndots)

head(fooA0)
tail(fooA0)

head(foondot)
tail(foondot)

plot(fooA0[, 1])
realA0 <- apply(testcase$A, 1, min)

ind <- nrow(fooA0)
ind <- 1
plot(realA0, fooA0[ind, ], log = "xy"); abline(0, 1)
```

```{r}
val_nd_plot(testcase)
bar <- val_anova_lm(testcase)
summary(bar)

testcase$Ahat <- swot_A(A0vec = fooA0[nrow(fooA0), ], dAmat = testcase$dA)
manning_qdot(testcase, Avar = "Ahat") %>% plot_DAWG() + scale_y_log10()

testcase$Ahat <- swot_A(A0vec = fooA0[1, ], dAmat = testcase$dA)
manning_qdot(testcase, Avar = "Ahat") %>% plot_DAWG() + scale_y_log10()
plot_DAWG(testcase$Q) + scale_y_log10()
```

Well that's just not working. Think more about it later. Now it's time to focus on lisflood stuff. 

### Lisflood stuff

```{r}

is_ss <- function(...) {
  arglist <- list(...)
  ssvars <- map(arglist, function(x) c(1, diff(x))) %>% 
    map(function(x) x == 0) %>% 
    Reduce(`*`, x = .)
  out <- as.logical(ssvars)
  out
}

sscase <- lis_profiles("../lisflood/toy_1/results_const_v2/") %>% 
  lis_reaches(slope_method = "s_median", agfun = median) %>% 
  filter(loc != 4) %>%
  arrange(loc, time) %>% 
  group_by(loc) %>% 
  mutate(steady = is_ss(W, S, Q, H, D)) %>% 
  group_by(time) %>% 
  filter(sum(steady) == 3) %>% 
  group_by(Q) %>% 
  filter(time == time[1]) %>% 
  mutate(A = W * D) %>% 
  select(-steady) %>% 
  swot_untidy()
```


Can I test the hypothesis that reach 3 is useless because it does not have any ws variability?

```{r}
sscase %>% 
  swot_sset(keeplocs = c(1, 2)) %>% 
  estA0()

sscase %>% 
  swot_sset(keeplocs = c(1, 3)) %>% 
  estA0()

sscase %>% 
  swot_sset(keeplocs = c(2, 3)) %>% 
  estA0()

```

Huh, so it does work with reaches 2 and 3, but not reaches 1 and 3. I wonder why. 

- Check out condition numbers
- Bootstrap times

```{r}
sscase %>% 
  swot_sset(keeplocs = c(1, 2)) %>% 
  estA0_lm() %>% 
  A0_condno()

sscase %>% 
  swot_sset(keeplocs = c(1, 3)) %>% 
  estA0_lm() %>% 
  A0_condno()

sscase %>% 
  swot_sset(keeplocs = c(2, 3)) %>% 
  estA0_lm() %>% 
  A0_condno()

```

Correct. Now bootstrap. 

```{r}

casei <- uscase %>% 
  # swot_sset()
  swot_sset(keeplocs = c(1, 2)) %>% 
  swot_timelag(c(0, -2))

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE))
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = .))

boot_nunique <- map_int(sampcollist, ~length(unique(.)))
boot_lms <- map(boot_cases, estA0_lm)
boot_A0 <- map(boot_lms, coef)
boot_R2 <- map_dbl(boot_lms, A0_R2)
boot_condno <- map_dbl(boot_lms, A0_condno)

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  map2(boot_R2, ~mutate(.x, R2 = .y)) %>% 
  map2(boot_condno, ~mutate(.x, condno = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = R2)) +
  geom_point(position = position_jitterdodge()) +
  # geom_boxplot(fill = NA) +
  scale_colour_viridis_c() 
  # geom_violin(fill = NA)

boot_df %>% 
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>% 
  glimpse() %>% 
  ggplot(aes(x = R2, y = A0_estimate, color = reach)) +
  geom_point()

boot_df %>% 
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>% 
  glimpse() %>% 
  ggplot(aes(x = condno, y = A0_estimate, color = reach)) +
  geom_point()

boot_df %>% 
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>% 
  glimpse() %>% 
  ggplot(aes(x = condno, y = R2, color = reach)) +
  geom_point()
```

How can it be that using reaches 2 and 3 only results in models with infinite condition number, yet the estimate is still correct? Is it because of zero intercept? Must be!

```{r}
sscase %>% 
  swot_sset(keeplocs = c(2, 3)) %>% 
  swot_plot()
```

It's time I made that termplot function. 

```{r}
markstats:::ggTermPlot.lm

# Partial fit -- use broom::augment?

auglm <- augment(estA0_lm(sscase)) %>% 
  gather(key = variable, value = value, -y, -.fitted, -.se.fit)

head(auglm)

```


```{r}
ggTermPlot <- function (object, xvar = NULL, data = NULL, ..., facet_scales = "free") {
    if (!is.null(data)) 
        object$call$data <- as.name("data")
    predvars <- names(stats::model.frame(object))[-1]
    vis = purrr::map(predvars, ~visreg::visreg(object, xvar = ., plot = FALSE)) %>% 
      setNames(predvars)
    
    if (!is.null(xvar)) {
      vis <- vis[xvar]
    }
    
    visfits <- map(vis, ~.[["fit"]]) %>% 
      bind_rows(.id = "variable") %>% 
      gather(key = varkey, value = x, -variable, -y, 
             -visregFit, -visregLwr, -visregUpr) %>% 
      dplyr::filter(variable == varkey) %>% 
      dplyr::select(-varkey)
    visres <- map(vis, ~.[["res"]]) %>% 
      bind_rows(.id = "variable")  %>% 
      gather(key = varkey, value = x, -variable, -y, 
             -visregRes, -visregPos) %>% 
      dplyr::filter(variable == varkey) %>% 
      dplyr::select(-varkey)
    
    out <- visfits %>% 
      ggplot() + 
      geom_ribbon(aes(x = x, ymin = visregLwr, ymax = visregUpr), alpha = 0.5) + 
      geom_line(aes(x = x, y = visregFit)) +
      geom_point(aes(x = x, y = visregRes), data = visres) +
      facet_wrap(~variable, scales = facet_scales)
    out
}

object <- estA0_lm(sscase %>% swot_sset(keeplocs = c(1, 3)), intercept = FALSE)
ggTermPlot(object, xvar = "x1")
ggTermPlot(object, facet_scales = "free_x") + theme_bw()
```

Cool, I like that that's finally taken care of. Probably I should make a lib/markstats.R to put that in. 

```{r}
A0_condno(estA0_lm(sscase))
ggTermPlot(sscase %>% estA0_lm, facet_scales = "fixed")
```



### Unsteady case

```{r}
uscase <- lis_profiles("../lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20)
```


```{r}
uscase %>% 
  swot_sset(keeplocs = 1:2) %>% 
  estA0()
```


### After speaking with Mike

Try a for-real inversion on steady-state data.

```{r}
casei <- sscase
times <- sample(1:ncol(casei$W), 2)
locs <- sample(1:nrow(casei$W), 2)

xmati <- manning_ws35(casei)[locs, times]
toinvmat <- xmati %*% diag(c(1, -1))
invmat <- solve(toinvmat)

rhsi <- (xmati * casei$dA[locs, times]) %*% c(-1, 1)

invmat %*% rhsi
```

So that's that. 

```{r}
ggTermPlot(estA0_lm(sscase))

pairs(model.frame(estA0_lm(sscase)))

pairs(model.frame(estA0_lm(sscase, random_omega = TRUE)))
```


Now try correcting things by dividing out real n. 

```{r}
ni <- with(sscase, manningN(A = A, W = W, S = S, Q = Q))
ni

w23n <- sscase$W ^ (-2/3) / ni


case_closed <- sscase
case_closed$W <- w23n ^ (-3/2)

estA0(swotlist = case_closed)
```


```{r}

casei <- case_closed
# casei <- sscase

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE))
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = .))

boot_nunique <- map_int(sampcollist, ~length(unique(.)))
boot_lms <- map(boot_cases, estA0_lm, random_omega = FALSE)
boot_A0 <- map(boot_lms, coef)
boot_R2 <- map_dbl(boot_lms, A0_R2)
boot_condno <- map_dbl(boot_lms, A0_condno)

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  map2(boot_R2, ~mutate(.x, R2 = .y)) %>% 
  map2(boot_condno, ~mutate(.x, condno = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = R2)) +
  geom_point(position = position_jitterdodge()) +
  # geom_boxplot(fill = NA) +
  scale_colour_viridis_c() 

```

```{r}
casei <- sscase
times <- sample(1:ncol(casei$W), 2)
locs <- sample(1:nrow(casei$W), 2)
# locs <- c(1, 3)

aoi <- apply(casei$A, 1, min)
dai <- casei$A - swot_vec2mat(aoi, casei$A)
# dai <- casei$dA

xmati <- manning_ws35(casei)[locs, times] %>% t()
toinvmat <- xmati %*% diag(c(1, -1))
invmat <- solve(toinvmat)

rhsi <- (xmati * t(dai[locs, times])) %*% c(-1, 1)

invmat %*% rhsi

```


```{r}
xmeas <- manning_ws35(sscase)
xreal0 <- manning_ws35(case_closed)
xreal <- scale(manning_ws35(case_closed), scale = FALSE)

xdiff <- xreal - xmeas




```

