---
title: "notebook20180612"
author: "Mark Hagemann"
date: "June 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm ready to check the Pepsi 1 cases using the new bamr version. It seems reasonable to use the following for starters:

- sigma_man = 0.1
- logA0_sd = 0.5
- meas_error= TRUE
- reparam = TRUE
- iter = 2000
- adapt_delta = 0.95

Again, load the data

```{r}
library(rstan)
library(bamr)
swotlists <- reachdata[1:15] # Omit StLaw, Tanana, Wabash
qhats <- Qhats[names(swotlists)] %>% 
  map(~.[1])
bamdatas <- map2(swotlists, qhats, ~bam_data(.x$W, .x$S, .x$dA, .y))
otimes <- map(bamdatas, ~.$omitTimes)

smartomit <- function(qvec, omit) {
  if (length(omit) == 0) {
    return(qvec)
  } else {
    return(qvec[-omit])
  }
}

qobs <- swotlists %>% 
  map(~apply(.$Q, 2, mean)) %>% 
  map2(otimes, smartomit)

bampriors <- map(bamdatas, ~bam_priors(., sigma_man = 0.1))
```

Runs!

```{r}
runs1 <- list()

for (i in 1:length(bamdatas)) {
  runs1[[i]] <- bam_estimate(bamdata = bamdatas[[i]], 
                             bampriors = bampriors[[i]], variant = "manning", 
                             meas_error = TRUE, cores = 3, chains = 3, 
                             iter = 2000, reparam = TRUE)
}

rhats <- map(runs1, ~check_rhat(., plot = FALSE))

map(rhats, as.data.frame) %>% 
  bind_rows(.id = "case") %>% 
  group_by(case) %>% 
  summarize(maxrhat = max(stat))

```

Problematic cases are 1, 11, 13. Also check hmc diagnostics.

```{r}
i <- i + 1
print(i)
check_hmc_diagnostics(runs1[[i]])
```

Problematic cases: 1, 7, 8, 13

So 11 must be a marginal case. 

```{r}
stan_trace(runs1[[11]], "A0")
stan_trace(runs1[[11]], pars = c("logn", "logQbar", "sigma_logQ"))
```

11 just needs more samples. 

```{r}
i <- 11
runs1[[i]] <- bam_estimate(bamdata = bamdatas[[i]], 
                             bampriors = bampriors[[i]], variant = "manning", 
                             meas_error = TRUE, cores = 3, chains = 3, 
                             iter = 10000, warmup = 8000, reparam = TRUE)

check_rhat(runs1[[11]])

stan_trace(runs1[[11]], pars = "sigma_logQ")
```

Still need to work on 1, 7, 8, 13

```{r}
check_rhat(runs1[[1]])
stan_trace(runs1[[1]], "A0")
stan_trace(runs1[[1]], c("logn", "sigma_logQ", "logQbar"))
pairs(runs1[[1]], c("logn", "sigma_logQ", "logQbar"))
plot_DAWG(bamdatas[[1]]$Sobs)

runs1[[1]] <- bam_estimate(bamdata = bamdatas[[1]], variant = "manning", 
                           bampriors = bam_priors(bamdatas[[1]], sigma_man = 0.2,
                                                  logA0_sd = 0.1),
                           meas_error = TRUE,
                           reparam = TRUE, iter = 10000, warmup = 8000)

plot_DAWG(bamdatas[[1]]$dAobs %>% rezero_dA("minimum"))
```

Might just be a loss. Now work on 7, 8, 13.

```{r}
check_hmc_diagnostics(runs1[[7]])

check_rhat(runs1[[7]])
stan_trace(runs1[[7]], "A0")
stan_trace(runs1[[7]], c("logn", "sigma_logQ", "logQbar"))

runs1[[7]] <- bam_estimate(bamdata = bamdatas[[7]], variant = "manning", 
                           bampriors = bam_priors(bamdatas[[7]], sigma_man = 0.2,
                                                  logA0_sd = 0.1),
                           meas_error = TRUE,
                           reparam = TRUE, iter = 10000, warmup = 8000)

bam_hydrograph(runs1[[7]])
```

```{r}
check_hmc_diagnostics(runs1[[8]])

check_rhat(runs1.8)
stan_trace(runs1.8, "A0")
stan_trace(runs1.8, c("logn", "sigma_logQ", "logQbar"))

runs1.8 <- bam_estimate(bamdata = bamdatas[[8]], variant = "manning", 
                           bampriors = bam_priors(bamdatas[[8]], sigma_man = 0.2,
                                                  logA0_sd = 0.1),
                           meas_error = TRUE,
                           reparam = TRUE, iter = 10000, warmup = 8000)
bam_hydrograph(runs1.8)

runs1[[8]] <- runs1.8
```

Last one!

```{r}
check_hmc_diagnostics(runs1[[13]])

check_rhat(runs1[[13]])
stan_trace(runs1[[13]], "A0")
stan_trace(runs1[[13]], c("logn", "sigma_logQ", "logQbar"))

runs1[[13]] <- bam_estimate(bamdata = bamdatas[[13]], variant = "manning", 
                           bampriors = bam_priors(bamdatas[[13]], sigma_man = 0.1,
                                                  logA0_sd = 0.1),
                           meas_error = TRUE,
                           reparam = TRUE, iter = 10000, warmup = 8000)

bam_hydrograph(runs1[[13]])
bam_hydrograph(runs1[[10]])
stan_trace(runs1[[10]], c("logn", "sigma_logQ", "logQbar"))
```


That should do it!

Time to validate.

```{r}
vals1 <- list()

for (i in 1:length(runs1)) {
  vals1[[i]] <- bam_validate(runs1[[i]], qobs[[i]])
}

vals1 %>% 
  setNames(names(bamdatas)) %>% 
  map(~.[["stats"]]) %>%
  map(~as.data.frame(as.list(.))) %>% 
  bind_rows(.id = "case")

vals1 %>% 
  setNames(names(bamdatas)) %>% 
  map(~.[["valdata"]]) %>%
  bind_rows(.id = "case") %>% 
  glimpse() %>% 
  ggplot(aes(x = qobs, y = qpred)) +
  geom_point(aes(color = case)) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(intercept = 0, slope = 1)
```

That's no good. Retry with a fixed sigma_logQ.

```{r}

```

