---
title: "notebook20180323"
author: "Mark Hagemann"
date: "March 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm going to do more function/package devel later, but first quantifying flow imbalance

From notebook work, 2 error sources: time granularity, space granularity

- $\frac{\Delta t}{\bar{Q}} \frac{\partial Q}{\partial t}$
- $\frac{\Delta x}{\bar{Q}} \frac{\partial Q}{\partial x}$

Total granularity error is sum of these.


```{r}

mycase <- "Platte"

granerr <- function(qmat) {
  myq <- qmat
  dqdt <- apply(myq, 1, function(x) abs(diff(x)))
  # dterr <- mean(dqdt) / mean(as.vector(myq))
  dterr <- mean(dqdt / t(myq)[-1, ])
  
  dqdx <- apply(myq, 2, function(x) abs(diff(x)))
  # dxerr <- mean(dqdx) / mean(as.vector(myq))
  dxerr <- mean(dqdx / myq[-1, ])
  
  out <- data.frame(dx = dxerr, dt = dterr)
  out
}

granerrs <- reachdata %>% 
  map(~.$Q) %>% 
  # map(~.^(3/5)) %>% # This to put into A0 inference space
  map(granerr) %>% 
  bind_rows(.id = "case")

ge_gg <- granerrs %>% 
  arrange(dx) %>% 
  ggplot(aes(x = dx, y = dt)) +
  geom_point(aes(color = case))

plotly::ggplotly(ge_gg)
```

Strange that Severn doesn't have the same issues as Platte when it is almost as out of balance. This should become clearer when I build out the functions some more. 

```{r}
plot_DAWG(reachdata$Severn$Q)

plot_DAWG(reachdata$Severn$Q[, -18]) + scale_y_log10()
```

What? That doesn't look right. 

```{r}
apply(reachdata$Severn$Q[, -18], 2, diff) %>% 
  abs() %>% 
  mean()

mean(as.vector(reachdata$Severn$Q))
```

So the anser appears to be the order of division, average operations. Now fixed. 

```{r}
granerr(reachdata$Severn$Q[, -18])
```

I'm reasonably satisfied with this, although it may be difficult to assess without knowing Q. 

Looks like I can estimate flow imbalance using S? Math looks to be:

$$
\frac{\partial Q}{\partial x} \frac{\Delta x}{Q} \approx \frac{\Delta S}{2 S}
$$

Try for comparison. 

```{r}
Smat <- reachdata$Platte$S
granerr2 <- function(swotlist) {
  rhsmat <- with(swotlist, manning_qdot(W, S, A))
  diffs <- apply(rhsmat, 2, function(x) abs(diff(x)))
  dxerr <- mean(diffs / rhsmat[-1, ]) / 2
  dxerr
}

dxerrs_s <- reachdata %>% 
  # map(~.$S) %>% 
  map_dbl(granerr2)

granerrs$dx_s <- dxerrs_s

ge_gg2 <- granerrs %>% 
  na.omit() %>% 
  ggplot(aes(x = dx_s, y = dx, color = case)) +
  geom_point()

ge_gg2 +
  scale_x_log10() +
  scale_y_log10()

plotly::ggplotly(ge_gg2)
```

Well that doesn't work. Clearly. Something wrong with math? Yes. RHS needs to account for spatial differences in A, W. 


```{r}
qdot <- with(reachdata$Platte, manning_qdot(W, S, A))
```

Next I need to know how this affects A0 inference. But first develop the functions more and maybe draft a vignette?

- Yes, I like the idea of a vignette

Realizing that my projects have gotten tangled. Think of an ecosystem:

- swotData
    - just i/o, basic utility funcitons, no inference
- McMan
    - Mass-conserved Manning equation fucntions
- Lisfloodr
    - i/o, possibly running, plotting, conversion to swotData, etc. 
- Everything else just happens in a project

But at this stage I shouldn't worry about making packages. Just put everything into its own lib/ script and work with a ***single project***

- Transfer A0_identifiability to manningEval or vice versa?
    - Yes, I *think* transfer to manningEval. 
    
### Back after working on project migration and work in lib/, reports/mcfli-swotr.Rmd

Workflow is really coming together

Now I can really look at lisflood output. 

```{r}
toyrches <- lis_reaches("../lisflood/toy_1/results_simple/")

toydawg <- toyrches %>% 
  filter(time > 17) %>%  # Added after looking at plot_DAWG below
  filter(loc != 4) %>% 
  swot_untidy()

coef(estA0_list(toydawg))

toydawg %>% 
  ccf_lag() %>% 
  swot_dqdx()

toydawg$A <- toydawg$dA + apply(toydawg$W, 1, function(x) min(x[x>0]))

with(toydawg, plot_DAWG(manning_qdot(W, S, A))) + scale_y_log10()
with(ccf_lag(toydawg), plot_DAWG(manning_qdot(W, S, A))) + scale_y_log10()

swot_dqdx(toydawg)
```

~~Huh, things work a lot better when I don't omit the first 17 time points.~~ Nvermind! That was including reach 4, which is a bad reach. 


Next build out and apply Manning assumptions checks

```{r}
toyA0ests <- toydawg %>% 
  ccf_lag() %>% 
  estA0_list() 

A0_R2(toyA0ests)
A0_R2(estA0_list(toydawg))
```

```{r}
A0_condno(toydawg %>% ccf_lag)
```

Not sure yet how to interpret these, but there they are. 

Now on to Q-dependent measures

- Exact flow imbalance

```{r}
swot_dqdx(toydawg)
true_dqdx(toydawg$Q)

swot_dqdx(toydawg, method = "reldiff")
true_dqdx(toydawg$Q, method = "reldiff")
```

