---
title: "notebook20180328"
author: "Mark Hagemann"
date: "March 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Spent some time organizing and updating functions. It will be more work to get full consistency. Not to mention writing tests...

Major things to do now:

- Establish connection between performance diagnostics and actual validations.
- Run inference, diagnostics, and validation on Lisflood
    - steady-state
    - unsteady
- Write up soup-to-nuts on Pepsi 1 cases
- Diagnose and fix joint A0-ndot inference
- Figure out a better dqdx metric that accounts for temporal and spatial sampling. 


### Lisflood results

I built out the lisflood functions somewhat. 

```{r}
lisdata1 <- lis_profiles("../lisflood/toy_1/results_simple/")
lisreaches1 <- lis_reaches("../lisflood/toy_1/results_simple/")

lisswot1 <- lisreaches1 %>% 
  mutate(A = D * W) %>% 
  swot_untidy()
```

Can I chain together the inference pipeline?

```{r}
lisswot2 <- lisswot1 %>% 
  swot_sset(keeplocs = -4, keeptimes = -1:-15)
plot_DAWG(lisswot2$Q)


lisswot3 <- lisswot2 %>% 
  ccf_lag(verbose = TRUE) 
plot_DAWG(lisswot3$Q)

estA0(lisswot3)

lisswot3 %>% 
  estA0_lm() %>% 
  A0_R2

```

diagnostics now

```{r}
manning_dqdx(lisswot2)
manning_dqdx(lisswot3)

val_dqdx(lisswot2$Q)
val_dqdx(lisswot3$Q)
```

Further validation

```{r}
testlist <- lisswot3
# testlist <- reachdata$Po

val_linA_R2(testlist, mc = FALSE)
val_log_R2(swotlist = testlist)
val_plot(testlist, log = TRUE)
val_plot(testlist, log = FALSE)

val_linA_plot(testlist, mc = TRUE) 
val_linA_plot(testlist, mc = FALSE)

val_log_termplot(testlist, mc = TRUE, scales = "free_x")
```


Now move on to steady-state case. I need to amend the boundary condition file and rerun. 

```{r}
q_impose <- c(50, 75, 100, 125, 150, 175, 200, 250, 300)

hfun <- lis_happroxfun(qinfile = "../lisflood/toy_1/bdy_qin.tsv", 
                       houtfile = "../lisflood/toy_1/bdy_hout.tsv")

tsus <- 186400 # How long to stay at each steady-flow condition
tchg <- 10368 # How long to take to transition to next steady-state condition


newbdry <- lis_bdry(q_impose = q_impose, h_impose = hfun(q_impose), 
                    time_sustain = tsus, time_change = tchg)

lis_write_bdry(newbdry, "../lisflood/toy_1/const_v2.bdy", header = "See notebook20180328.Rmd")

newbdry$q
```

Churning through lisflood now. And done!

```{r}
lis_ss <- lis_profiles("../lisflood/toy_1/results_const_v2/")
rch_ss <- lis_reaches("../lisflood/toy_1/results_const_v2/")

glimpse(rch_ss)

ggplot(rch_ss, aes(x = time, y = Q, color = loc, group = loc)) + 
  geom_point()

is_ss <- function(...) {
  arglist <- list(...)
  ssvars <- map(arglist, function(x) c(1, diff(x))) %>% 
    map(function(x) x == 0) %>% 
    Reduce(`*`, x = .)
  out <- as.logical(ssvars)
  out
}

rch_ss2 <- rch_ss %>% 
  filter(loc != 4) %>% 
  arrange(loc, time) %>%
  group_by(loc) %>% 
  mutate(steady = is_ss(W, S, Q, H, D)) %>% 
  ungroup() %>% 
  glimpse()

summary(rch_ss2)
n_rch <- length(unique(rch_ss2$loc))
rch_ss3 <- rch_ss2 %>% 
  filter(steady) %>% 
  group_by(time) %>% 
  filter(sum(steady) == n_rch) %>% 
  ungroup() %>% 
  # glimpse() %>% 
  group_by(Q) %>% 
  filter(time == time[1]) %>% 
  glimpse()


rch_ss3 %>% 
  ggplot(aes(x = time, y = Q, color = loc, group = loc)) + 
  geom_point()

```

Finally ready to make it into swot-like untidy data

```{r}
ss3data <- rch_ss3 %>% 
  mutate(A = D * W) %>% 
  swot_untidy()

estA0(ss3data, zero = "minimum")

val_linA_plot(ss3data)
val_linA_plot(lisswot3)

val_log_termplot(ss3data, scales = "free_x")
val_plot(ss3data, log = FALSE)
val_log_termplot(lisswot3)

```


Do ANOVA

```{r}
anova1 <- val_anova(lisswot3)

val_nd_plot(lisswot3)
val_nd_plot(ss3data)
```

This reveals something not good in unsteady data. logn is > 2?!

```{r}
with(lisswot3)
```

