---
title: "notebook20180515"
author: "Mark Hagemann"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I have a list of tasks. First make a simple likelihood-based model for a single pair of reaches. I can modify yesterday's function. 

```{r}
nllfun_v4 <-  function(swotlist, maxval = Inf) {
  W <- swotlist$W
  stopifnot(nrow(W) == 2)
  S <- swotlist$S
  X <- 2/3 * log(W) - 1/2 * log(S)
  dA <- swotlist$dA
  
  llfun <- function(pars) {
    A01 <- pars[1] 
    A02 <- pars[2] 
    logQn <- pars[3]
    sigma <- pars[4]
    
    if (sigma <= 0) return(maxval)
    
    A0mat <- swot_vec2mat(c(A01, A02), dA)
    A <- A0mat + dA
    
    mat1 <- (X - 5/3 * log(A) + logQn)^2 / (2 * sigma^2)
    mat2 <- 5/3 * log(A)
    const <- log(sqrt(2 * pi) * sigma)
    
    out <- sum(mat1 + mat2 + const)  
    out
  }
  
  grfun <- function(pars) {
    A01 <- pars[1] 
    A02 <- pars[2] 
    logQn <- pars[3]
    sigma <- pars[4]
    
    A0mat <- swot_vec2mat(c(A01, A02), dA)
    A <- A0mat + dA
    
    dA0 <- 5 / (3 * sigma^2) * (-X + 5/3 * log(A) - logQn + sigma^2) / A
    dlogQn <- - sigma^-2 * (-X + 5/3 * log(A) - logQn)
    dsigma <- 1 / sigma - sigma^-3 * (-X + 5/3 * log(A) - logQn)^2
    
    grad <- c(sum(dA0[1, ]), sum(dA0[2, ]), sum(dlogQn), sum(dsigma))
    grad
  }
  
  attr(llfun, "gradfun") <- grfun
  llfun
}



```


Here's a version with a prior on sigma (half-gaussian):

```{r}
nllfun_v4.1 <-  function(swotlist, sigma_priorsd = 1) {
  W <- swotlist$W
  stopifnot(nrow(W) == 2)
  S <- swotlist$S
  X <- 2/3 * log(W) - 1/2 * log(S)
  dA <- swotlist$dA
  
  llfun <- function(pars) {
    A01 <- pars[1] 
    A02 <- pars[2] 
    logQn <- pars[3]
    sigma <- pars[4]
    
    if (sigma <= 0) return(Inf)
    
    A0mat <- swot_vec2mat(c(A01, A02), dA)
    A <- A0mat + dA
    
    mat1 <- (X - 5/3 * log(A) + logQn)^2 / (2 * sigma^2)
    mat2 <- 5/3 * log(A)
    const <- log(sqrt(2 * pi) * sigma)
    
    priorsum <- -dnorm(sigma, 0, sigma_priorsd, log = TRUE)
    
    out <- sum(mat1 + mat2 + const) + priorsum
    
    out
  }
  
  
  
  llfun
}

foo <- nllfun_v4.1(swot_sset(sscase, keeplocs = 1:2))
bar <- nllfun_v4(swot_sset(sscase, keeplocs = 1:2))
optim(par = c(1000, 1000, 4, 2), foo)
optim(par = c(1000, 1000, 4, 2), bar)
```

Great! After some minor debugging that seems to work like a charm! It still may be advantageous to manually specify the gradient, but wait until that becomes necessary. 

Now test on Pepsi cases.

```{r}
rd_maxlik <- reachdata %>% 
  map(swot_sset, keeplocs = 1:2) %>% 
  map(nllfun_v4) %>% 
  map(possibly(~optim(par = c(10000, 10000, 6, 3), fn = ., 
                      control = list(maxit = 1000)), otherwise = NA))

which(is.na(rd_maxlik))

rd_maxlik$Connecticut
rd_maxlik$MississippiDownstream

map_dbl(rd_maxlik[!is.na(rd_maxlik)], ~.$convergence)

rd_maxlik$SacramentoDownstream
```

Yeah, I should put in the gradient. OK, that should be done. Now test it out.

```{r}
bar <- nllfun_v4(swot_sset(reachdata$GaronneDownstream, keeplocs = 1:2), maxval = 1e10)
foo <- nllfun_v4.1(swot_sset(sscase, keeplocs = 1:2))
optim(par = c(1000, 1000, 4, 2), bar, control = list(maxit = 10000))
optim(par = c(1000, 1000, 4, 2), foo, control = list(maxit = 10000))


optim(par = c(1000, 1000, 4, 2), bar, method = "BFGS", 
      gr = attr(bar, "gradfun"),
      control = list(trace = 1))
optim(par = c(1000, 1000, 4, 2), bar, method = "BFGS", 
      # gr = attr(bar, "gradfun"),
      control = list(trace = 1))
optim(par = c(1000, 1000, 4, 2), bar, method = "CG", gr = attr(bar, "gradfun"),
      control = list(trace = 2))

optim(par = c(1000, 1000, 4, 2), fn = bar, gr = attr(bar, "gradfun"),
      method = "L-BFGS-B",
      control = list(trace = 3))


gfun <- attr(bar, "gradfun")
gfun(c(100, 100, 6, 10))

gfun(c(100, 100, 0, 0.4))


```

Alright, gradient doesn't help much, and in some cases hurts. Better let the machine do its thing. 


## Plotting functions

```{r}
swotlist <- reachdata$Po %>% 
  swot_sset(keeplocs = c(1, 5))
llfun <- nllfun_v4
A01lim <- A02lim <- c(300, 25000)
p <- c(10000, 10000, 10, 10)
log = FALSE

llplot1 <- function(swotlist, llfun, A01lim, A02lim, otherpars = "optim",
                    log = FALSE, 
                    p = c(10000, 10000, 10, 10), plot = TRUE, ...) {
  
  ll <- llfun(swotlist)
  
  if (otherpars == "optim") {
    optims <- optim(par = p, fn = ll, ...)
    if (!(optims$convergence == 0))
      stop("Optimization did not converge")
    otherpars <- optims$par[-1:-2]
  } 
  stopifnot(is.numeric(otherpars))
  
  
  if (log) {
    A01lim <- log(A01lim)
    A02lim <- log(A02lim)
    
    A01vals <- exp(seq(A01lim[1], A01lim[2], length.out = 100))
    A02vals <- exp(seq(A02lim[1], A02lim[2], length.out = 100))
  } else {
    A01vals <- seq(A01lim[1], A01lim[2], length.out = 100)
    A02vals <- seq(A02lim[1], A02lim[2], length.out = 100)
  }
  

  
  pargrid <- expand.grid(A01 = A01vals, A02 = A02vals)
  pargrid$logQn <- otherpars[1]
  pargrid$sigma <- otherpars[2]
  
  parlist <- split(pargrid, f = 1:nrow(pargrid)) %>% 
    map(unlist)
  nllvec <- map_dbl(parlist, ll)
  
  pargrid$nll <- nllvec
  
  if (!plot) {
    return(pargrid)
  }
  
  out <- ggplot(pargrid, aes(x = A01, y = A02, z = nll, fill = nll)) +
    geom_raster() +
    geom_contour() +
    scale_fill_viridis_c()
  
  if (log)
    out <- out + scale_x_log10() + scale_y_log10()

  out
}



llplot1(swotlist, nllfun_v4, A01lim, A02lim, log = FALSE, method = "BFGS")
llplot1(swotlist, nllfun_v4, A01lim, A02lim, log = TRUE, method = "BFGS")

```

Now make some plots for various cases on hand. 

```{r}
getpars <- function(swotlist) {
  W <- swotlist$W
  stopifnot(nrow(W) == 2)
  S <- swotlist$S
  X <- 2/3 * log(W) - 1/2 * log(S)
  dA <- swotlist$dA
  A <- swotlist$A
  
  A0 <- (A - dA)[, 1]
  
  rhs <- log(W^(-2/3) * A^(5/3) * S^(1/2))
  
  logQn <- mean(rhs)
  sigma <- sd(rhs)
  
  out <- list(A0 = A0, logQn = logQn, sigma = sigma)
  out
}

sscase %>% 
  swot_sset(keeplocs = c(1, 3)) %>% 
  getpars

ssgg <- sscase %>% 
  swot_sset(keeplocs = c(1, 3)) %>% 
  llplot1(llfun = nllfun_v4, 
          # otherpars = c(1.709, 0.56),
          A01lim = c(30, 500), A02lim = c(30, 500), 
          p = c(300, 300, 5, 4), 
          method = "BFGS",
          control = list(maxit = 10000, trace = 2),
          plot = TRUE)
ssgg

```

Curious that sscase doesn't converge. Not enough times? 

Returning now to Po case, now that I can manually specify other params. 

```{r}
foocase <- reachdata$Po %>% 
  swot_sset(keeplocs = c(1, 5))
getpars(foocase)
llplot1(foocase, nllfun_v4, 
        A01lim = c(300, 2000), A02lim = c(300, 3000), 
        log = FALSE, otherpars = c(3.336, 0.573))
```

That looks to be perfect! So if I can get the other parameters, then it looks like I can estimate A0. But can I get the other parameters? I should make a nll function that has only a single A0, plus logQn and sigma. 

```{r}
# This version only uses a single A0. Gradient hasn't been verified. 
nllfun_v4.2 <-  function(swotlist, maxval = Inf) {
  W <- swotlist$W
  S <- swotlist$S
  X <- 2/3 * log(W) - 1/2 * log(S)
  dA <- swotlist$dA
  
  llfun <- function(pars) {
    A0 <- pars[1] 
    # A02 <- pars[2] 
    logQn <- pars[2]
    sigma <- pars[3]
    
    if (sigma <= 0) return(maxval)
    
    A <- A0 + dA
    
    mat1 <- (X - 5/3 * log(A) + logQn)^2 / (2 * sigma^2)
    mat2 <- 5/3 * log(A)
    const <- log(sqrt(2 * pi) * sigma)
    
    out <- sum(mat1 + mat2 + const)  
    out
  }
  
  grfun <- function(pars) {
    A0 <- pars[1] 
    # A02 <- pars[2] 
    logQn <- pars[2]
    sigma <- pars[3]
    
    A <- A0 + dA
    
    dA0 <- 5 / (3 * sigma^2) * (-X + 5/3 * log(A) - logQn + sigma^2) / A
    dlogQn <- - sigma^-2 * (-X + 5/3 * log(A) - logQn)
    dsigma <- 1 / sigma - sigma^-3 * (-X + 5/3 * log(A) - logQn)^2
    
    grad <- c(sum(dA0), sum(dlogQn), sum(dsigma))
    grad
  }
  
  attr(llfun, "gradfun") <- grfun
  llfun
}

foofun <- nllfun_v4.2(swotlist = swot_sset(reachdata$Po, keeplocs = -14))
optim(par = c(3900, 10, 10), fn = foofun, 
      # gr = attr(foofun, "gradfun"),
      method = "BFGS",
      control = list(maxit = 10000))
```

Try putting a prior on sigma. 

```{r}
nllfun_v4.3 <-  function(swotlist, sigma_priorsd = 1, maxval = Inf) {
  W <- swotlist$W
  S <- swotlist$S
  X <- 2/3 * log(W) - 1/2 * log(S)
  dA <- swotlist$dA
  
  llfun <- function(pars) {
    A0 <- pars[1] 
    # A02 <- pars[2] 
    logQn <- pars[2]
    sigma <- pars[3]
    
    if (sigma <= 0) return(maxval)
    
    A <- A0 + dA
    
    mat1 <- (X - 5/3 * log(A) + logQn)^2 / (2 * sigma^2)
    mat2 <- 5/3 * log(A)
    const <- log(sqrt(2 * pi) * sigma)
 
    priorsum <- -dnorm(sigma, 0, sigma_priorsd, log = TRUE)
 
    out <- sum(mat1 + mat2 + const) + priorsum
    out
  }
  
  llfun
}

barfun <- nllfun_v4.3(swotlist = swot_sset(reachdata$Po, keeplocs = -14))
optim(par = c(3900, 10, 10), fn = barfun, 
      # gr = attr(foofun, "gradfun"),
      method = "BFGS",
      control = list(maxit = 10000))
```

Now make a plot showing tradeoff of A0 vs logQn. 


## EM algorithm approach

I should have done this earlier!

```{r}
casei <- reachdata$Po
# casei <- sscase


logW <- log(casei$W)
logS <- log(casei$S)
dA <- casei$dA
minA0 <- -apply(dA, 1, min) + 1
 
A0i <- minA0 * 1 # initialize A0 vector

for (i in 1:1000) {
  Ai <- casei$dA + swot_vec2mat(A0i, casei$dA)
  rhsi <- -2/3 * logW + 1/2 * logS + 5/3 * log(Ai)
  
  logQtni <- apply(rhsi, 2, mean)
  # logQni <- mean(logQtni)
  sigma1i <- apply(rhsi, 2, sd)
  # sigma2i <- sd(sigma1i)
  
  mu_Ati <- 3/5 * (swot_vec2mat(logQtni, dA) + 2/3 * logW - 1/2 * logS)
  sd_Ati <- 3/5 * sigma1i
  # sd_A <- sd(log(Ai))
  
  
  Anew <- exp(mu_Ati + sd_Ati^2 / 2)
  # Anew <- exp(mu_Ati)
  
  A0new <- Anew - dA
  
  badinds <- A0new < swot_vec2mat(minA0, A0new)
  A0new[badinds] <- swot_vec2mat(minA0, A0new)[badinds]
  A0i <- apply(A0new, 1, mean)
  
  # A0new <- apply(Anew - dA, 1, mean)
  # A0i <- map2_dbl(minA0, A0new, max)
  
  
  lli <- map2_dbl(split(as.data.frame(t(mu_Ati)), 1:ncol(mu_Ati)), 
              sd_Ati, ~sum(dnorm(unlist(.x), 0, .y, log = TRUE))) %>% 
    sum()
  
  if (i %% 10 == 0) cat(i, mean(A0i), lli, "\n")
  # A0i
}

```

Doesn't look like straight-up EM is going to work with analytical maximization. I can, however, do univariate optimization within EM via optimize(). That would be a fun function to write. 

```{r}

casei <- reachdata$Po %>%
  swot_sset (keeplocs = -14)
# casei <- sscase


logW <- log(casei$W)
logS <- log(casei$S)
X <- -2/3 * logW + 1/2 * logS
dA <- casei$dA
minA0 <- -apply(dA, 1, min) + 1
 
A0i <- minA0 * 10 # initialize A0 vector

for (i in 1:10000) {

  # Calculate logQtn
  Ai <- casei$dA + swot_vec2mat(A0i, casei$dA)
  rhsi <- X + 5/3 * log(Ai)
  logQtni <- apply(rhsi, 2, mean)
  # sigma1i <- apply(rhsi, 2, sd)
  
  # Function to optimize over A0
  nllfun <- function(A0, Xrow, dArow) {
    A <- dArow + A0
    rhs <- Xrow + 5/3 * log(A)
    
    # out <- sum((logQtni - rhs)^2 / (2 * mean(sigma1i^2)))
    
    out <- mean((logQtni - rhs)^2)
    out
  }
  
  optimsi <- data.frame(A0 = numeric(length(A0i)), sigsq = numeric(length(A0i)))
  for (r in 1:nrow(logW)) {
    optimr <- optimize(nllfun, interval = c(minA0[r], A0i[r] * 100),
                       Xrow = X[r, ], dArow = dA[r, ])
    optimsi$A0[r] <- optimr[["minimum"]]
    optimsi$sigsq[r] <- optimr[["objective"]]
  }
  
  A0i <- optimsi$A0
  
  if (i %% 100 == 0)
    cat(i, mean(optimsi$A0), mean(optimsi$sigsq), "\n")
}


```

Looks like it works!

```{r}
plot(A0i, apply(casei$A, 1, median)); abline(0, 1)
plot(estA0(casei), apply(casei$A, 1, median)); abline(0, 1)
```



```{r}
rhsmat <- (X + 5/3 * log(swot_vec2mat(A0i, X) + dA)) %>% 
  exp()

truerhsmat <- (X + 5/3 * log(casei$A)) %>% 
  exp()

plot_DAWG(rhsmat)
plot_DAWG(truerhsmat)

plot_DAWG(rhsmat - swot_vec2mat(apply(rhsmat, 2, mean), rhsmat))
plot_DAWG(truerhsmat - swot_vec2mat(apply(truerhsmat, 2, mean), rhsmat))


apply(rhsmat, 2, sd) %>% mean()
apply(truerhsmat, 2, sd) %>% mean()

apply(rhsmat, 2, sd) %>% max()
apply(truerhsmat, 2, sd) %>% max()

apply(rhsmat, 2, sd) %>% plot()
apply(truerhsmat, 2, sd) %>% plot()

```

Works for sscase, but not for Po. Po pushes everything too high, because this tightens the rhs matrix beyond where it should reasonbly go. My instinct is that this needs to be fully Bayesian, since a max likelihood will always tighten the mass conservation too much. Let's take a break, then come back and work on gathering and tidying work done so far. 

See what happens when I minimize in cartesian space.

```{r}

casei <- reachdata$Po %>%
  swot_sset (keeplocs = -14)
# casei <- sscase


logW <- log(casei$W)
logS <- log(casei$S)
X <- -2/3 * logW + 1/2 * logS
dA <- casei$dA
minA0 <- -apply(dA, 1, min) + 1
 
A0i <- minA0 * 1 # initialize A0 vector

for (i in 1:10000) {

  # Calculate logQtn
  Ai <- casei$dA + swot_vec2mat(A0i, casei$dA)
  rhsi <- X + 5/3 * log(Ai)
  logQtni <- apply(exp(rhsi), 2, mean)
  # sigma1i <- apply(rhsi, 2, sd)
  
  # Function to optimize over A0
  nllfun <- function(A0, Xrow, dArow) {
    A <- dArow + A0
    rhs <- Xrow + 5/3 * log(A)
    
    # out <- sum((logQtni - rhs)^2 / (2 * mean(sigma1i^2)))
    
    out <- mean((logQtni - exp(rhs))^2)
    out
  }
  
  optimsi <- data.frame(A0 = numeric(length(A0i)), sigsq = numeric(length(A0i)))
  for (r in 1:nrow(logW)) {
    optimr <- optimize(nllfun, interval = c(minA0[r], A0i[r] * 100),
                       Xrow = X[r, ], dArow = dA[r, ])
    optimsi$A0[r] <- optimr[["minimum"]]
    optimsi$sigsq[r] <- optimr[["objective"]]
  }
  
  A0i <- optimsi$A0
  
  if (i %% 100 == 0)
    cat(i, mean(optimsi$A0), mean(optimsi$sigsq), "\n")
}

```

Converges right away, which makes sense. Now show that I can use the same approach to get the A0est result. 

```{r}
casei <- reachdata$Po %>%
swot_sset (keeplocs = -14)
# casei <- sscase


W <- casei$W
S <- casei$S
X <- W^(-2/5) * S^(3/10) #* 1000
dA <- casei$dA
minA0 <- -apply(dA, 1, min) + 1
 
A0i <- minA0 * rnorm(length(minA0), 10, 0.2) # initialize A0 vector

for (i in 1:1000) {

  Ai <- dA + swot_vec2mat(A0i, dA)
  Qtni35 <- apply(Ai * X, 2, mean)
  
  
  parenmat <- Ai * X - swot_vec2mat(Qtni35, X)
  
  # plot(Qtni35, type = "l")
  
  tss1 <- sum((Ai * X - swot_vec2mat(Qtni35, X))^2)
  
  
  dAXbar <- apply(dA * X, 1, mean)
  Xbar <- apply(X, 1, mean)
  
  A0i <- (mean(Qtni35) - dAXbar) / Xbar
  # A0i <- (mean(Qtni35) - Xbar * apply(dA, 1, mean)) / Xbar
  
  Qtnmat <- swot_vec2mat(Qtni35, dA)
  
  # A0i <- apply(Qtnmat / X - dA, 1, mean)
  A0i <- map2_dbl(A0i, minA0, max)
  
  tss2 <- sum(((dA + swot_vec2mat(A0i, X)) * X - swot_vec2mat(Qtni35, X))^2)
  
  if (i %% 100 == 0)
    cat(i, mean(A0i), mean(Qtni35), tss1, tss2, "\n")
}

plot(Qtni35)


# emsd <- apply(swot_vec2mat(Qtni35, X) - (Ai * X), 2, sd)
# 
# Ai_lm <- swot_vec2mat(estA0(casei), X) + casei$dA
# 
# lmsd <- apply(swot_vec2mat(Qtni35, X) - (Ai_lm * X), 2, sd)
# 
# plot(emsd, lmsd); abline(0, 1)

```

It converges as long as it starts in clos to the right place, but doesn't agree with A0est.

```{r}
plot_DAWG(Ai * X)

plot_DAWG(casei$A * X)

lmmat <- (swot_vec2mat(estA0(casei), X) + casei$dA) * X
plot_DAWG(lmmat)

apply(lmmat, 2, sd) %>% mean()
apply(Ai * X, 2, sd) %>% mean()

```


Try a gradient version.

```{r}

linA_llfun <- function(swotlist) {
  W <- swotlist$W
  S <- swotlist$S
  X <- W^(-2/5) * S^(3/10)
  dA <- swotlist$dA
  minA0 <- -apply(dA, 1, min) + 1
  
  objfun <- function(Qn, A0) {
    
    # if (sum(A0 < minA0) > 0) return(1e15)
    # if (sum(Qn <= 0) > 0) return(1e15)
    
    A <- dA + swot_vec2mat(A0, dA)
    Qn35 <- swot_vec2mat(Qn, dA)^(3/5)
    objmat <- (A * X - Qn35)^2
    obj <- sum(objmat)
    obj
  }
  gradfun <- function(Qn, A0) {
    A <- dA + swot_vec2mat(A0, dA)
    Qnmat <- swot_vec2mat(Qn, dA)
    Qngradmat <- 2 * (-3/5) * Qnmat^(-2/5) * (A * X - Qnmat^(3/5))
    A0gradmat <- 2 * X * (A * X - Qnmat^(3/5))
    
    Qngrad <- apply(Qngradmat, 2, sum)
    A0grad <- apply(A0gradmat, 1, sum)
    
    # Qngrad[Qn <= 0] <- -1e5
    # A0grad[A0 < minA0] <- -1e5
    
    out <- list(Qn = Qngrad, A0 = A0grad) 
    out
  }
  
  out <- list(objfun = objfun, gradfun = gradfun)
}




gradclimb <- function(objfun, gradfun, initQn, initA0, alpha = 1) {
  
}

```

```{r}
casei <- sscase
ofi <- linA_llfun(casei)

initQn <- rep(500 * 0.03, ncol(casei$W))
initA0 <- rep(30, nrow(casei$W))


Qn <- initQn
A0 <- initA0
alpha <- 1
minQn <- 1 * 0.02
minA0 <- -apply(casei$dA, 1, min) + 1

for (i in 1:34) {
  gradi <- ofi$gradfun(Qn = Qn, A0 = A0)
  newQn <- Qn - gradi$Qn * alpha
  
  # newQn[newQn < minQn] <- minQn
  
  gradi <- ofi$gradfun(Qn = newQn, A0 = A0)
  newA0 <- A0 - gradi$A0 * alpha
  

  
  Qn <- newQn
  # Qn[Qn < minQn] <- minQn
  A0 <- newA0
  # A0[A0 < minA0] <- minA0
  
  # if (i %% 10 == 0) 
    cat(i, ofi$objfun(Qn, A0), mean(Qn), mean(A0), "\n")
}



```

One where parameter is transformed Qn--avoids negative issue

```{r}
linA_llfun2 <- function(swotlist) {
  W <- swotlist$W
  S <- swotlist$S
  X <- W^(-2/5) * S^(3/10)
  dA <- swotlist$dA
  minA0 <- -apply(dA, 1, min) + 1
  
  objfun <- function(Qn35, A0) {
    
    # if (sum(A0 < minA0) > 0) return(1e15)
    # if (sum(Qn <= 0) > 0) return(1e15)
    
    A <- dA + swot_vec2mat(A0, dA)
    Qn35mat <- swot_vec2mat(Qn35, dA)
    objmat <- (A * X - Qn35mat)^2
    obj <- sum(objmat)
    obj
  }
  gradfun <- function(Qn35, A0) {
    A <- dA + swot_vec2mat(A0, dA)
    Qn35mat <- swot_vec2mat(Qn35, dA)
    Qngradmat <- -2 * (A * X - Qn35mat)
    A0gradmat <- 2 * X * (A * X - Qn35mat)
    
    Qngrad <- apply(Qngradmat, 2, sum)
    A0grad <- apply(A0gradmat, 1, sum)
    
    # Qngrad[Qn <= 0] <- -1e5
    # A0grad[A0 < minA0] <- -1e5
    
    out <- list(Qn35 = Qngrad, A0 = A0grad) 
    out
  }
  
  out <- list(objfun = objfun, gradfun = gradfun)
}

```



```{r}
casei <- sscase
ofi <- linA_llfun2(casei)

initQn <- rep(500 * 0.03, ncol(casei$W))
initA0 <- rep(300, nrow(casei$W))

Qn35 <- initQn^(3/5)
A0 <- initA0
alpha <- 1
minQn <- 1 * 0.02
minA0 <- -apply(casei$dA, 1, min) + 1

oldobj <- ofi$objfun(Qn35 = Qn35, A0 = A0)

for (i in 1:80) {
  gradi <- ofi$gradfun(Qn35 = Qn35, A0 = A0)
  newQn35 <- Qn35 - gradi$Qn35 * alpha
  
  if (sum(is.na(newQn35)) > 0) break()
  
  # newQn[newQn < minQn] <- minQn
  
  # gradi <- ofi$gradfun(Qn35 = newQn35, A0 = A0)
  newA0 <- A0 - gradi$A0 * alpha
  
  Qn35 <- newQn35
  # Qn[Qn < minQn] <- minQn
  A0 <- newA0
  # A0[A0 < minA0] <- minA0
  
  newobj <- ofi$objfun(Qn35 = Qn35, A0 = A0)
  if (newobj > oldobj) {
    # break
    alpha <- alpha * 0.5
    next
  } else {
    alpha <- alpha * 1.2
  }
  
  oldobj <- newobj
  # if (i %% 10 == 0) 
    cat(i, newobj, mean(Qn35), mean(A0), alpha,  "\n")
}

Ai <- swot_vec2mat(A0, casei$dA) + casei$dA
Xi <- casei$W^(-2/5) * casei$S^(3/10)
rhsi <- Ai * Xi

plot_DAWG(rhsi)

newgrad <- ofi$gradfun(A0, Qn35 = Qn35)


```

This still isn't making sense. I need a version that I can let R optimize. 

```{r}
linA_llfun3 <- function(swotlist) {
  W <- swotlist$W
  S <- swotlist$S
  X <- W^(-2/5) * S^(3/10)
  dA <- swotlist$dA
  minA0 <- -apply(dA, 1, min) + 1
  
  objfun <- function(pars) {
    
    # if (sum(A0 < minA0) > 0) return(1e15)
    # if (sum(Qn <= 0) > 0) return(1e15)
    
    Qn35 <- pars[1:9]
    A0 <- pars[10:12]
    
    A <- dA + swot_vec2mat(A0, dA)
    Qn35mat <- swot_vec2mat(Qn35, dA)
    objmat <- (A * X - Qn35mat)^2
    obj <- sum(objmat)
    obj
    
    
    
  }
  out <- objfun
}

objfun3 <- linA_llfun3(sscase)

inits <- c(setNames(rep(500, 9), paste0("mu", 1:9)), 
           setNames(rep(30, 3), paste0("A0", 1:3)))
foo <- optim(par = inits, objfun3, method = "BFGS", control = list(maxit = 500))
foo <- optim(par = inits, objfun3, method = "CG", 
             control = list(maxit = 50000))
bar <- nlm(objfun3, p = inits, method = )
```

So that works! Why doesn't my gradient method? And why doesnt' my EM method?

- It's got to be numerical. I can show that the solution they converge to has a small gradient--it's just not optimal. 

How about a general-purpose minimizable function? I think I have an idea how.

```{r}
llfun4 <- function(swotlist) {
  W <- swotlist$W
  S <- swotlist$S
  X <- W^(-2/5) * S^(3/10)
  dA <- swotlist$dA
  minA0 <- -apply(dA, 1, min) + 1
  
  times <- 1:ncol(X)
  reaches <- 1:nrow(X)
  
  objfun <- function(pars) {
    
    # if (sum(A0 < minA0) > 0) return(1e15)
    # if (sum(Qn <= 0) > 0) return(1e15)
    
    Qn35 <- pars[times]
    A0 <- pars[max(times) + reaches]
    
    A <- dA + swot_vec2mat(A0, dA)
    Qn35mat <- swot_vec2mat(Qn35, dA)
    objmat <- (A * X - Qn35mat)^2
    obj <- sum(objmat)
    obj
    
  }
  
  initA0 <- apply(dA, 1, function(x) max(x) - min(x))
  
  initQn35 <- apply(X * (dA + swot_vec2mat(initA0, dA)), 2, mean)

  out <- list(objfun = objfun, inits = c(initQn35, initA0))
}

casei <- reachdata$Po
llfi <- llfun4(casei)

optimi <- nlm(llfi$objfun, p = llfi$inits)

```

Works! nlm looks like the way to go. 

Next: 

- Meet Lark for lunch
- Write up lisflood cases for blog
- Make these results into functions, document
- Package swotr functions, put on github