---
title: "notebook20180419"
author: "Mark Hagemann"
date: "April 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Since I can't get an intuitive or mathematical understanding of correlation and other error structure in uscase, pepsi cases, it's time to go ahead with the simulation / sensitivity analysis. 

- split closure term into gamma, nu. 
- Split gamma into constant dq/dx and other
- split nu into time (depth) and space components
- Any further decomposition?
    - I still need to characterize this stuff!

Also, swot_gamma needs to be changed to manning_gamma. In fact, I think it's time to put all the closure funcitons into lib/.

```{r}
casei <- uscase
casei <- reachdata$MississippiDownstream

gammai <- mannning_gamma(casei)

plot_DAWG(gammai)

dgi <- decomp_gamma(gammai, casei$x)

plot_DAWG(dgi$gammahat)
plot_DAWG(dgi$gammaerr)


```

Does nu calculation work via manning_ndot? Nope, that's just the space-vector of means. How about by manning_closure?

```{r}
nui <- manning_closure(casei, log = TRUE, mc = FALSE) %>% 
  `-`(mean(.))

plot_DAWG(nui)

closi <- manning_closure(casei, log = TRUE, mc = TRUE)
plot_DAWG(closi)
plot_DAWG(closi - gammai)
plot_DAWG(manning_nu(casei) %>% exp())

dni <- decomp_nu(manning_nu(casei))

plot_DAWG(dni$nuhat %>% exp())
plot_DAWG(dni$nuerr %>% exp())

plot(casei$A / casei$W, dni$nuerr)

plot_DAWG(findif_x(dni$nuerr))

cor(t(dni$nuerr))
```

There is correlation between these different tau timeseries, but I can't account for it. 

```{r}
val_nd_plot(casei)
```

Looking at what I have so far, I need to modify:

- temporal variability of dgamma/dx - sigma_dgdx
- spatial variability of nu - sigma_alpha
- magnitude of residuals - sigma_epsilon

Approach:

- Specify magnitudes of errors described above
- Decompose case into these components
- Modify components accordingly 
- Reassemble
- estimate A0

```{r}
# Characterize
characterize_closure <- function(swotlist) {
  gma <- manning_gamma(swotlist)
  gd <- decomp_gamma(gma, swotlist$x)
  
  sig_dgdx <- sd(gd$dgdx)
  
  nu <- manning_nu(swotlist)
  nd <- decomp_nu(nu)
  
  sig_alpha <- sd(nd$nuhat[, 1])
  sig_err <- sd(gd$gammaerr + nd$nuerr)
  
  out <- c(dgdx = sig_dgdx, alpha = sig_alpha, err = sig_err, 
           dx = sd(swotlist$x[, 1]))
  out
}

plot_DAWG(gd$gammaerr + nd$nuerr)
plot_DAWG(nd$nuerr)
plot_DAWG(gd$gammaerr)
plot_DAWG(gd$gammahat)
sd(gd$gammahat)
```

How do I relate sig_dgdx to sig_gammahat? Need mean of dgdx as well. But this is zero if dist'n is symmetric, and I'm modeling it as normal, so it's symmetric. 

Anyway, the answer is it's the product of mean(dgdx^2) and mean(deltax^2). Show that it is. 

```{r}
mean(gd$dgdx^2) * mean((findif_x(casei$x)[, 1])^2)

sd(gd$gammahat)^2

mean(gd$gammahat^2)

mean((swot_vec2mat(gd$dgdx, gd$gammahat) * 
        swot_vec2mat(findif_x(casei$x)[, 1], gd$gammahat))^2)
mean((swot_vec2mat(gd$dgdx, gd$gammahat) * 
        swot_vec2mat(findif_x(casei$x)[, 1], gd$gammahat)))^2
```


Close, but not perfect.

Bah, right. deltax isn't the same as finite difference. 


```{r}
xlocs <- casei$x[, 1]
mean(gd$dgdx^2) * mean((xlocs - mean(xlocs))^2)
```

There it is. Good. Now finish characterie_closure function.

```{r}
characterize_closure(casei)
```

Now what? Do a test case where I modify these and recreate the case. 

```{r}
newdgdx <- 1.6e-5
newalpha <- 4e-3
newerr <- 2e-2

modify_closure <- function(swotlist, sigma_dgdx, sigma_alpha, sigma_err) {
  
  ptrn <- swotlist$W
  
  xlocs <- swotlist$x[, 1]
  xdevs <- xlocs - mean(xlocs)
  xdevs_mat <- swot_vec2mat(xdevs, ptrn)
  
  nx <- nrow(ptrn)
  nt <- ncol(ptrn)
  
  sim_dgdx <- rnorm(nt, 0, sigma_dgdx)
  gammahat <- swot_vec2mat(sim_dgdx, ptrn) * xdevs_mat
  
  sim_alpha <- rnorm(nx, 0, sigma_alpha)
  nuhat <- swot_vec2mat(sim_alpha, ptrn)
  
  sim_err <- rnorm(nx * nt, 0, sigma_err)
  
  out <- gammahat + nuhat + matrix(sim_err, nr = nx)
  out
}

plot_DAWG(gammahat)


newclosi <- modify_closure(casei, sigma_dgdx = newdgdx, 
                        sigma_alpha = newalpha, sigma_err = newerr)


```

Now need to make something to go into estA0_linA_ws35.

```{r}
newWS35 <- manning_linA_closed(casei) * exp(newclosi)

plot_DAWG(newWS35)

estA0_lm_ws35(newWS35, dAmat = casei$dA)

```

I should express dgdx in a more intuitive unit. Like percent per reach. 


