---
title: "notebook20180313"
author: "Mark Hagemann"
date: "March 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Much of today was continuing yesterday's work in yesterday's notebook. 

Following skype call with Mike and Steve: look at Willamette data from A0 inversion perspective

```{r}
hc_h <- read.csv("../data/HydraulicControlReaches/H.txt", sep = "\t", header = FALSE) %>% 
  as.matrix()
hc_w <- read.csv("../data/HydraulicControlReaches/W.txt", sep = "\t", header = FALSE) %>% 
  as.matrix()
hc_s <- read.csv("../data/HydraulicControlReaches/S.txt", sep = "\t", header = FALSE) %>% 
  as.matrix()
hc_dA <- calcdA_mat(hc_w, hc_h)
hc_dA_median <- matrix(rep(apply(hc_dA, 1, median), ncol(hc_dA)), ncol = ncol(hc_dA), byrow = FALSE)
hc_dA_minimum <- matrix(rep(apply(hc_dA, 1, min), ncol(hc_dA)), ncol = ncol(hc_dA), byrow = FALSE)
hc_dA2 <- hc_dA - hc_dA_median
hc_dA3 <- hc_dA - hc_dA_minimum

summary(t(hc_h))
summary(t(hc_s))
summary(t(hc_w))
summary(t(hc_dA))
summary(t(hc_dA2))
summary(t(hc_dA3))
```

Here goes!

```{r}
A0est_med <- estA0(wmat = hc_w, smat = hc_s, damat = hc_dA2)
A0est_min <- estA0(wmat = hc_w, smat = hc_s, damat = hc_dA3)
summary(A0est_med)
summary(A0est_min)

coef(A0est_med)
coef(A0est_med) - apply(hc_dA3, 1, median)

coef(A0est_min) + apply(hc_dA3, 1, median)

plot(coef(A0est_med), coef(A0est_min) + apply(hc_dA3, 1, median))
abline(0, 1)

termplots <- map(paste0("x", 1:16), ~ggTermPlot(foo, xvar = .))
termplots[[4]]

A0hats <- coef(foo) 

plot(apply(hc_w, 1, median), A0hats, xlab = "Reach width (m)", ylab = "A0 estimate (m^2)")
abline(0, 1)
plot(apply(hc_s, 1, median), A0hats, xlab = "Median reach width (unitless)", ylab = "A0 estimate (m^2)")

```


PCA on data matrix

```{r}
bar <- eigen(cor(foo$model[, -1]))
bar$values
bar$vectors[, 16]
```

Hatvalues of matrix

```{r}
foox <- as.matrix(foo$model[,-1])
solve(t(foox) %*% foox)


```



I think that's adequate for today. To be revisited in the future.

Now on to playing with lisflood parameters.

- Previously I did this in airSWOT notebook20180307.Rmd.

```{r}
read_lisProfile <- function(file) {
  infile <- read_fwf(file, 
                     fwf_empty(file, 
                               c("ChanX", "ChanY", "Chainage", "Width", 
                                 "Mannings", "Slope", "BankZ", "BedElev", 
                                 "WaterElev", "WaterDepth", "Flow"), skip = 2), 
                     skip = 2, col_types = "ddddddddddd")
  infile
}


```

I need to look at the relationship between inflow Q and outflow H so I can modify the boundary condition. 

```{r}
qin <- read.csv("../lisflood/toy_1/bdy_qin.tsv", sep = "\t")
hout <- read.csv("../lisflood/toy_1/bdy_hout.tsv", sep = "\t")

bdy <- left_join(qin, hout, by = "time_s")


bdy %>% 
  ggplot(aes(x = h_m, y = q_m3s)) + 
  geom_point()
```

Based on this, use the following heights and flows:

```{r}
q_impose <- c(50, 75, 100, 125, 150, 175, 200, 250, 300)

hfun <- approxfun(x = bdy$q_m3s, y = bdy$h_m)

h_impose <- hfun(q_impose)
h_impose
```

Now make into data.frame that I can write and easily coerce into lisflood format. 

```{r}
plot(bdy$time_s)
```

Looks like the thing to do would be to do a step function for Q and H. 

```{r}
time_sustain <- 86400 # How long to stay at each steady-flow condition
time_change <- 10368 # How long to take to transition to next steady-state condition

q_write <- rep(q_impose, each = 2)
h_write <- rep(h_impose, each = 2)
t_write <- cumsum(c(0, rep(c(time_sustain, time_change), 
                           length.out = length(h_write)))) %>% 
  `[`(1:length(h_write))

plot(t_write, q_write, type = "l")
```

Looks good! Write it. 

```{r}
library(readr)
write_tsv(data.frame(q_m3s = q_write, t_s = t_write), 
          path = "../lisflood/q_ss.tsv")

write_tsv(data.frame(h_m = h_write, t_s = t_write), 
          path = "../lisflood/h_ss.tsv")
```


Done. Now run it in Lisflood. Done. Now read the results.

```{r}
prof_const <- read_lisProfile(file = "../lisflood/toy_1/results/res-river0-0006.profile")

allprof_files <- list.files("../lisflood/toy_1/results/", pattern = "\\.profile$", 
                            full.names = TRUE)
allprof_const <- map(allprof_files, read_lisProfile) %>% 
  setNames(1:length(allprof_files)) %>% 
  bind_rows(.id = "time") %>% 
  mutate(time = as.numeric(time))


allprof_const %>% 
  filter(ChanX == 2100) %>% 
  glimpse() %>% 
  ggplot(aes(x = time, y = Slope)) +
  geom_point()

allprof_const %>% 
  filter(time > 2) %>%
  glimpse() %>% 
  ggplot(aes(x = time, y = ChanX, fill = Flow)) +
  geom_raster() +
  scale_fill_gradient(low = "#ff0000", high = "#3300ff")

allprof_const$Flow
```

