---
title: "notebook20180614"
author: "Mark Hagemann"
date: "June 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Going with original bam files for starters. 




```{r}
library(batchtools)

oldreg <- loadRegistry("../../SWOT/src/bam_experiments/runs3/reg_reach_manning")

getJobStatus(1:20, oldreg)

```

So in order to use the model from the paper I'd need to go to bamr version 0.0.12. 

- Commit: be5e9375c42d65890bf503b32619884c11c4ffe1

I could go back further to original files in SWOT/src/defunct.

```{r}
oldreslist <- map(1:19, ~loadResult(id = ., reg = oldreg))

oldreslist[[1]] %>% 
  get_posterior_mean() %>% 
  rownames()


pairs(oldreslist[[2]], pars = c("logQ[1]", "logn", "A0[1]", "A0[2]"))
```

Trying to reproduce these tighter distributions using vectorized qhat.

```{r}
paremod <- stan_model("../src/manning_pared.stan")
paremod_vec <- stan_model("../src/manning_pared_vecq.stan")


paresamps1 <- baminps$Po %>% 
  pare_baminps() %>% 
  sampling(paremod, data = ., chains = 3, cores = 3, iter = 2000)

paresamps2 <- baminps$Po %>% 
  pare_baminps() %>% 
  sampling(paremod_vec, data = ., chains = 3, cores = 3, iter = 2000)

paresamps3 <- baminps$Po %>% 
  within(., {logA0_sd = 0.1}) %>% 
  pare_baminps() %>% 
  sampling(paremod_vec, data = ., chains = 3, cores = 3, iter = 2000)

```



```{r}

get_posterior_mean(paresamps1, pars = "A0")
stan_trace(paresamps1, pars = "A0")
get_posterior_mean(paresamps1, pars = c("truesigma_err", "sigma_y"))
pairs(paresamps1, pars = c("truesigma_err", "sigma_y", "mu", "A0[1]", "A0[2]"))
check_hmc_diagnostics(paresamps1)
check_rhat(paresamps1)

get_posterior_mean(paresamps2, pars = "A0")
stan_trace(paresamps2, pars = "A0")
get_posterior_mean(paresamps2, pars = c("truesigma_err", "sigma_y"))
pairs(paresamps2, pars = c("truesigma_err", "sigma_y", "mu", "A0[1]", "A0[2]"))
check_hmc_diagnostics(paresamps2)
check_rhat(paresamps2)

get_posterior_mean(paresamps3, pars = "A0")
stan_trace(paresamps3, pars = "A0")
get_posterior_mean(paresamps3, pars = c("truesigma_err", "sigma_y"))
pairs(paresamps3, pars = c("truesigma_err", "sigma_y", "mu", "A0[1]", "A0[2]"))
check_hmc_diagnostics(paresamps3)
check_rhat(paresamps3)
```

Verily, it does impact the tightness of the posterior. 


I'm now wondering whether I get the same correlation structure in my matrix version. 

```{r}
lmmod <- stan_model("../stan/A0_v2-1.stan")

lminps0 <- baminps$Po %>% 
  within(., {logA0_sd = 0.1})

omg <- make_omegas(lminps0$nx) %>% 
  as.data.frame()
lminps <- within(lminps0, {
  x = 1:nx
  omega = t(omg)
  dA = t(dAobs)
  W = t(Wobs)
  S = t(Sobs)
  logA0_hat = mean(logA0_hat)
  sigma_logA0 = logA0_sd
  sigma_err = 0.1 * (max(dA) - min(dA))
})

lmopt1 <- optimizing(lmmod, lminps, as_vector = FALSE)
lmopt1$par$A0

lmsamps <- sampling(lmmod, lminps, cores = 3, chains = 3, iter = 2000)
check_hmc_diagnostics(lmsamps)

get_posterior_mean(lmsamps)
pairs(lmsamps, pars = c("A0[1]", "A0[2]", "A0[3]", "truesigma_err"))
```

Does sampling work if I substantially loosen the prior on A0?

```{r}
lminps2 <- within(lminps0, {
  x = 1:nx
  omega = t(omg)
  dA = t(dAobs)
  W = t(Wobs)
  S = t(Sobs)
  logA0_hat = mean(logA0_hat)
  sigma_logA0 = 2
  sigma_err = 0.1 * (max(dA) - min(dA))
})

lmopt2 <- optimizing(lmmod, lminps2, as_vector = FALSE)
lmopt2$par$A0

lmsamps2 <- sampling(lmmod, lminps2, cores = 3, chains = 3, iter = 2000)
check_hmc_diagnostics(lmsamps)

check_rhat(lmsamps2)
get_posterior_mean(lmsamps2)
stan_trace(lmsamps2, "A0")

stan_ess(lmsamps2)
pairs(lmsamps2, pars = c("A0[1]", "A0[2]", "A0[3]", "truesigma_err"))
```

```{r}
stan_trace(lmsamps, c("A0[1]", "A0[2]", "truesigma_err", "lp__"), inc_warmup = TRUE)
stan_trace(lmsamps2, c("A0[1]", "A0[2]", "truesigma_err", "lp__"), inc_warmup = TRUE)
# cache("lmsamps")
# cache("lmsamps2")
```


Is this working right? Check using sscase.

```{r}
ssinp <- within(sscase, {
  nt = ncol(W)
  nx = nrow(W)
  x = x[, 1]
  omega = t(as.data.frame(make_omegas(nx)))
  dA = t(dA)
  W = t(W)
  S = t(S)
  logA0_hat = log(500)
  sigma_logA0 = 2
  sigma_err = 0.1 * (max(dA) - min(dA))
}
)

optimizing(lmmod, ssinp)

```

Apparently! Can I verify log probability?

```{r}

```


### Bamr previous version

```{r}
dir.create("./templib")
devtools::install_github("markwh/bamr@be5e937", lib.loc = "./templib", local = FALSE, args = "--preclean")


```

