---
title: "notebook20180425"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Before I have the skype call with Mike and group, I want to try and see if I can reproduce things using Mike's simplest McFli. I had worked on this recently. When was it? Ah yes, Friday.

```{r}
h1fun <- function(w) {
  bs1 <- 1/5
  bw1 <- 100
  
  h1 <- (w - bw1) * bs1 / 2
  h1
}

a1fun <- function(w) {
  bw1 <- 100
  h1 <- h1fun(w)
  a1 <- bw1 * h1 + 1/2 * (w - bw1) * h1
  a1
}

h2fun <- function(w) {
  bs1 <- 1/20
  bw1 <- 50
  
  h1 <- (w - bw1) * bs1 / 2
  h1
}

a2fun <- function(w) {
  bw1 <- 50
  h1 <- h2fun(w)
  a1 <- bw1 * h1 + 1/2 * (w - bw1) * h1
  a1
}

h1fun(100)
h1fun(103.1)
a1fun(103.1)
a2fun(68)
```

I need to be able to go from Q to geometry.

```{r}


bs <- 20 * 1e-5
Q <- 5
ss <- 1/5
bw <- 100
n <- 0.04

respfun <- function(Q) {
  
  lhs <- Q * n * bs ^ (-1/2)
  obfun <- function(W) {
    rhs <- (ss / 4 * (W^2 - bw^2)) ^ (5/3) * W^(-2/3)
    out <- (lhs - rhs)^2
  }
  
  optw <- optimize(f = obfun, interval = c(bw, bw + 10 * bw / ss))
  W <- optw$minimum
  A <- (Q * n * bs^(-1/2) * W^(2/3))^(3/5)
  H <- (W - bw) * ss / 2
    
  out <- list(A = A, W = W, H = H)
}

1 / n * W ^ (-2/3) * A ^ (5/3) * bs ^ (1/2)
1 / 0.04 * 103.1 ^ (-2/3) * 31.3 ^ (5/3) * bs ^ (1/2)

```

Maybe it's down to exactitude of hydraulic radius?

Anyway, I can run with what I have. 

```{r}
makeRespFun <- function(n, bw, ss, bs) {
  scalfun <- function(Q) {
  
  lhs <- Q * n * bs ^ (-1/2)
  obfun <- function(W) {
    rhs <- (ss / 4 * (W^2 - bw^2)) ^ (5/3) * W^(-2/3)
    out <- (lhs - rhs)^2
  }
  
  optw <- optimize(f = obfun, interval = c(bw, bw + 10 * bw / ss + 10000 / ss))
  W <- optw$minimum
  A <- (Q * n * bs^(-1/2) * W^(2/3))^(3/5)
  H <- (W - bw) * ss / 2
    
  out <- data.frame(A = A, W = W, H = H)
  out
  }
  vecfun <- Vectorize(scalfun, SIMPLIFY = FALSE)
  out <- function(Q) {
    dplyr::bind_rows(vecfun(Q))
  } 
}


r1fun <- makeRespFun(0.04, 100, 1/5, 20 * 1e-5)
r2fun <- makeRespFun(0.04, 50, 1/20, 20 * 1e-5)

r1fun(5)
r2fun(5)
r1fun(20)
r2fun(20)
```

Cool. It doesn't vectorize though, huh?

- It does now!

```{r}
r1fun(c(5, 10))
```

Now I can do all kinds of crazy experiments!


```{r}
set.seed(26348)
bs <- 20 * 1e-5
n <- 0.04
bws <- runif(10, 100, 200)
sss <- 1 / runif(10, 5, 20)
rfuns <- map2(bws, sss, ~makeRespFun(n = n, bw = .x, ss = .y, bs = bs))
Qs <- rlnorm(11, log(25), 1)

resps <- map(rfuns, function(x) x(Qs)) %>% 
  map(~mutate(., Q = Qs, time = 1:nrow(.))) %>% 
  bind_rows(.id = "loc") %>% 
  mutate(loc = as.numeric(loc),
         S = bs) %>% 
  group_by(loc) %>% 
  mutate(dA = A - min(A)) %>% 
  glimpse()

resp_swot <- swot_untidy(resps)

gg1 <- swot_plot(resp_swot) +
  theme_bw()
ggsave("../graphs/poster/widthonly1.png", plot = gg1, width = 4, height = 3, dpi = 500)
```

Simulate error and observe A0 error. 

```{r}
add_dAerror <- function(swotlist, sd_frac) {
  dA0 <- swotlist$dA
  sd0 <- mean(apply(dA0, 1, sd))
  swotlist$dA <- dA0 + rnorm(nrow(dA0) * ncol(dA0), 0, sd_frac * sd0)
  swotlist
}

errfracs <- rep(1:10 / 200, each = 300)
simerrlists <- map(errfracs, ~add_dAerror(resp_swot, .)) %>% 
  map(~suppressMessages(estA0(.)))

realA0df <- as.data.frame(as.list(realA0(resp_swot))) %>% 
  setNames(paste0("x", 1:ncol(.))) %>% 
  gather(key = "loc", value = "realA0")
  

errsimdf <- simerrlists %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(errfracs, ~mutate(.x, err_frac = .y)) %>% 
  bind_rows() %>% 
  gather(key = "loc", value = "A0est", -err_frac) %>% 
  left_join(realA0df, by = "loc") %>% 
  mutate(relerr = (A0est - realA0) / realA0 * 100) 

gg6 <- ggplot(errsimdf, aes(x = err_frac * 100, y = relerr)) +
  geom_point(position = position_jitter(), size = 0.2, alpha = 0.2) + 
  theme_bw() +
  ylab("Relative A0 error, percent") +
  xlab("Relative dA error, percent") +
  scale_x_continuous(breaks = 1:10)

ggsave("../graphs/poster/widthonly2.png", width = 4, height = 3, dpi = 500)

```


```{r}
trueA0 <- resp_swot

foolm <- estA0_lm(swotlist = resp_swot)

foolm %>% 
  summary()

xmat <- as.matrix(model.frame(foolm)[-1])

invmat <- solve(t(xmat) %*% xmat)

diag(invmat)
```

Just a little aside on how R is handling the matrix inversions. From looking at `summary.lm`.


```{r}
dput(stats:::qr.lm)
Qr <- foolm$qr

Qr$qr
rnk <- foolm$rank

R <- chol2inv(Qr$qr[1:rnk, 1:rnk, drop= FALSE])
diag(R)
```

This leaves the Qr decomposition as a black box (which extends down to fortran level), as well as the inversion from this decomp.

```{r}
pairs(t(manning_ws35(resp_swot)))
```

Can I do the condition-number pruning?
  

First, try an optimization. 

- Fix Q (several of them)
- Fix m1, b1, 




