---
title: "notebook20180315"
author: "Mark Hagemann"
date: "March 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today:

- Chase this idea of using matrix properties to determine if we can infer A0. 
- Design and build out interactive checks on:
    - Manning
    - Mass conservation
    - other A0 inference assumptions (more direct--in 3/5 space)
    
    
As per handwritten work, condition number seems like a reasonable thing to check for A0 inference. 

```{r}
condNo <- function(mat) {
  eigvals <- eigen(mat)[["values"]]
  out <- abs(eigvals[1]) / abs(eigvals[length(eigvals)])
  out
}

lmCondNo <- function(mod) {
  cormat <- cor(mod$model[, -1])
  out <- condNo(cormat)
  out
}

map_dbl(A0lms, lmCondNo) %>% 
  sort()
```


Found a paper (Tamura et al., 2016) discussing variable subset methods for reducing collinearity. 

It seems there may be a tradeoff between eliminating (Manning) model and measurement error and eliminating collinearity. The former would require more locations, the latter would require fewer. 

OK, reading intro to paper, the approach is pretty straightforward. 

1. Calculate condition number of full model
2. For i in 1:(number of locations):
    - remove location i
    - Calculate condition number of model without location i
    - Remember this value
3. New model is model from part 2 with smallest condition number
4. Return to step 2 using new model.
5. Repeat the above until some threshold is reached--either condition number floor or location number floor. 

Easy. Test this out on one of the A0lms. 

```{r}
testcase <- reachdata$Kanawha

# modmat <- testlm$model[, -1]
modmat <- with(testcase, na.omit(t((W^(-2/3) * S^(1/2))^(3/5))))
nlocs <- ncol(modmat)

elims <- numeric(0)
condnos <- condNo(cor(modmat))
for (i in 1:nlocs) {
  nlocs_remain <- nlocs - i + 1
  if (nlocs_remain <= 2) break
  condnos_i <- map_dbl(1:nlocs_remain, ~condNo(cor(modmat[, -.])))
  elim_i <- which.min(condnos_i)
  
  modmat <- modmat[, -elim_i]
  
  elims <- c(elims, elim_i)
  condnos <- c(condnos, condNo(cor(modmat)))

}
```

I need to think about whether this is affected by weight matrix / vector. 

- Yes, of course it is. Really I should investigate correlation structure of WS matrix in 3/5 space. 
- OK, modified the above accordingly. 

```{r}
ws35 <- map(reachdata, function(x) t(x$W^(-2/3) * x$S^(1/2))^(3/5))

pairs(ws35$Connecticut)
pairs(ws35$Severn[-1, ])

which.max(ws35$Severn[, 4])
```

Spin-up issue in Severn. NEED TO FIX THIS. I should build a shiny app for visualizing Pepsi cases. 

But first, finish characterizing condition numbers.

```{r}
recondition <- function(modmat) {
  nlocs <- ncol(modmat)
  
  elims <- numeric(0)
  remains <- 1:nlocs
  condnos <- condNo(cor(modmat))
  for (i in 1:nlocs) {
    nlocs_remain <- nlocs - i + 1
    if (nlocs_remain <= 2) break
    condnos_i <- map_dbl(1:nlocs_remain, ~condNo(cor(modmat[, -.])))
    elim_i <- remains[which.min(condnos_i)]
    
    remains <- remains[-which.min(condnos_i)]
    modmat <- modmat[, -which.min(condnos_i)]
    
    elims <- c(elims, elim_i)
    condnos <- c(condnos, condNo(cor(modmat)))
  
  }
  
  out <- data.frame(elim = c(NA_integer_, elims),
                    n_elim = 0:length(elims),
                    condno = condnos)
  out
}

recres <- recondition(ws35$Po)
plot(recres$condno)

pairs(ws35$Po[, -recres$elim[2:13]])

condNo(cor(ws35$Po[, -recres$elim[2:8]]))
condNo(cor(ws35$Po[, -recres$elim[2:3]]))
```

That seems to work. Ready to put it into action?

- That would require setting a threshold either on condition number or number of locations. 
- Start with individual cases, showing A0 inversion as funciton of number of locations remaining when pruning this way. Alternatively, as function of condition number. 

```{r}
testcase <- "GaronneUpstream"
rmtimes <- apply(ws35[[testcase]], 2, function(x) which(is.na(x))) %>% 
  unlist() %>% 
  unique()
if (length(rmtimes) == 0)
  rmtimes <- -(1:nrow(ws35[[testcase]]))

recres <- recondition(ws35[[testcase]][-rmtimes, ])
rmlocs <- c(list(NA), Reduce(c, recres$elims, accumulate = TRUE))
# keeplocs <- map(rmlocs, function(x) (1:ncol(ws35[[testcase]]))[-x])

estA0_sset <- function(rmlocs = NA, rmtimes = NA, lst) {
  # browser()
  force(rmlocs)
  force(rmtimes)
  
  if (length(rmtimes) == 1L && is.na(rmtimes))
    rmtimes <- -(1:ncol(lst$W))
  
  if (length(rmlocs) == 1L && is.na(rmlocs))
    rmlocs <- -(1:nrow(lst$W))
  
  # print(map_int(lst, ~nrow(.[-rmlocs, -rmtimes])))
  estA0(wmat = lst$W[-rmlocs, -rmtimes], 
        smat = lst$S[-rmlocs, -rmtimes], 
        damat = lst$dA[-rmlocs, -rmtimes])
}
  

A0list <- rmlocs %>% 
  map(~estA0_sset(rmlocs = ., rmtimes = rmtimes, 
                  lst = reachdata[[testcase]])) %>% 
  map(coef)

putInCoefs <- function(coefs, rmlocs) {
  if (length(rmlocs) == 1L && is.na(rmlocs)) {
    rmlocs <- -(1:length(coefs))
  }
  starter <- rep(NA_real_, ncol(ws35[[testcase]]))
  starter[-rmlocs] <- coefs
  starter
}

realA0 <- with(reachdata[[testcase]], A - dA)[, 1] %>% 
  data.frame(location = 1:length(.), A0_real = .)
condnodf <- data.frame(nremoved = (1:length(recres$condno)) - 1, 
                       condno = recres$condno)

resdf <- map2(A0list, rmlocs, putInCoefs) %>% 
  as.data.frame() %>% 
  setNames((1:length(A0list)) - 1) %>% 
  mutate(location = 1:nrow(.)) %>% 
  gather(key = "nremoved", value = "A0_est", -location) %>% 
  left_join(realA0, by = "location") %>% 
  mutate(A0_err = (A0_est - A0_real) / A0_real, 
         nremoved = as.numeric(nremoved)) %>% 
  left_join(condnodf, by = "nremoved")

ggplot(resdf, aes(x = nremoved, y = A0_err)) +
  geom_point()

ggplot(resdf, aes(x = condno, y = A0_err)) +
  geom_point()

```


Make a function. 

```{r}

obslist <- reachdata$Seine

estA0_prune_condno <- function(obslist) {
  
  wmat <- obslist$W
  smat <- obslist$S
  amat <- obslist$A
  dAmat <- obslist$dA
  
  ws35 <- t(wmat ^ (-2/3) * smat ^ (1/2)) ^ (3/5)
  recres <- recondition(ws35)
  
  rmlocs <- Reduce(c, recres$elim[-1], accumulate = TRUE)
  keeplocs <- c(list("all"), map(rmlocs, `-`))
  
  A0lms <- map(keeplocs, ~estA0_list(obslist = obslist, keeplocs = .))
  lmpreds <- map(A0lms, predict)
  lmyobs <- map(A0lms, ~.$model[, 1])
  
  A0r2 <- map2_dbl(lmyobs, lmpreds,
               function(x, y) 1 - sum((x - y)^2)/sum((x - mean(x))^2)) %>%
    data.frame(n_elim = (1:length(.) - 1), R2 = .)
  # A0r2 <- map_dbl(A0lms, markstats::Q2) %>% 
  #   data.frame(n_elim = (1:length(.) - 1), R2 = .)
  
  
  allocs <- 1:nrow(wmat)
  coeflocs <- c(list(allocs), map(rmlocs, ~allocs[-.]))
  A0ests <- map2(A0lms, coeflocs, function(x, y) setNames(coef(x), y)) %>% 
    map(~data.frame(loc = names(.), A0_est = .)) %>% 
    setNames(0:(length(.) - 1)) %>% 
    bind_rows(.id = "n_elim") %>% 
    mutate(n_elim = as.integer(n_elim), 
           loc = as.integer(loc))
  
  out <- left_join(A0ests, recres, by = "n_elim") %>% 
    left_join(A0r2, by = "n_elim") %>% 
    select(loc, condno, R2, elim, n_elim, A0_est)
  
  if (!is.null(amat)) {
    A0reals <- data.frame(loc = 1:nrow(wmat), A0_real = (amat - dAmat)[, 1])
    out <- out %>% 
      left_join(A0reals, by = "loc")
  }
  
  out
}

estA0_prune_condno(reachdata$Ganges)
```

Now I can do the whole lot!

```{r}
omitnas <- function(obslist) {
  obsnas <- map(obslist, ~which(is.na(.), arr.ind = TRUE)) %>% 
    map(as.data.frame) %>% 
    bind_rows()
  rmthese <- unique(obsnas$col)
  if(length(rmthese) == 0L)
    return(obslist)
  out <- map(obslist, ~.[, -rmthese])
  out
}

foo <- reachdata %>% 
  map(omitnas) %>% 
  map(possibly(estA0_prune_condno, otherwise = NA))

foodf <- foo %>% 
  bind_rows(.id = "case") %>% 
  mutate(A0_relerr = abs((A0_est - A0_real) / A0_real)) %>% 
  group_by(case, n_elim) %>% 
  mutate(medianerr = median(A0_relerr))

ggplot(foodf, aes(x = R2, y = A0_relerr, color = case)) +
  geom_point() +
  geom_line(aes(y = medianerr, group = case))

ggplot(foodf, aes(x = condno, y = A0_relerr, color = case)) +
  geom_point() +
  geom_line(aes(y = medianerr, group = case)) +
  scale_x_log10()


foolm <- lm(A0_relerr ~ R2 + log(condno), data = foodf)

```


Now a prune function for R2. 

```{r}
obslist <- reachdata$Connecticut
estA0_prune_R2 <- function(obslist) {
  
  wmat <- obslist$W
  smat <- obslist$S
  amat <- obslist$A
  dAmat <- obslist$dA

  fullm <- estA0_list(obslist = obslist)
    
  nlocs <- nrow(obslist$W)
  
  myR2 <- function(model) {
    markstats::R2(model$model[, 1], predict(model))
  }
  
  locs_remain <- 1:nlocs
  r2s <- myR2(fullm)
  elims <- NA
  coefs <- list(setNames(coef(fullm), 1:nlocs))
  for (i in 1:nlocs) {
    
    print(sprintf("iteration %s of %s ", i, nlocs - 1))
    
    nlocs_remain <- nlocs - i + 1
    if (nlocs_remain <= 2) 
        break
    
    A0lms_i <- map(1:length(locs_remain), ~estA0_list(obslist, keeplocs = locs_remain[-.]))
    
    r2_i <- map_dbl(A0lms_i, myR2)
    elimind <- which.max(r2_i)
    elim_i <- locs_remain[elimind]
    locs_remain <- locs_remain[-elimind]
    coefs_i <- setNames(coef(A0lms_i[[elimind]]), locs_remain)
    
    elims <- c(elims, elim_i)
    r2s <- c(r2s, max(r2_i))
    coefs <- c(coefs, list(coefs_i))
    
  }
  
  coefdf <- coefs %>% 
    map(~data.frame(loc = names(.), A0_est = .)) %>% 
    setNames((1:length(coefs)) - 1) %>% 
    bind_rows(.id = "nremoved") %>% 
    mutate(nremoved = as.integer(nremoved), 
           loc = as.integer(loc))
  r2df <- data.frame(nremoved = (1:length(coefs)) - 1,
                     R2 = r2s, elim = elims)
  
  Arealdf <- data.frame(loc = 1:nlocs, A0_real = (amat - dAmat)[, 1])
  
  out <- left_join(coefdf, r2df, by = "nremoved") %>% 
    left_join(Arealdf, by = "loc")
  
  out
}


out <- estA0_prune_R2(reachdata$Ganges)
out %>% 
  mutate(relerr = (A0_est - A0_real) / A0_real) %>% 
  ggplot(aes(x = R2, y = relerr)) +
  geom_point()


```

Now do them all!

```{r}

bar <- reachdata %>% 
  map(omitnas) %>% 
  map(possibly(estA0_prune_R2, otherwise = NA))

bardf <- bar[!is.na(bar)] %>% 
  bind_rows(.id = "case") %>% 
  mutate(A0_relerr = (A0_est - A0_real) / A0_real) %>% 
  group_by(case, nremoved) %>% 
  mutate(medianerr = median(A0_relerr))

ggplot(bardf, aes(x = R2, y = abs(A0_relerr), color = case)) +
  geom_point() +
  geom_line(aes(y = abs(medianerr), group = case))
  scale_x_log10()

bardf %>% 
  group_by(case, nremoved) %>% 
  mutate(n = n()) %>% 
  ggplot(aes(x = n, y = A0_relerr, color = case)) + 
  geom_point() +
  geom_line(aes(y= medianerr, group = case))
```

