---
title: "notebook20180830"
author: "Mark Hagemann"
date: "August 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Working on getting stan to play nice with ragged swot data. Herein I shall test the changes. 

```{r}
library(rstan)
library(bamr)
library(swotr)
library(dplyr)
load("../cache/reachdata.RData")

mod1 <- stan_model("../../bamr/exec/master.stan")

testdat1 <- swot_bamdata(reachdata$Ganges)
testinps1 <- bamr:::compose_bam_inputs(testdat1) %>% 
  within(., {
    ntot_man <- ntot_amhg <- nx * nt
    hasdat_man <- hasdat_amhg <- matrix(1, nr = nx, nc = nt)
    inc_m <- 0
    inc_a <- 1
    meas_err <- 0
  })

opts1 <- optimizing(mod1, data = testinps1, as_vector = FALSE)
samps1 <- sampling(mod1, data = testinps1, chains = 1, 
                   pars = c("Wact", "Sact", "dApos_act", "man_lhs", "man_rhs",
                            "amhg_rhs", "logA_man", "logQ_man", "logQ_amhg"),
                   include = FALSE)

samps2 <- sampling(bamr:::stanmodels$master, data = testinps1, chains = 1, 
                   pars = c("Wact", "Sact", "dApos_act", "man_lhs", "man_rhs",
                            "amhg_rhs", "logA_man", "logQ_man", "logQ_amhg"),
                   include = FALSE)

```

From a first-pass, ~~the new version takes a little over 2x as long as the old.~~ Nope, it's the same! 

How do the results compare?

```{r}
bam_hydrograph(samps1)
bam_hydrograph(samps2)

stan_trace(samps1, pars = "A0", inc_warmup = TRUE)
stan_trace(samps1, pars = "logQ[1]", inc_warmup = TRUE)

stan_trace(samps2, pars = "A0", inc_warmup = TRUE)
stan_trace(samps2, pars = "logQ[1]", inc_warmup = TRUE)

```

Test out the ragged_vec and rep functions

```{stan output.var = "raggedtest"}

functions {
  // Convert an array to a vector based on a binary matrix 
  // indicating non-missing data
  
 vector ragged_vec(vector[] x, int[,] bin) {
    vector[num_elements(x)] out;
    int ind;
    
    ind = 1;
    for (i in 1:size(x)) {
      for (t in 1:num_elements(x[1])) {
        // print(bin[i, t]);
        // print(ind);
        if (bin[i, t] == 1) {
          out[ind] = x[i, t];
          ind += 1;
        }
      }
    }
    // print(out);
    return(out[1:(ind - 1)]);
  }
}

data {
  int nx;
  int nt;
  int ntot;
  int isobs[nx, nt];
  vector[nt] xmat[nx];
}

transformed data {
  vector[ntot] xvec;
  
  xvec = ragged_vec(xmat, isobs);
  print(xvec);
}

parameters {
  real mu; 
}

model {
  xvec ~ normal(mu, 1);
  mu ~ normal(0, 1);
}

```

```{r}

nx <- 3
nt <- 4
ntot <- 10
isobs <- matrix(sample(c(rep(1, ntot), rep(0, nx * nt - ntot))),
                nr = nx, nc = nt)
xmat <- matrix(rnorm(nx * nt), nc = nt, nr = nx)

raginps <- list(
  nx = nx, 
  nt = nt, 
  ntot = ntot,
  xmat = xmat,
  isobs = isobs
)

ragopt <- optimizing(raggedtest, data = raginps)
ragsamp <- sampling(raggedtest, data = raginps, chains = 1)
```



```{stan output.var = "reptest"}

functions {
  // repeat each element of a vector a given number of times
  vector rep(vector x, int[] each) {
    vector[sum(each)] out;
    int startind;
    int endind;
    
    endind = 0;
    startind = 1;
    for (i in 1:size(each)) {
      endind = endind + each[i];
      out[startind:endind] = rep_vector(x[i], each[i]);
      startind = endind + 1;
    }
    
    return(out);
  }
  
  // Number of reps
  int[] nreps(int[,] bin, int margin) {
    int nt;
    int nx;
    int nn;
    int out[max(size(bin), num_elements(bin[1]))];
    
    nx = size(bin);
    nt = num_elements(bin[1]);
    nn = margin == 1 ? nx : nt;

    // initialize vector to zero
    for (ind in 1:nn) {
      out[ind] = 0;
    }

    for (i in 1:nx) {
      for (t in 1:nt) {
        if (margin == 1)
          out[i] = out[i] + bin[i, t];
        else
          out[t] = out[t] + bin[i, t];
      }
    }
    return(out[1:nn]);
  }
}

data {
  int nx;
  int nt;
  int ntot;
  int isobs[nx, nt];
  vector[nx] xvec1;
  vector[nt] xvec2;
}

transformed data {
  int times1[nx];
  int times2[nt];
  vector[ntot] longvec1;
  vector[ntot] longvec2;
  
  times1 = nreps(isobs, 1);
  times2 = nreps(isobs, 2);
  
  longvec1 = rep(xvec1, times1);
  longvec2 = rep(xvec2, times2);
  
  print(times1);
  print(times2);
  print(longvec1);
  print(longvec2);
  print(num_elements(longvec1));
  print(num_elements(longvec2));
}

parameters {
  real mu;
}

model {
  mu ~ normal(0, 1);
}
```


```{r}
nx <- 3
nt <- 4
ntot <- 10
isobs <- matrix(sample(c(rep(1, ntot), rep(0, nx * nt - ntot))),
                nr = nx, nc = nt)
xmat <- matrix(rnorm(nx * nt), nc = nt, nr = nx)

repinps <- list(
  nx = nx, 
  nt = nt, 
  ntot = ntot,
  xvec1 <- rnorm(nx),
  xvec2 <- rnorm(nt),
  isobs = isobs
)

repopt <- optimizing(reptest, data = repinps)

```

That appears to work! Now go back and retest the new implementation. 

Making progress. I scrapped the above functions when I realized I had to be more deliberate about vectorizing arrays and repeating vectors--rows first! then columns. Anyway, I now need to test a new function that is needed for defining Wact_man using Wact and the integer arrays giving non-missing Manning and AMHG indices. 

```{stan output.var = "indtest"}

functions {
  vector ragged_vec(vector[] x, int[,] bin) {
    vector[num_elements(x)] out;
    int ind;
    
    ind = 1;
    for (i in 1:size(x)) {
      for (t in 1:num_elements(x[1])) {
        // print(bin[i, t]);
        // print(ind);
        if (bin[i, t] == 1) {
          out[ind] = x[i, t];
          ind += 1;
        }
      }
    }
    // print(out);
    return(out[1:(ind - 1)]);
  }

  // indices of vectorized bin1 that are also in vectorized bin2
  int[] commoninds(int[,] bin1, int[,] bin2) {
    int out[num_elements(bin1)];
    int vecinds[size(bin1), num_elements(bin1[1])];
    int ctr;
    int ind;

    ctr = 0;
    for (i in 1:size(bin1)) {
      for (t in 1:num_elements(bin1[1])) {
        if (bin1[i, t] == 1) {
          ctr += 1;
          vecinds[i, t] = ctr;
        }
      }
    }

    ind = 0;
    for (i in 1:size(vecinds)) {
      for (t in 1:size(vecinds[1])) {
        if (bin2[i, t] == 1) {
          ind += 1;
          out[ind] = vecinds[i, t];
        }
      }
    }
    return(out[1:ind]);
  }
}

data {
  int nx;
  int nt;
  //  int ntot;
  int isobs1[nx, nt];
  int isobs2[nx, nt];
  int nobs1;
  int nobs2;
  vector[nt] xmat[nx];
}

transformed data {
  vector[nobs1] xvec1;
  int inds12[nobs2];
  
  xvec1 = ragged_vec(xmat, isobs1);
  inds12 = commoninds(isobs1, isobs2);
  
  print(xvec1);
  print(inds12);
  print(xvec1[inds12]);
}

parameters {
  real mu;
}

model {
  mu ~ normal(0, 1);
}

```



```{r}
  int nx;
  int nt;
  int isobs1[nx, nt];
  int isobs2[nx, nt];
  int nobs1;
  int nobs2;
  vector[nt] xmat[nx];

nx <- 3
nt <- 4
isobs1 <- matrix(c(1, 1, 0, 0, 
                   1, 0, 1, 1, 
                   1, 1, 0, 1), byrow = TRUE, nrow = 3)
isobs2 <- isobs1
isobs2[2, 1] <- isobs2[2, 4] <- isobs2[3, 2] <- 0
xmat <- matrix(rnorm(12), nr = 3, nc = 4)

indinps <- list(
  nx = nx, 
  nt = nt, 
  isobs1 = isobs1, 
  isobs2 = isobs2, 
  nobs1 = sum(isobs1),
  nobs2 = sum(isobs2),
  xmat = xmat
)

indopts <- optimizing(indtest, data = indinps)
```

OK!

