

Need to check that my AR(1) likelihood funciton and gradient are OK. 

```{r}

llik1 <- logPost_ar1(datadf = data.frame(obs, obsdates))

# library(stats4)
# library(bbmle)
# test1 <- stats4::mle(llik1, start = list(params = c(7, 2)))
# 
# test2 <- mle2(llik1, start = list(params = c(7, 2)))

test3 <- nlm(llik1, p = c(7, 2))
```

How to test if this works? Use synthetic data? Plot the data?

```{r}
test3
hist(log(obs + exp(6.0605)))
```

I wonder if it's that one anomalous dA that's throwing this off. 

```{r}
which.min(obs)

obsdf_omit <- data.frame(obs[-394], obsdates[-394])
llik2 <- logPost_ar1(datadf = obsdf_omit)

test4 <- nlm(llik2, p = c(7, 2))
test4


hist(log(obs[-394] + exp(6.588)))

```

```{r}
xv1 <- seq(from = log(-min(obs - 1)), to = 10, length.out = 100)
yv1 <- 10^seq(from = -1.5, to = 1, length.out = 100)
gridar2 <- llikGrid(xvec = xv1, yvec = yv1, nllfun = llik2)


priorvals <- list(priors$logA0_hat, 
                  priors$sigmalogA_hat, 
                  priors$sigmalogA_sd)

llik3 <- logPost_ar1(datadf = obsdf_omit, 
                     priors = priorvals)
gridar3 <- llikGrid(xvec = xv1, yvec = yv1, nllfun = llik3)

llikContour(gridar2, realx = log(real_A0), realy = real_sigmalogA)
llikContour(gridar3, realx = log(real_A0), realy = real_sigmalogA)

```

That's... reasonable.

```{r}
foo <- mledf(logPost_ar1, 
             datadf = data.frame(obs, obsdates), 
             priors = priorvals,
             p = c(7, 0.2), cumulative = TRUE)
head(foo)

plot(exp(foo$V1), ylim = c(0, 1000))
abline(h = real_A0)
plot(foo$V2, ylim = c(0, 0.65))
abline(h = real_sigmalogA)
```

Compare against no-AR version

```{r}
mle_noar <- mledf(logPost, 
             datadf = data.frame(obs, obsdates), 
             priors = priorvals,
             p = c(7, 0.2), cumulative = TRUE)

head(mle_noar)


plot(exp(mle_noar$V1), ylim = c(0, 1000))
abline(h = real_A0)
plot(mle_noar$V2, ylim = c(0, 0.65))
abline(h = real_sigmalogA)
```

Because of the impact of this outlier, I think I need to go back through HydroSWOT and check for outliers in logQ ~ logA. Use studentized deleted residuals. 

```{r}
qalm_train <- train_full %>% 
  mutate(logA = log(area_m2), logQ = ) %>% 
  split(.$xs) %>% 
  map(~lm(logQ ~ logA, data = .))
```

Potential issue that this does not check logA if logQ is missing or Inf. Can I use logA alone?

```{r}
hist(testdata$area_m2 %>% log() %>% scale() %>% `[`(-394))
```

Sure, but it's more difficult. Just do version using Q for now. Then redo log posteriors for test set. If there are still problem cases I can revisit.

```{r}
qalm_train <- train_full %>% 
  mutate(logA = log(area_m2), logQ = ifelse(is.infinite(logQ), NA, logQ)) %>% 
  split(.$xs) %>% 
  map(~lm(logQ ~ logA, data = .))

delres_train <- map(qalm_train, ~delResid(.) / sqrt(delMSE(.))) %>% 
  map(~data.frame(sdr = .)) %>% 
  bind_rows(.id = "xs") %>% 
  group_by(xs) %>% 
  mutate(obsno = 1:n()) %>% 
  ungroup()

train_full_sdr <- train_full %>% 
  group_by(xs) %>% 
  mutate(obsno = 1:n()) %>% 
  ungroup() %>% 
  mutate(xs = as.character(xs)) %>% 
  left_join(delres_train, by = c("xs", "obsno"))



```

See some plots based on this.

```{r}
minsdrs <- train_full_sdr %>% 
  group_by(xs) %>% 
  summarize(minsdr = min(sdr), maxsdr = max(sdr)) %>% 
  arrange(minsdr)

maxsdrs <- train_full_sdr %>% 
  group_by(xs) %>% 
  summarize(minsdr = min(sdr), maxsdr = max(sdr)) %>% 
  arrange(desc(maxsdr))

i <- 1

# train_full_sdr %>% 
#   # filter(xs == "1053600") %>% 
#   filter(xs == minsdrs$xs[i]) %>% #`[[`("area_m2")
#   plot(logQ ~ log(area_m2), ., main = minsdrs$minsdr[i])
# i <- i + 1

train_full_sdr %>% 
  # filter(xs == "1053600") %>% 
  filter(xs == maxsdrs$xs[i]) %>% #`[[`("area_m2")
  plot(logQ ~ log(area_m2), ., main = maxsdrs$maxsdr[i])
i <- i + 1
```

6 seems like a reasonable threshold. Make a function that can be caled multiple times on hswot2 to remove these points. 

```{r}

refitRemove <- function(df, exclude = rep(FALSE, nrow(df)), thresh = 8) {
  
  
  lm1 <- lm(logQ ~ logA, df[!exclude, ])
  
  nas <- is.na(df$logQ | is.na(df$logA))
  
  dsrs <- delResid(lm1) / sqrt(delMSE(lm1))
  dsrbad <- abs(dsrs) > thresh
  dsrbad[is.na(dsrbad)] <- FALSE
  
  exclude[!exclude & !nas] <- dsrbad
  
  out <- df
  out$exclude <- exclude
  
  if (sum(dsrbad) > 0)
    out <- refitRemove(out, exclude = exclude, thresh = thresh)
  out
}

cleanse_A_outliers <- function(hswot2, thresh = 8) {
  lms <- hswot2 %>% 
    mutate(logA = log(area_m2), logQ = ifelse(is.infinite(logQ), NA, logQ)) %>% 
    split(.$xs) %>% 
    map(refitRemove, thresh = thresh) %>% 
    bind_rows()

  # hswot2_sdr <- hswot2 %>% 
  #   group_by(xs) %>% 
  #   mutate(obsno = 1:n()) %>% 
  #   ungroup() %>% 
  #   mutate(xs = as.character(xs)) %>% 
  #   left_join(delres, by = c("xs", "obsno")) %>% 
  #   mutate(dumpit = lt(thresh, abs(sdr)))
  
  print(sprintf("Removed %s area outliers based on logQ ~ logA sdr", 
                sum(lms$exclude)))
  
  out <- lms %>% 
    filter(!exclude) %>% 
    select(-exclude, -logA) %>% 
    group_by(xs) %>% 
    mutate(n = n()) %>% 
    ungroup() %>% 
    filter(n > 30)
  out
}

hswot3.1 <- hswot2

# nr_old <- nrow(hswot3.1)
# nr_new <- 0
# while(nr_old != nr_new) {
#   nr_old <- nrow(hswot3.1)
hswot3.1 <- cleanse_A_outliers(hswot3.1)
#   nr_new <- nrow(hswot3.1)
# }


foo <- filter(hswot2, xs %in% sample(xs, 10))
bar <- cleanse_A_outliers(foo)
# bar <- refitRemove(foo %>% mutate(logA = log(area_m2), logQ = ifelse(is.infinite(logQ), NA, logQ)))

cache("hswot3")
```

This is done. Next I can redo the maximum likelihood estimates of A0. 

```{r}

noprior_train_ar1 <- train_full %>% 
  split(.$xs) %>% 
  map(~logPost_ar1(datadf = data.frame(.$dA, .$Date)))

starts <- train_full %>% 
  group_by(xs) %>% 
  summarize(logA0_start = log(-min(dA)) + 1) %>% 
  split(.$xs) %>% 
  map(~c(.$logA0_start, 0.2))

starts[[1]]  

mynlm <- function(x, y) nlm(f = x, p = y)

mle1_noprior <- map2(noprior_train_ar1, starts, nlm) %>%
  map(~.$estimate) %>%
  map(~as.data.frame(as.list(.))) %>%
  map(setNames, c("logA0", "sigma_logA")) %>%
  bind_rows(.id = "xs") %>% 
  mutate(xs = as.numeric(xs))

mle1_noprior

train_val <- train_smry %>% 
  transmute(xs, logA0_real = log(A0))
train_val %>% 
  left_join(mle1_noprior, by = "xs") %>% 
  glimpse() %>% 
  plot(logA0 ~ logA0_real, .)
  


```

Neat, now try one with priors. 

```{r}

priordf <- stanPriorList %>% 
  map(as.data.frame) %>% 
  bind_rows(.id = "xs") %>% 
  transmute(xs = as.numeric(xs), logA0_hat, sigmalogA_hat, sigmalogA_sd)

train_ar2 <- train_full %>% 
  left_join(priordf, by = "xs") %>% 
  glimpse() %>% 
  split(.$xs) %>% 
  map(~logPost_ar1(datadf = data.frame(.$dA, .$Date), 
                   priors = list(.$logA0_hat[1], 
                                 .$sigmalogA_hat[1], 
                                 .$sigmalogA_sd[1])))

starts <- train_full %>% 
  group_by(xs) %>% 
  summarize(logA0_start = log(-min(dA)) + 1) %>% 
  split(.$xs) %>% 
  map(~c(.$logA0_start, 0.2))


mynlm <- function(x, y) nlm(f = x, p = y)

mle1_withprior <- map2(train_ar2, starts, nlm) %>%
  map(~.$estimate) %>%
  map(~as.data.frame(as.list(.))) %>%
  map(setNames, c("logA0", "sigma_logA")) %>%
  bind_rows(.id = "xs") %>% 
  mutate(xs = as.numeric(xs))

mle1_noprior

train_val <- train_smry %>% 
  transmute(xs, logA0_real = log(A0))
train_val %>% 
  left_join(mle1_withprior, by = "xs") %>% 
  glimpse() %>% 
  plot(logA0 ~ logA0_real, .)
  

```


