---
title: "notebook20180517"
author: "Mark Hagemann"
date: "May 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Working on A' variance approximation. 

I need a way to approximate var(p + epsilon_p) when cov is large. When cov is small, I have the following:

$$
var((p + \epsilon_p)^{1/2}) = p(\exp(\frac{\sigma^2_p}{2p^2}) - \exp(\frac{\sigma^2_p}{4 p^2})) 
$$


Which I can show empirically:

```{r}
sigmas <- 1:10 / 10

simdats <- map(sigmas, ~rnorm(1000, 1, .))

simvars <- map_dbl(simdats, ~var(sqrt(.[.>0])))

# approxvars <- (1 - exp(sigmas^2 / 4))
approxvars <- exp(sigmas^2 / 2) - exp(sigmas^2 / 4)

plot(simvars)
plot(simvars, approxvars); abline(0, 1)
```

Now the high cov version is :

$$
var(p + \epsilon_p) = \sigma^2_p / 2 - 0.2655 \sigma^3_p
$$


```{r}
sigmas <- 1:3 0 / 5

simdats <- map(sigmas, ~rnorm(1000, 1, .))

simvars <- map_dbl(simdats, ~var(sqrt(.[.>0])))

newsigs <- sqrt(sigmas^2 + 1)
approxvars <- (newsigs * 0.1219)

plot(simvars)
plot(simvars, approxvars); abline(0, 1)2
```



Great, now I've got that worked out. Update equation 1. Done. Now make into functions. 


```{r}
approxvar1 <- function(p, sigma_p) {
  
  expbit <- exp(sigma_p^2 / (4 * p^2))
  out <- (expbit - 1) * expbit * p
  out
}

approxvar2 <- function(p, sigma_p) {
  
  newsigma <- sqrt(sigma_p^2 + p^2)
  newsigma * 0.1219
}


p <- 5
cvs <- 10^seq(-2, 1.5, length.out = 21)
sigmas <- cvs * p
simvars <- map(sigmas, ~abs(rnorm(10000, p, .))) %>% 
  map_dbl(~var(sqrt(.)))

# simvars <- simcvs * p^2

valdf <- data.frame(p = p, sigma_p = sigmas, cv = cvs, 
                    sim_var = simvars) %>% 
  mutate(approx1 = approxvar1(p, sigma_p),
         approx2 = approxvar2(p, sigma_p))

ggplot(valdf, aes(x = cv, y = sim_var)) +
  geom_point(aes(shape = "empirical")) +
  geom_line(aes(y = approx1, color = "approx1")) +
  geom_line(aes(y = approx2, color = "approx2")) +
  # ylim(0, max(valdf$sim_var * 1.5)) +
  scale_x_log10(breaks = 10^(-2:1),
                labels = 10^(-2:1)) +
  scale_y_log10(limits = c(min(valdf$sim_var) / 2,
                           max(valdf$sim_var * 1.5))) +
  annotation_logticks() +
  xlab("Coefficient of Variation") +
  ylab("Empirical Variance") +
  theme_bw() +
  theme(legend.title = element_blank())
  
```

Great! Now repeat for (p + e_p)^(-1/2)

OK, all is done. See reports/Aprime_variance.Rmd. Note that math above is not correct.
