---
title: "notebook20180404"
author: "Mark Hagemann"
date: "April 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I need to discern exactly how error in McMan assumptions get translated into model errors:

- Errors in model matrix
- Errors in response

Start with Manning

$$
Q_{it} = \frac{1}{n_{it}} A_{it}^{5/3}W_{it}^{-2/3}S_{it}^{1/2}
$$

Alternatively:

$$
Q_{it}^{3/5} = n_{it}^{-3/5} A (W_{it}^{-2/3}S_{it}^{1/2})^{3/5}
$$


I usually to take flow and n to be exactly constant across space, and n to be constant in time. This is not strictly speaking true. Compare:

$$
X_{it} = (W_{it}^{-2/3}S_{it}^{1/2})^{3/5}
$$

In theory, a mass-conserved observation matrix. (That is, $\mathbf{X} \omega = \mathbf{0}$ for any vector $\omega$ such that $\mathbf{1}'\omega = 0$)


But really we need to figure in closure term. In order to determine (additive) error, this must be additive in linear-A space, so it will be somewhat contrived. 

Per (handwritten) notebook work:

$$
\mathbf{X}_{real} = \mathbf{X}_{meas} + \mathbf{\Gamma}
$$

where the elements of $\mathbf{Gamma}$ are $\gamma_{it} = (Q_{it}n_{it})^{3/5} -  A (W_{it}^{-2/3}S_{it}^{1/2})^{3/5}$

Now calculate this for sscase. 

```{r}
ssnmat <- with(sscase, manningN(A = A, W = W, S = S, Q = Q))
ssgamma <- (sscase$Q * ssnmat)^(3/5) - manning_ws35(sscase) * sscase$A

plot_DAWG(ssgamma)
```

That can't be right. Rather, look at deviation from mean (in space) of true Qn matrix.

```{r}
Qnmean <- apply((sscase$Q * ssnmat)^(3/5), 2, mean) %>% 
  swot_vec2mat(pattern = sscase$W)

Qndev <- (sscase$Q * ssnmat)^(3/5) - Qnmean

plot_DAWG(ssnmat)
plot_DAWG(sscase$S)
plot_DAWG(Qndev)
plot_DAWG(manning_ws35(sscase))
plot_DAWG(Qndev / manning_ws35(sscase))
```

That's getting close, but still needs to be better elucidated and formalized. There is not much slope variability compared to closure error. 


Check what the model looks like in the corrected case. 

```{r}
xmat_real <- manning_ws35(sscase) - Qndev

omegas <- make_omegas(nrow(xmat_real))
modmat_real <- map(omegas, function(x) t(xmat_real) %*% diag(x)) %>% 
  map(as.data.frame) %>% 
  bind_rows()
resp_real <- map(omegas, function(x) t(xmat_real * sscase$dA) %*% x) %>% 
  unlist()

moddf_real <- mutate(setNames(modmat_real, paste0("x", 1:ncol(modmat_real))),
                     y = resp_real)

lm_real <- lm(y ~ 0 + ., data = moddf_real)

summary(lm_real)
```

OK, math not working here...