---
title: "notebook20180407"
author: "Mark Hagemann"
date: "April 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Still struggling with how to handle "error in variables" model. But I'm making progress by stripping things down to components. Starting with closure term. I have that math jotted down, but good to put it in tex form here. Important parts:


$$
(Q_{t} \bar{n}_{\cdot t})^{3/5} = A_{it} (W_{it}^{-2/3} S_{it}^{1/2} \dot{n}_{it}^{-1})^{3/5}
$$

where:

- $\bar{n}_{\cdot t} \times \dot{n}_{it} = c_{it}$
- $c_{it} = \frac{1}{Q_t} A_{it}^{5/3} W_{it}^{-2/3} S_{it}^{1/2}$ 

Note that:

- space-invariant (mass-conserved) part is on left-hand side
    - Any averaging function can be used for to transfer $Q_{it}$ to $Q_t$.  
- RHS of course must also be mass-conserved, but here that is accomplished by complicated product of quantities that do vary in time (including closure term)
- closure term $c_{it}$ encompasses multiple sources of error, including:
    - Spatial variability in Manning's n (could be accounted for via $\dot{n}$)
    - Temporal (e.g. depth) variability in Manning's n
    - Flow imbalance (not mass conserved and/or not steady state)
    - Manning equation error
    
Now I still need to apply this to get the following:

$$
\mathbf{W = X + U}
$$

Where W is observed, X is real, and U is error. 

Suppose I try to use W, S, and dA observations to estimate A0. Assume no measurement error. No a priori $\dot{n}_{it}$. Then what is W and what is U? W is ws35 %*% Omega. X could be ws35 / closure, but this would be very different from W. But I can scale ws35 / closure by multiplying it by anything that preserves conservation of mass (steady-state). That means I can mutliply ws35 / closure by any matrix created from an arbitrary time vector. Optimally I would do this to match row means of W. 

```{r}
casei <- sscase

wmati <- manning_ws35(casei)
closi <- manning_closure(casei)
xmati_1 <- wmati * closi ^ (-3/5)

```

Verify that algorithm works for X. 

```{r}
summary(estA0_lm_ws35(xmati_1, casei$dA))
```

Good. Now modify X to match row means of W. 

```{r}
rmeans_W <- apply(wmati, 2, mean)
rmeans_X <- apply(xmati_1, 2, mean)
xmati_2 <- xmati_1 * swot_vec2mat(rmeans_W / rmeans_X, xmati_1)
```

Plot and verify similarity. Also verify performance of new X matrix. 

```{r}
plot_DAWG(xmati_1)
plot_DAWG(xmati_2)
plot_DAWG(wmati)
```

That much is good. Now does it work?

```{r}
summary(estA0_lm_ws35(xmati_2, casei$dA))
```

Yes! Now what I really need is difference after it's been multiplied by omega. 

```{r}
omegaProduct <- function(ws35mat, symmetric = FALSE) {
  omegas <- make_omegas(nrow(ws35mat), symmetric = symmetric)
  out <- map(omegas, function(x) t(ws35mat) %*% diag(x)) %>% 
    Reduce(rbind, .)
  out
}

truewmat <- omegaProduct(wmati, symmetric = TRUE)
truexmat <- omegaProduct(xmati_2, symmetric = TRUE)
trueumat <- truexmat - truewmat
```

How does U behave?

```{r}
mean(truexmat)
mean(truewmat)
mean(trueumat)

sd(truexmat)
sd(truewmat)
sd(trueumat)
```

Doesn't quite have zero mean! Which is to be expected. 

```{r}
pairs(trueumat)
```

Not quite independent either. But close enough. 

What if I use symmetric omegas? Then will it have same mean? Yes! And mean will always be zero! 

- Later on I may need to revisit this, since I'm artificially increasing the number of rows in the data matrix, without adding any further information. 

Do I have a way to quickly get the U matrix from a given W matrix? 

- Not yet. What should it do?
- Given a ws35 matrix, calculate the truth matrix that has the same mean row-by-row.
- Rely on manning_closure function.

```{r}

# Obsmat should be a WS35 matrix, and by default will be calculated as ws35 for 
# the supplied swotlist. The output will be coerced to have the same row-by-row 
# mean as obsmat. 

manning_linA_closed <- function(swotlist, obsmat = NULL, adjust = TRUE) {
  if (is.null(obsmat)) {
    obsmat <- manning_ws35(swotlist = swotlist)
  }
  clos <- manning_closure(swotlist)
  xmat <- obsmat * closi ^ (-3/5)
  
  rmeans_W <- apply(obsmat, 2, mean)
  rmeans_X <- apply(xmat, 2, mean)
  
  if (adjust) {
    out <- xmat * swot_vec2mat(rmeans_W / rmeans_X, xmat)
  } else {
    out <- xmat
  }
  
  out
}

xmat3 <- manning_linA_closed(casei)

op3 <- omegaProduct(xmat3, symmetric = TRUE)

op3.1 <- omegaProduct(manning_ws35(casei), symmetric = TRUE)

summary(op3)
summary(op3.1)
```

Now I can make the closure!

```{r}
manning_linA_closure <- function(swotlist, obsmat = NULL) {
  if (is.null(obsmat)) {
    obsmat <- manning_ws35(swotlist = swotlist)
  }
  
  closedmat <- manning_linA_closed(swotlist = swotlist, obsmat = obsmat)
  out <- closedmat - obsmat
  out
}
```

OK, *now* can I do the simulation? Almost. Where does the error magnitude stand currently?

```{r}
sd(omegaProduct(manning_linA_closure(casei)))
sd(omegaProduct(manning_ws35(casei)))

```

Really small, apparently! Am I sure this closure works?

```{r}
(manning_linA_closure(casei) + manning_ws35(casei)) %>% 
  estA0_lm_ws35(dAmat = casei$dA) %>% 
  summary()

(manning_ws35(casei)) %>% 
  estA0_lm_ws35(dAmat = casei$dA) %>% 
  summary()
```

Yup! It's all working! Figure out parameters of simulation study. 

```{r}
(manning_linA_closure(casei)) %>% 
  # mean()
  apply(1, mean)

(manning_ws35(casei)) %>% 
  # mean()
  apply(1, mean)
  
(manning_linA_closure(casei)) %>% 
  t() %>% 
  pairs()
```

Definitely some dependency stucture here. Also column-wise means of closure matrix are not zero. But for simulation purposes I'll make everything zero-mean and independent. 

Figure will plot sd of closure matrix against error in A0. I'll need to do a statistically large number of A0 inversions for each level of error. Do that tonight, then tomorrow work on doing similar for synthetic cases. 

```{r}
simsds <- 10^seq(-8, -2, length.out = 20) %>% 
  rep(each = 100)

nx <- nrow(casei$W)
nt <- ncol(casei$W)

simerrmats <- map(simsds, ~rnorm(nx * nt, 0, .)) %>% 
  map(~matrix(., nrow = nx, ncol = nt))
simwsmats <- map(simerrmats, ~(manning_linA_closed(casei) + .))

simA0ests <- map(simwsmats, ~estA0_lm_ws35(., dAmat = casei$dA)) %>% 
  map(~as.data.frame(as.list(coef(.)))) %>% 
  bind_rows(.id = "simno")

trueA0df <- data.frame(loc = c("x1", "x2", "x3"), A0true = c(100, 122.15, 100))

simoutdf <- simA0ests %>% 
  mutate(sd = simsds) %>% 
  select(-simno) %>% 
  gather(key = "loc", value = "A0est", -sd) %>% 
  left_join(trueA0df, by = c("loc")) %>% 
  mutate(relerr_pct = (A0true - A0est) / A0true * 100)
```

Now plot it and call it a night!

```{r}
simoutdf %>% glimpse() %>% 
  mutate(absrelerr = abs(relerr_pct)) %>% 
  # ggplot(aes(x = sd, y = relerr_pct)) +
  ggplot(aes(x = sd, y = absrelerr)) +
  geom_point() +
  scale_x_log10() + 
  scale_y_log10()
```

Last thing: Put actual error from sscase in there. 

```{r}

valdf <- data.frame(loc = c("x1", "x2", "x3"),
                    sd = apply(manning_linA_closure(casei), 1, sd),
                    A0est = estA0(casei)) %>% 
  left_join(trueA0df, by = "loc") %>% 
  mutate(relerr_pct = (A0true - A0est) / A0true * 100,
         absrelerr = abs(relerr_pct))
  


simoutdf %>% glimpse() %>% 
  mutate(absrelerr = abs(relerr_pct)) %>% 
  # ggplot(aes(x = sd, y = relerr_pct)) +
  ggplot(aes(x = sd, y = absrelerr)) +
  geom_point() +
  geom_point(data = valdf, color = "red", size = 5) +
  scale_y_log10() +
  scale_x_log10() +
  theme_bw()

# ggsave("../graphs/A0_ss_sim1_log.png")

```

Nice! That's a wrap! 