---
title: "notebook20180522"
author: "Mark Hagemann"
date: "May 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Trying weighted regression based on distance for estA0. See how well it improves things. 

```{r}

realA0 <- map(reachdata, function(x) (x$A - x$dA)[, 1])
# A0hat1 <- map(reachdata, possibly(estA0, otherwise = NA))

nse1 <- map2_dbl(A0hat1, realA0, NSE)
rrmse1 <- map2_dbl(A0hat1, realA0, RRMSE)
nrmse1 <- map2_dbl(A0hat1, realA0, NRMSE)

# Now I sloppily change the estA0 code to add weights
# A0hat2 <- map(reachdata, possibly(estA0, otherwise = NA))

nse2 <- map2_dbl(A0hat2, realA0, NSE)
rrmse2 <- map2_dbl(A0hat2, realA0, RRMSE)
nrmse2 <- map2_dbl(A0hat2, realA0, NRMSE)

plot(nse1, nse2); abline(0, 1)
plot(rrmse1, rrmse2); abline(0, 1)
plot(nrmse1, nrmse2); abline(0, 1)

# Try inverse square weights.
# A0hat3 <- map(reachdata, possibly(estA0, otherwise = NA))

nse3 <- map2_dbl(A0hat3, realA0, NSE)
rrmse3 <- map2_dbl(A0hat3, realA0, RRMSE)
nrmse3 <- map2_dbl(A0hat3, realA0, NRMSE)

plot(nse1, nse3); abline(0, 1)
plot(rrmse1, rrmse3); abline(0, 1)
plot(nrmse1, nrmse3); abline(0, 1)


plot(nse2, nse3); abline(0, 1)
plot(rrmse2, rrmse3); abline(0, 1)
plot(nrmse2, nrmse3); abline(0, 1)

 
```

Inverse weights are the way to go. 

```{r}
wtdf <- data.frame(nse = c(nse1, nse2, nse3), 
                   rrmse = c(rrmse1, rrmse2, rrmse3),
                   nrmse = c(nrmse1, nrmse2, nrmse3),
                   case = rep(names(nse1), 3),
                   weights = rep(c("none", "inverse", "inv.sq"), 
                                 each = length(nse1))) %>% 
  gather(key = "stat", value = "value", -weights, -case) %>% 
  mutate(weights = factor(weights, levels = c("none", "inverse", "inv.sq")),
         stat = factor(stat, levels = c("rrmse", "nrmse", "nse")))
wtdf %>% 
  ggplot(aes(x = stat, y = value, color = weights)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge()) +
  theme_bw()
ggsave("estA0_wtcompare_boxplot.png")

wtdf %>% 
  filter(stat == "rrmse") %>% 
  ggplot(aes(x = weights, y = value, group = case, color = case)) +
  geom_point() + 
  geom_line(size = 0.5) +
  theme_bw()
ggsave("estA0_wtcompare_rrmse.png")

wtdf %>% 
  filter(stat == "nrmse") %>% 
  ggplot(aes(x = weights, y = value, group = case, color = case)) +
  geom_point() + 
  geom_line(size = 0.5) +
  theme_bw()
ggsave("estA0_wtcompare_nrmse.png")

wtdf %>% 
  filter(stat == "nse") %>% 
  ggplot(aes(x = weights, y = value, group = case, color = case)) +
  geom_point() + 
  geom_line(size = 0.5) +
  theme_bw()
ggsave("estA0_wtcompare_nse.png")

```

OK, that's that. Now I need to fix up these stan models so I can present something on them.

### Working on Bayesian models

in batman.R

- nlm works if I give it a low enough steptol. 
- Compare across estimation methods

```{r}
poA0_logbm <- batman_log(reachdata$Po, iterlim = 500)
poA0_linQbm <- batman_linQ(reachdata$Po, iterlim = 500)

A0ests_log <- reachdata %>% 
  map(possibly(~batman_log(., iterlim = 500), otherwise = NA))
map_int(A0ests_log[!is.na(A0ests_log)], ~.[["code"]])
estslog <- map(A0ests_log[!is.na(A0ests_log)], ~.[["A0"]])

A0ests_linA <- reachdata %>% 
  map(possibly(~batman_linA(., iterlim = 500), otherwise = NA))
map_int(A0ests_linA[!is.na(A0ests_linA)], ~.[["code"]])
estslinA <- map(A0ests_linA[!is.na(A0ests_linA)], ~.[["A0"]])

A0ests_linQ <- reachdata %>% 
  map(possibly(~batman_linQ(., iterlim = 500), otherwise = NA))
map_int(A0ests_linQ[!is.na(A0ests_linQ)], ~.[["code"]])
estslinQ <- map(A0ests_linQ[!is.na(A0ests_linQ)], ~.[["A0"]])

trueA0 <- map(reachdata, function(x) (x$A - x$dA)[, 1])

A0ests_lm <- reachdata %>% 
  map(possibly(estA0, otherwise = NA))
estslm <- A0ests_lm[!is.na(A0ests_lm)]


rrmselog <- map2_dbl(estslog, trueA0[names(estslog)], RRMSE)
rrmselinQ <- map2_dbl(estslinQ, trueA0[names(estslinQ)], RRMSE)
rrmselinA <- map2_dbl(estslinA, trueA0[names(estslinA)], RRMSE)
rrmselm <- map2_dbl(estslm, trueA0[names(estslm)], RRMSE)

coms1 <- intersect(names(rrmselog), names(rrmselinA))
plot(rrmselog[coms1], rrmselinA[coms1]); abline(0, 1)

coms2 <- intersect(names(rrmselinA), names(rrmselm))
plot(rrmselinA[coms2], rrmselm[coms2]); abline(0, 1)

coms3 <- intersect(names(rrmselinQ), names(rrmselinA))
plot(rrmselinQ[coms3], rrmselinA[coms3]); abline(0, 1)
which.max(rrmselinA[coms3])
```

Next make these models in Stan and validate results.

A0_v4-1.stan should be the same as A0ests_log. 

```{r}
library(rstan)
stanmodlog <- stan_model("../stan/A0_v4-1.stan")

prepdata <- function(swotlist) {
  list0 <- with(swotlist, list(
  nt = ncol(W),
  nx = nrow(W),
  dA = dA,
  W = W, 
  S = S, 
  logA0_hat = log(max(dA) - min(dA)),
  logn_hat = -3.5,
  sigma_logn = 0.5))
  
  list1 <- with(c(swotlist, list0), 
    list(sigma_logA0 = 1, 
         logQ_hat = mean(5/3 * logA0_hat - 2/3 * log(W) + 1/2 * log(S) - logn_hat), 
         sigma_logQ = 1))
  out <- c(list0, list1)
  out
}

reachdata$Po %>%
  prepdata() %>% 
  optimizing(stanmodlog, data = ., as_vector = FALSE) %>% 
  (function(x) x$par$A0)

sscase %>%
  prepdata() %>% 
  optimizing(stanmodlog, data = ., as_vector = FALSE) %>% 
  (function(x) x$par$A0)

uscase %>%
  prepdata() %>% 
  optimizing(stanmodlog, data = ., as_vector = FALSE) %>% 
  (function(x) x$par$A0)

sssamps <- sscase %>%
  prepdata() %>%
  sampling(stanmodlog, data = ., chains = 3, cores = 3, iter = 10000)

stan_trace(sssamps, par = "A0", inc_warmup = TRUE)$data %>%
  # glimpse() %>%
  ggplot(aes(x = iteration, y = value, color = chain, group = chain)) +
  geom_line(size = 0.5) +
  facet_wrap(~parameter, ncol = 2)

stan_trace(sssamps, par = "Qn", inc_warmup = TRUE)$data %>%
  # glimpse() %>%
  ggplot(aes(x = iteration, y = value, color = chain, group = chain)) +
  geom_line(size = 0.5) +
  facet_wrap(~parameter, ncol = 2)

ussamps <- uscase %>%
  prepdata() %>%
  sampling(stanmodlog, data = ., chains = 3, cores = 3, iter = 10000)

stan_trace(ussamps, par = "A0", inc_warmup = TRUE)$data %>%
  # glimpse() %>%
  ggplot(aes(x = iteration, y = value, color = chain, group = chain)) +
  geom_line(size = 0.5) +
  facet_wrap(~parameter, ncol = 2)

ussamps_adj <- uscase %>%
  swot_timelag(c(0, -2, -4)) %>%
  prepdata() %>%
  sampling(stanmodlog, data = ., chains = 3, cores = 3, iter = 10000)

stan_trace(ussamps_adj, par = "A0", inc_warmup = TRUE)$data %>%
  # glimpse() %>%
  ggplot(aes(x = iteration, y = value, color = chain, group = chain)) +
  geom_line(size = 0.5) +
  facet_wrap(~parameter, ncol = 2)

```

Now estimate for all Pepsi cases

```{r}

rddata <- reachdata %>% 
  map(prepdata)

rdsamps <- rddata %>% 
  map(possibly(~sampling(stanmodlog, data = ., cores = 3, chains = 3, iter = 10000), 
               otherwise = NA))

rdpostmeans <- rdsamps %>% 
  map(get_posterior_mean, par = "A0") %>% 
  map(~.[, "mean-all chains"])
s
nulls <- map_lgl(rdpostmeans, is.null)

trueA0_min <- reachdata %>% 
  map(function(x) (x$A - rezero_dA(x$dA, zero = "minimum"))[, 1])

makedf <- function(x, y) data.frame(est = x, truth = y)

map2_lgl(rdpostmeans[!nulls], trueA0_min[!nulls], function(x, y) length(x) == length(y))

compdf <- map2(rdpostmeans[!nulls], trueA0_min[!nulls], makedf) %>% 
  bind_rows(.id = "case")

ggplot(compdf, aes(x = est, y = truth)) +
  geom_point(aes(color = case)) +
  geom_abline(slope = 1, intercept = 0)

```

Check convergence

```{r}
rhats <- map(rdsamps, possibly(~stan_rhat(., "A0")$data, otherwise = NA))

rhatdf <- rhats[!is.na(rhats)] %>%
  bind_rows(.id = "case") %>% glimpse()

# rhatdf_2k <- rhats[!is.na(rhats)] %>% 
#   bind_rows(.id = "case") %>% glimpse()

ggplot(rhatdf, aes(x = case, y = stat)) +
  geom_boxplot() + 
  geom_point()

ggplot(rhatdf_2k, aes(x = case, y = stat)) +
  geom_boxplot() + 
  geom_point()


plot(rhatdf$stat - rhatdf_2k$stat)
```

