---
title: "notebook20180410"
author: "Mark Hagemann"
date: "April 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Still trying to tease out effects of error (and specific parts of error structure) on A0 inversions. 

I think it's time to add in distance and timelag information. This will allow me to approximate Q variability in space. 

- This is likely to be a very crude estimate, and biased low, for long time intervals (since dA/dt will be smaller than it should be.)
- Math is:

$$
\Delta Q \approx \frac{\Delta A}{\Delta t} \Delta x
$$


The objective is to get "characteristic values" for dA/dt for each river, then get the corresponding for Q variability. Then take 

$$
Qvar = sum(\Delta Q)
$$

Let's try it! How far back do I need to go to get this info?

```{r}
nclists$Connecticut$Reach_Timeseries.t
nclists$Connecticut$River_Info.rch_bnd

nci <- nclists$Connecticut
dati <- reachdata$Connecticut
tt <- nci$Reach_Timeseries.t
xx <- nci$River_Info.rch_bnd
xx <- (xx[-1] + xx[-length(xx)]) / 2
xx <- xx[nci$River_Info.gdrch]

dA <- apply(dati$dA, 1, diff)
dt <- diff(tt * 3600 * 24)

dQreal <- median(abs(apply(dati$Q, 2, diff)))

dx <- diff(xx[-1] + xx[-length(xx)]) / 2

dadt <- median(abs(dA / swot_vec2mat(dt, dA)))
dQest <- dadt * sum(dx)

dQreal
```

Doesn't look like a great estimate of dQ. I think I'm better off (somewhat) by estimating A0 and backing out dQ. 

OK, I'm going to make a short writeup about dQ/dx = dA/dt validation. 

Done with the analysis for now (to my satisfaction). Will need polishing before sharing, but that's for a later time. 


### depth-variability of Manning's n

Is this a relationship that can be transferred from one case to another?

```{r}
# nddf <- reachdata %>%
#   map(possibly(val_anova_lm, otherwise = NULL)) %>% 
#   map(tidy) %>% 
#   bind_rows(.id = "case")

nddf <- reachdata %>% 
  map(~val_nd_plot(., plot = FALSE)) %>% 
  bind_rows(.id = "case")

ndlms <- nddf %>% 
  filter(!is.na(loc), !is.na(logn), !is.na(logd)) %>% 
  split(f = .$case) %>% 
  map(~mutate(., loc = as.factor(loc))) %>% 
  map(~lm(logn ~ logd + loc, .))

ndevaldf <- ndlms %>% 
  map(tidy) %>% 
  bind_rows(.id = "case")

ndevaldf %>%
  filter(term == "logd") %>% 
  glimpse() %>% 
  ggplot(aes(x = case, y = estimate, ymin = estimate - 2 * std.error,
             ymax = estimate + 2 * std.error)) +
  geom_pointrange() +
  theme(axis.text.x = element_text(angle = 90))

```

```{r}
val_nd_plot(reachdata$SacramentoUpstream)

plot_DAWG(manning_ndot(reachdata$SacramentoUpstream))
```

