---
title: "notebook20180418"
author: "Mark Hagemann"
date: "April 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Yesterday I finally got the Po validation off my plate. I also made the decomposition of gamma term better, and shored up the math. Still *some* work to do on that front, but first I need to get to the bottom of another finding from yesterday: apparent bias in closure term. Since this is multiplicative, it results in sloping residuals. What causes this? Can I fix this?

First, observe what I mean:

```{r}
casei <- uscase
# casei <- sscase
# casei <- reachdata$Po
# casei <- reachdata$Severn

yobs <- omegaProduct_y(manning_ws35(casei), casei$dA, symmetric = TRUE)
ytru <- omegaProduct_y(manning_linA_closed(casei), casei$dA, symmetric = TRUE)

plot(yobs ~ ytru); abline(0, 1)

```

Can I show this comes from biased closre term?

```{r}
xclos <- manning_linA_closure(casei)

xclos <- manning_linA_closed(casei) / manning_ws35(casei)

xclos %>% 
  hist()

hist(casei$dA)
hist(manning_ws35(casei) * casei$dA)
hist(manning_linA_closed(casei) * casei$dA)
```

Hmm. Let's be more clear about what I want. From notebook work:

$$
M_{it}\delta A_{it}E[C_{it} - 1] - M_{i't}\delta A_{i't}E[C_{i't} - 1] = M_{it}\delta A_{it} - M_{i't}\delta A_{i't}
$$

So really just $E[C_{it}] = 1$

```{r}
mean(manning_closure(casei, log = FALSE))
```

Do I need a new way of making the closure term? Probably.

```{r}
foo <- manning_ws35(casei) / manning_linA_closed(casei)

mean(foo)
geomMean(foo)
hist(rowMeans(foo))
hist(rowMeans(t(foo)))
```

### After consuldting previous notes on closure

See notebook20180407. I can multiply closure by any time vector. 

```{r}
plot(yreal)
plot(yobs)
plot(yreal / yobs)
plot(yobs / yreal)
```

```{r}
yreal <- omegaProduct_y(manning_linA_closed(casei), casei$dA, symmetric = TRUE)
closi <- manning_ws35(casei) / manning_linA_closed(casei)

rhserri <- omegaProduct_y(closi, dAmat = casei$dA, symmetric = TRUE)

omegsi <- make_omegas(nrow(casei$W), symmetric = TRUE)

rhserri <- map(omegsi, function(x) t(closi) %*% x) %>% 
  unlist()

plot(rhserri)
plot(yreal, rhserri)

plot(closi, manning_linA_closed(casei))
```


Can I change this last relationship by modifying closi?

```{r}
closi %>% dim()

mclosi <- manning_linA_closed(casei, adjust = TRUE)
munclosi <- manning_ws35(casei)
closi <- mclosi / munclosi

plot(closi, mclosi)
plot(closi, munclosi)

cor(as.vector(closi), as.vector(mclosi))
cor(as.vector(closi), as.vector(munclosi))

plot(closi[, 1:10], mclosi[, 1:10])
plot(closi[, 18], mclosi[, 18])

plot_DAWG(munclosi)
plot_DAWG(closi)
plot_DAWG(mclosi)
```

Can I scale the columns of X?


