---
title: "notebook20180828"
author: "Mark Hagemann"
date: "August 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I finally have a workable gce + docker workflow for comparing results of different bamr branches. Now it's time to do those comparisons so I can push forward on improving bamr. Start by making a utility function for cluster planning.

```{r}
# Based on work in notebook20180823 and online tutorial linked therein.

library(future)
docker_clustfun <- function(ext_ip, docker_image, user = "markwh", 
                            keyfile = "~/.ssh/google_compute_engine", ...) {
  cl <- makeClusterPSOCK(
  ext_ip,

  # User name; DigitalOcean droplets use root by default
  user = user,

  # Use private SSH key registered with DigitalOcean
  rshopts = c(
    "-o", "StrictHostKeyChecking=no",
    "-o", "IdentitiesOnly=yes",
    "-i", keyfile
  ),

  # Command to run on each remote machine
  # The script loads the tidyverse Docker image
  # --net=host allows it to communicate back to this computer
  rscript = c("sudo", "docker", "run", "--net=host", 
              docker_image, "Rscript"),

  # These are additional commands that are run on the remote machine. 
  # At minimum, the remote machine needs the future library to work—installing furrr also installs future.
  rscript_args = c(
    # Create directory for package installation
    "-e", shQuote("local({p <- Sys.getenv('R_LIBS_USER'); dir.create(p, recursive = TRUE, showWarnings = FALSE); .libPaths(p)})"),
    # Install furrr and future
    "-e", shQuote("if (!requireNamespace('furrr', quietly = TRUE)) install.packages('furrr')")
  ),

  # Actually run this stuff. Set to TRUE if you don't want it to run remotely.
  dryrun = FALSE,
  ...
)
}

```

Recall: I need to make a snapshot. Alternatively I can just clone the instances from the console. I think I'll do that. 

The workflow will then be:

```{r}
testcases <- reachdata[c("Cumberland", "Ganges", "SacramentoUpstream")]

# Start the vm's (do this externally on command line)

clust1 <- docker_clustfun(ext_ip = gce_get_external_ip("rocker2", 
                                                       zone = "us-east4-b"), 
                          user = "markwh",
                          docker_image = "markwh/bam-docker:master", 
                          # makeNode = myMakenode,
                          # keyfile = "C:/Users/markh/.ssh/google_compute_engine",
                          verbose = TRUE)

clust2 <- docker_clustfun(ext_ip = gce_get_external_ip("rocker3", 
                                                       zone = "us-east4-b"),
                          docker_image = "markwh/bam-docker:devel")
clust3 <- docker_clustfun(ext_ip = gce_get_external_ip("rocker4", 
                                                       zone = "us-east4-b"),
                          docker_image = "markwh/bam-docker:consolidate")
```

Now run!

```{r}

futureAssign("evals1_master", value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_be <- purrr::map(test_bd, 
                            ~bamr::bam_estimate(., 
                              variant = "manning", 
                              reparam = TRUE, 
                              meas_error = TRUE))}, 
             evaluator = tweak(cluster, workers = clust1))
futureAssign("evals1_devel", 
   value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_be <- purrr::map(test_bd, 
                            ~bamr::bam_estimate(., 
                              variant = "manning", 
                              reparam = TRUE, 
                              meas_error = TRUE))},
             evaluator = tweak(cluster, workers = clust2))
futureAssign("evals1_consolidate",  
    value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_be <- purrr::map(test_bd, 
                            ~bamr::bam_estimate(., 
                              variant = "manning", 
                              reparam = TRUE, 
                              meas_error = TRUE))},
             evaluator = tweak(cluster, workers = clust3))


```

```{r}

futureAssign("evals2_master", value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_bp <- purrr::map(test_bd, 
                      ~bam_priors(., variant = "manning_amhg",
                                  logA0_sd = 0.5, logn_sd = 1))
      test_be <- purrr::map2(test_bd, test_bp,
                            ~bamr::bam_estimate(.x, bampriors = .y,
                              variant = "manning_amhg", 
                              reparam = TRUE, 
                              meas_error = TRUE))}, 
             evaluator = tweak(cluster, workers = clust1))
futureAssign("evals2_devel", 
   value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_bp <- purrr::map(test_bd, 
                            ~bam_priors(., variant = "manning_amhg",
                                        logA0_sd = 0.5, logn_sd = 1))
      test_be <- purrr::map2(test_bd, test_bp,
                            ~bamr::bam_estimate(.x, bampriors = .y,
                              variant = "manning_amhg", 
                              reparam = TRUE, 
                              meas_error = TRUE))},
             evaluator = tweak(cluster, workers = clust2))
futureAssign("evals2_consolidate",  
   value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_bp <- purrr::map(test_bd, 
                      ~bam_priors(., variant = "manning_amhg",
                                  logA0_sd = 0.5, logn_sd = 1))
      test_be <- purrr::map2(test_bd, test_bp,
                            ~bamr::bam_estimate(.x, bampriors = .y, 
                              variant = "manning_amhg", 
                              reparam = TRUE, 
                              meas_error = TRUE))},
             evaluator = tweak(cluster, workers = clust3))

```



```{r}

futureAssign("evals3_master", value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_be <- purrr::map(test_bd, 
                            ~bamr::bam_estimate(., 
                              variant = "manning", 
                              # reparam = TRUE, 
                              meas_error = FALSE))}, 
             evaluator = tweak(cluster, workers = clust1))
futureAssign("evals3_devel", 
   value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_be <- purrr::map(test_bd, 
                            ~bamr::bam_estimate(., 
                              variant = "manning", 
                              # reparam = TRUE, 
                              meas_error = FALSE))},
             evaluator = tweak(cluster, workers = clust2))
futureAssign("evals3_consolidate",  
   value = {
      test_bd <- purrr::map(testcases, ~swot_bamdata(.))
      test_be <- purrr::map(test_bd, 
                            ~bamr::bam_estimate(., 
                              variant = "manning", 
                              reparam = TRUE, 
                              meas_error = FALSE))},
             evaluator = tweak(cluster, workers = clust3))
```


devel not working. Check a single case. OK, so rerunning the cluster definition fixed it. 

Now compare the evals. 

```{r}
evals1_master$Cumberland %>% 
  bam_hydrograph()
evals1_devel$Cumberland %>% 
  bam_hydrograph()
evals1_consolidate$Cumberland %>% 
  bam_hydrograph()

evals1_master$Ganges %>% 
  bam_hydrograph()
evals1_devel$Ganges %>% 
  bam_hydrograph()
evals1_consolidate$Ganges %>% 
  bam_hydrograph()


evals1_master$SacramentoUpstream %>% 
  bam_hydrograph()
evals1_devel$SacramentoUpstream %>% 
  bam_hydrograph()
evals1_consolidate$SacramentoUpstream %>% 
  bam_hydrograph()


```


```{r}
evals2_master$Cumberland %>% 
  bam_hydrograph()
evals2_devel$Cumberland %>% 
  bam_hydrograph()
evals2_consolidate$Cumberland %>% 
  bam_hydrograph()

evals2_master$Ganges %>% 
  bam_hydrograph()
evals2_devel$Ganges %>% 
  bam_hydrograph()
evals2_consolidate$Ganges %>% 
  bam_hydrograph()

evals2_master$SacramentoUpstream %>% 
  bam_hydrograph()
evals2_devel$SacramentoUpstream %>% 
  bam_hydrograph()
evals2_consolidate$SacramentoUpstream %>% 
  bam_hydrograph()

```


Jeez, that took forever! But now I have the answers I need. Make sure I can cache the objects in this notebook. 
