---
title: "notebook20180402"
author: "Mark Hagemann"
date: "April 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have some more to do on writing up and pushing forward A0 inversion from toy case, but for now it's time to tie up some loose ends.

First, Wilammette discharge using A0 ests and n = 0.03. 

```{r}
hc_h <- read.csv("../../A0_identifiability/data/HydraulicControlReaches/H.txt", sep = "\t", header = FALSE) %>% 
  as.matrix()
hc_w <- read.csv("../../A0_identifiability/data/HydraulicControlReaches/W.txt", sep = "\t", header = FALSE) %>% 
  as.matrix()
hc_s <- read.csv("../../A0_identifiability/data/HydraulicControlReaches/S.txt", sep = "\t", header = FALSE) %>% 
  as.matrix()
hc_q <- read.csv("../../A0_identifiability/data/HydraulicControlReaches/Qobs.txt", sep = "\t", header = FALSE) %>% as.matrix()

hclist <- list(H = hc_h, W = hc_w, S = hc_s, Q = hc_q)

swot_plot(hclist)

hclist$dA <- calcdA_mat(w = hclist$W, h = hclist$H, zero = "minimum")

swot_plot(hclist)
```

Now estimate A0

```{r}
foo <- hclist %>% 
  ccf_lag(Ahat = TRUE, verbose = TRUE)

plot_DAWG(manning_qdot(hclist, Avar = "Ahat")) + scale_y_log10()
```

```{r}
hclist$Ahat <- swot_A(A0vec = estA0(swotlist = hclist, zero = "minimum"), 
                      dAmat = hclist$dA)
qhat <- with(hclist, manningQ(A = Ahat, W = W, S = S, n = 0.03))

with(hclist, 1 / 0.03 * Ahat^(5/3) * W^(-2/3) * S^(1/2)) %>% 
  plot_DAWG()

plot_DAWG(qhat)

```


That's not so good. Based on qdot plot, let's try subsetting the reaches. 

```{r}
hcl_ss1 <- swot_sset(hclist, keeplocs = 1:10)

estA0(hcl_ss1)

hcl_ss1$Ahat <- swot_A(A0vec = estA0(swotlist = hcl_ss1, zero = "minimum"), 
                      dAmat = hcl_ss1$dA)
qhat <- with(hcl_ss1, manningQ(A = Ahat, W = W, S = S, n = 0.03))

with(hcl_ss1, 1 / 0.03 * Ahat^(5/3) * W^(-2/3) * S^(1/2)) %>% 
  plot_DAWG()

plot_DAWG(qhat)

manning_nplot(hcl_ss1)
```

Nothing doing. I could try stepwise. First, look at VIF

```{r}
library(car)

hcl_lm <- estA0_lm(hclist)
coef(hcl_lm)
A0_condno(hcl_lm)


vif(hcl_lm)

```

Now I have Q data. If n = 0.03, then 

- What must A0, A be?
- How much worse (or better?) does real A0 fit data than estimated A0?

```{r}
hcl2 <- hclist
hcl2$A <- with(hcl2, (Q * 0.03 * W^(2/3) * S^(-1/2))^(3/5))

swot_plot(hcl2)

# realA0s <- apply(hcl2$A, 1, min)
realA0s <- hcl2$A - hcl2$dA
plot_DAWG(realA0s)

hcl2_A0lms <- estA0_lm(hcl2)
hcl2_A0ests <- estA0(hcl2)

plot(realA0s, estA0(hcl2))

omegas <- make_omegas(length(realA0s))
Omegas <- map(omegas, diag)


ws35 <- manning_ws35(hcl2)
lhs_x <- map(Omegas, function(x) t(ws35) %*% x) %>% 
  Reduce(f = rbind)
lhs <- lhs_x %*% realA0s
rhs <- map(omegas, function(x) t(-hcl2$dA * ws35) %*% x) %>% 
  Reduce(f = c)

plot(lhs, rhs); abline(0, 1)

plot(lhs, predict(hcl2_A0lms))
```

This should give a perfect fit, and yet it doesn't. Why?

```{r}

omind <- 6
ws35 <- manning_ws35(hcl2)
testlhs <- t(ws35) %*% Omegas[[omind]] %*% realA0s
testrhs <- (t(ws35) * t(hcl2$dA)) %*% -omegas[[omind]]

plot(testlhs, testrhs); abline(0, 1)


foo <- with(hcl2, A == (dA + swot_A(A0vec = realA0s, dAmat = dA)))
plot_DAWG(hcl2$A)
plot_DAWG(with(hcl2, (dA + swot_A(A0vec = realA0s, dAmat = dA))))
```

Try again with an updated A0-ndot joint estimate.  But first, revisit Po and Niger estimates.


## Po validation, Niger prediction

```{r}
load("../../airSWOT/cache/est_po.RData")
load("../../airSWOT/cache/est_niger.RData")

class(est_po)
library(bamr)
bam_hydrograph(est_po)
```

I now have validation data for Po. But it's an excel file with multiple gages listed. Time to look at the raw data?

```{r}

podata <- read_metroman("../../airSWOT/data/Po.txt")
swot_plot(podata, vars = -1)

head(t(podata$H))


foo <- R.matlab::readMat("~/Downloads/gauge_all.mat")

str(foo[[1]])
str(foo[[1]][[1]][[1]], 1)


podsch <- foo[[1]] %>% 
  map(~.[[1]]) %>% 
  map(as.data.frame) %>% 
  map(~setNames(., c("Date", "stage", "level", "depth", "Q", "A"))) %>% 
  bind_rows(.id = "reach") %>% 
  transmute(Date = as.Date(Date, origin = "0000-01-01"),
            Date = Date - 1, reach, 
            H = level, D = depth, Q, A)

glimpse(podsch)
```

So I need to figure out which days are represented in the Po data. I have them numbered, but no idea where the numbers reference to, date-wise. 

```{r}
podsch %>% 
  ggplot(aes(x = Date, y = Q, color = ))
```


### Revisiting joint A0, ndot estimation

Try to visuialize why things fall apart.

```{r}
A01 <- estA0(hcl2)
hcl2$Ahat <- swot_A(A0vec = A01, dAmat = hcl2$dA)

plot_DAWG(manning_ws35(hcl2))
ndot1 <- swot_vec2mat(manning_ndot(hcl2, Avar = "Ahat"), pattern = hcl2$W)
plot_DAWG(ndot1)

plot_DAWG(manning_qdot(hcl2))
plot_DAWG(manning_qdot(hcl2) / ndot1)

A02 <- estA0(hcl2, ndot = ndot1)

ndot2 <- manning_ndot(swotlist = )
```

I fixed something (make ndot be a matrix), but now things are blowing up in the other direction.

```{r}
foo <- estA0_ndot(hcl2 %>% swot_sset(keeplocs = 1:6), iters = 100)

foo$Qbarn %>% unlist() %>% plot()
foo$Qbarn %>% unlist() %>% `/`(0.03) %>% plot(log = "y")
```


```{r}
hcl2$W %>% apply(1, min)
```


I need to put this in a stan program, probably.


I say that weights are 1, -1 because:

- mass is conserved at steady state
- n does not change in space

But really weights are different. If I knew them perfectly I could get A0. Compare 1, -1 to real weights. 

Also, how does sd(qdot) change as I iteratively change A0?

Things to track as I iterate:

- log-space sd(qdot)
- log-space mean(A0)
- log-space mean