---
title: "notebook20171207"
author: "Mark Hagemann"
date: "December 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today I'm making that darn presentation. 

Animation showing dA

```{r}
# copying this from SWOT project
calcdA_vec <- function(w, h) {
  words <- order(w)
  
  warr <- w[words]
  harr <- h[words]
  
  delh <- c(0, diff(harr))
  delA <- cumsum(warr * delh)
  
  dA <- 1:length(w)
  dA[words] <- delA
  dA
}
```

Maybe start with an image

```{r}
dA_ggdf <- train_full %>% 
  filter(!is.na(w_m), !is.na(h_m)) %>% 
  group_by(xs) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n < nn) %>%
  filter(n == max(n)) %>% 
  filter(xs == xs[1]) %>% 
  select(Date, w_m, h_m, dA, n, xs, xsname) %>% 
  mutate(w = w_m / 2)

# nn <- dA_ggdf$n[1]
nn <- 190.1 #


glimpse(dA_ggdf)

ggplot(dA_ggdf, aes(x = w_m, y = dA)) +
  geom_point()

plot(dA_ggdf$h_m)

ggplot(dA_ggdf, aes(x = w, y = h_m, xend = -w, yend = h_m)) +
  geom_segment()

ggplot()
```

Hmm. Try Pepsi data?

```{r}
load("../../SWOT/cache/nc_r.RData")
# load("../../swotData/data/Pepsi_v2.rda")

sapply(nc_r, function(x) ncol(x[["h"]]))

dA_ggdf <- with(nc_r$MississippiUpstream, 
  data.frame(w_m = as.vector(w[1, ]),
             h_m = as.vector(h[1, ]),
             dA = as.vector(dA[1, ]))) %>% 
  mutate(w = w_m / 2)

ggplot(dA_ggdf, aes(x = w, y = h_m, xend = -w, yend = h_m)) +
  geom_segment()
```

Picking up the next day. I'm ready to start making the animations for the AGU talk. I need functions to create the following:

- timeseries animation `ts_anim(obs)`
    - easy, fast, no MLE required
- histogram animation (no density) `hist_anim(obs)`
    - easy, fast, no MLE
- histogram animation (with density) `mle_hist_anim(obs)`
    - requires MLE for each timestep `mle_ts(factory, ...)`
- A0 estimate timeseries animation `mle_ts_anim(obs, mlets)`
    - requires MLE for each timestep `mle_ts(obs)`
    - return list of gg objects, one for each parameter
- Gridded likelihood animation `nll_contour_anim`
    - requires nll for each gridpoint, each timestep `nll_grid(xvec, yvec, nllFactory, ..., accum = FALSE)`


Make all of these simple functions of:

- dA vector
- prior vector (A0, sigmalogA_hat, sigmalogA_sd)

result should be gg object, ready for gganimate()


```{r}

```


