---
title: "notebook20180326"
author: "Mark Hagemann"
date: "March 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Forging ahead with function development, application to Pepsi and lisflood cases. 

- I think I can account for spatial differences in Manning's n when estimating A0. 
- Finish diagnostics on Manning applicability
    - with and without steady-state mass conservation
    - With and without Manning n adjustment
        - in time and space
- Be more explicit about swotlist structure
    - A vs Ahat
    - Keep separate estimated vs measured parts
- Come up with consistent naming scheme for funcitons (with eventual packages in mind)

### Manning's n spatial variability adjustment. 

Equation is

$$
\mathbf{X} \Omega \mathbf{a}_0 = (\mathbf{X \circ} \Delta) \mathbf{\omega}
$$

This comes from the equation 

$$
(Qn)^{3/5} = A(W^{-2/3}S^{1/2})^{3/5}
$$

But suppose n is not constant. Then the equation still holds, but my implementation is wrong if I assume constant n. Being explicit about location, I can write:

$$
(Q_in_i)^{3/5} = A_i(W_i^{-2/3}S_i^{1/2})^{3/5}
$$

Now write $n_i = c_i n^*$ such that $c_i$ is the geometric mean of the RHS. Then the following would be equal across locaitons:

$$
A_i(W_i^{-2/3}S_i^{1/2})^{3/5} \div \text{geomMean}(A_i(W_i^{-2/3}S_i^{1/2})^{3/5})
$$

But the geomMean is not knowable. Perhaps it could be converged upon? Algorithm would be:

- Estimate A0, assuming steady-state mass conservation and constant n. 
- Optionally ajdust via ccf_lag
- Center RHS in log space
- Repeat until convergence or max number of iterations is reached. 

Cool, at least that's squared away conceptually. I'll push on that this afternoon. Now better to work on Manning applicability and ANOVA


### Manning diagnostics using known Q and A


```{r}
swotlist <- reachdata$Platte

manning_lm_linA <- function(swotlist, mc = TRUE) {
  
  W <- swotlist$W
  S <- swotlist$S
  A <- swotlist$A
  Q <- swotlist$Q
  
  if (mc) {
    Qvec <- apply(Q, 2, median, na.rm = TRUE)
    Q <- matrix(rep(Qvec, nrow(Q)), nrow = nrow(Q), byrow = TRUE)
  }
  
  rhs <- as.data.frame(t((W^(-2/3) * S^(1/2))^(3/5)))
  lhs <- as.data.frame(t(Q^(3/5) / A))
  
  dfs <- map2(rhs, lhs, function(x, y) data.frame(x = x, y = y))
  lms <- map(dfs, ~lm(y ~ 0 + x, data = ., y = TRUE))
}

manning_R2_linA <- function(swotlist, mc = TRUE) {
  
  lms <- manning_lm_linA(swotlist, mc = mc)
  
  out <- map_dbl(lms, A0_R2)
  out
}
```

Let's see that in action!

```{r}
map(reachdata, possibly(~manning_R2_linA(., mc = TRUE), otherwise = NULL)) %>% 
  map(~data.frame(reach = names(.), R2 = .)) %>% 
  bind_rows(.id = "case") %>% 
  ggplot(aes(x = case, y = 1 - R2)) +
  geom_point() +
  theme(axis.text = element_text(angle = 90)) +
  ylim(0, 2) 

map(reachdata, possibly(~manning_R2_linA(., mc = FALSE), otherwise = NULL)) %>% 
  map(~data.frame(reach = names(.), R2 = .)) %>% 
  bind_rows(.id = "case") %>% 
  ggplot(aes(x = case, y = 1 - R2)) +
  geom_point() +
  theme(axis.text = element_text(angle = 90)) +
  ylim(0, 2)

```

It apparently works, but I'm a little stymied about mc toggling results. 


```{r}
manning_lm_log <- function(swotlist, mc = TRUE) {
  
  W <- swotlist$W
  S <- swotlist$S
  A <- swotlist$A
  Q <- swotlist$Q
  
  if (mc) {
    Qvec <- apply(Q, 2, median, na.rm = TRUE)
    Q <- matrix(rep(Qvec, nrow(Q)), nrow = nrow(Q), byrow = TRUE)
  }
  
  rhs <- as.data.frame(t(manning_qdot(W, S, A, log = TRUE)))
  lhs <- as.data.frame(t(log(Q)))
  
  dfs <- map2(rhs, lhs, function(x, y) data.frame(x = x, y = y))
  lms <- map(dfs, ~lm(y ~ x, data = ., y = TRUE))
}

manning_R2_log <- function(swotlist, mc = TRUE) {
  
  lms <- manning_lm_log(swotlist, mc = mc)
  
  out <- map_dbl(lms, A0_R2)
  out
}
```

That was only a slight modification of the previous one. Again, applied to Pepsi:


```{r}
map(reachdata, possibly(~manning_R2_log(., mc = TRUE), otherwise = NULL)) %>% 
  map(~data.frame(reach = names(.), R2 = .)) %>% 
  bind_rows(.id = "case") %>% 
  ggplot(aes(x = case, y = 1 - R2)) +
  geom_point() +
  theme(axis.text = element_text(angle = 90))

map(reachdata, possibly(~manning_R2_log(., mc = FALSE), otherwise = NULL)) %>% 
  map(~data.frame(reach = names(.), R2 = .)) %>% 
  bind_rows(.id = "case") %>% 
  ggplot(aes(x = case, y = 1 - R2)) +
  geom_point() +
  theme(axis.text = element_text(angle = 90))
```

OK, that appears fine. Not sure when I'll use this. More useful probably to have some plots. 

- Partial residual plots
- Prediction vs truth plots

```{r}
manning_valplot <- function(swotlist, mc = TRUE, plot = TRUE, log = TRUE) {
  lms <- manning_lm_log(swotlist, mc = mc)
  names(lms) <- 1:length(lms)
  valdfs <- map(lms, ~data.frame(true = .$y, pred = predict(.)))
  valdf <- bind_rows(valdfs, .id = "loc")
  valdf$loc <- as.factor(as.integer(valdf$loc))
  
  if (!log) {
    valdf$true <- exp(valdf$true)
    valdf$pred <- exp(valdf$pred)
  }
  
  if (!plot) 
    return(valdf)
  
  out <- ggplot(valdf, aes(x = pred, y = true)) +
    geom_point(aes(color = loc)) +
    geom_abline(aes(slope = 1, intercept = 0))
  out
}

#' Generate termplots for Manning variables
#'
#' Modified from a similar function in 20180312 notebook.
manning_termplot <- function(swotlist, mc = TRUE, plot = TRUE, scales = "free") {
  W <- swotlist$W
  S <- swotlist$S
  A <- swotlist$A
  Q <- swotlist$Q
  
  if (mc) {
    Qvec <- apply(Q, 2, median, na.rm = TRUE)
    Q <- matrix(rep(Qvec, nrow(Q)), nrow = nrow(Q), byrow = TRUE)
  }
  
  fullresids <- log(Q) - manning_qdot(W, S, A, log = TRUE)
  
  wresids <- fullresids - 2/3 * log(W)
  sresids <- fullresids + 1/2 * log(S)
  aresids <- fullresids + 5/3 * log(A)
  residlist <- list(W = wresids, S = sresids, A = aresids)
  
  datadf <- swotlist[c("W", "S", "A")] %>% 
    map(log) %>% 
    swot_tidy() %>% 
    gather(key = "variable", value = "logval", -time, -loc)
  residdf <- swot_tidy(residlist) %>% 
    gather(key = "variable", value = "p_resid", -time, -loc) %>% 
    left_join(datadf, by = c("variable", "time", "loc")) %>% 
    group_by(variable) %>% 
    mutate(p_resid = p_resid - mean(p_resid, na.rm = TRUE)) %>% 
    ungroup()
  
  if (!plot)
    return(residdf)
  
  plotdf <- residdf %>% 
    group_by(variable) %>% 
    mutate(meanval = mean(logval, na.rm = TRUE),
           # meanresid = mean(p_resid, na.rm = TRUE),
           slope = ifelse(variable == "W", -2/3, 
                          ifelse(variable == "S", 1/2, 
                                 ifelse(variable == "A", 5/3, 
                                        stop("variable not recognized")))),
           intercept = - slope * meanval,
           loc = as.factor(as.integer(loc))) 
  
  out <- ggplot(plotdf, aes(x = logval, y = p_resid, color = loc)) +
    geom_point() + 
    geom_abline(aes(slope = slope, intercept = intercept)) +
    facet_wrap(~variable, scales = scales)
  
  out
}

manning_termplot(reachdata$Kanawha, mc = TRUE, plot = TRUE, scales = "free_x")
manning_valplot(reachdata$Kanawha, mc = TRUE, plot = TRUE, log = TRUE)
manning_valplot(reachdata$Kanawha, mc = TRUE, plot = TRUE, log = FALSE)

```

That looks good. Now put these into lib/

### ANOVA on manning's n

Next step is to package up ANOVA functions. 

- Already have a lib/anova.R script. Just need to add some stuff from the reports/n_anova.Rmd file. 
- Specifically, make a function that does the ggplot for the depth-dependent Manning's n. 

```{r}
manning_nplot <- function(swotlist, plot = TRUE) {
  swotdf <- swot_tidy(swotlist) %>% 
    mutate(logn = -2/3 * log(W) + 5/3 * log(A) + 1/2 * log(S) - log(Q), 
           logd = log(A / W),
           loc = as.factor(loc))
  
  if (!plot) 
    return(swotdf)
  
  outgg <- swotdf %>% 
    ggplot(aes(x = logd, y = logn)) +
    geom_point(aes(color = loc), size = 0.2) + 
    stat_smooth(aes(group = loc, color = loc), method = "lm", se = FALSE)
  outgg
}

manning_nplot(reachdata$Connecticut)
head(manning_nplot(reachdata$Connecticut, plot = FALSE))
```


Alright, that's all taken care of. On to the next. 

I should draft a document outlining the inference pipeline. Briefly, I think this would be:

- read in data
- calculate dA (if necessary)
- Lag times via ccf_lag
- Subset locations, e.g. using condition number or A0 R2
- Iterate:
    - estimate A0
    - Estimate ndot
    
Partway through this, as well as at the end, diagnostics would be in order, including:

- A0 inference GoF (R2)
- Condition number
- dq/dx (estimated)

If validation data are available, the following could also be looked at:

- Manning GoF
    - in log space
    - in linear A space
- ANOVA on Manning n
- dq/dx 
- Manning termplots (log space)
- Manning prediction validation plots (log, linear Q)


### Picking up the next day

I tried implementing joint estimation of ndot and A0, but it's blowing up. Testing now.

```{r}
swotlist$Ahat <- NULL
ndot <- 1
ndots <- list()
qdots <- list()

for (i in 1:iters) {
  # estimate A0
  A0hat <- coef(estA0_list(swotlist, ndot = ndot))
  # A0hat[A0hat < 0] <- 0
  swotlist$Ahat <- swotlist$dA + swot_vec2mat(A0hat, swotlist$dA)
  # estimate ndot
  ndot <- manning_ndot(swotlist, avar = "Ahat", log = FALSE)
  ndots[[i]] <- log(ndot)
  qdots[[i]] <- manning_qdot(swotlist$W, swotlist$S, swotlist$Ahat, ndot = ndot, log = FALSE)
  if (trace) {
    cat(sprintf("iter: \t%s \nA0: \t%s \nndot: \t%s \n\n", i, 
                paste(round(A0hat), collapse = "\t"), 
                paste(round(ndot, digits = 3), collapse = "\t")))
  }
}
```

Inspect. Is difference between locations on qdots changing?

```{r}

# checkdf <- map(ndots, ~swot_tidy(list(dotval = swot_vec2mat(., qdots[[1]])))) %>%
checkdf <- map(qdots, ~swot_tidy(list(dotval = .))) %>%
  bind_rows(.id = "iter")

checkdf %>%
  mutate(iter = as.numeric(iter)) %>% 
  # glimpse() %>% 
  group_by(iter, loc) %>% 
  summarize(meanqdot = mean(dotval, na.rm = TRUE), 
            q1 = quantile(dotval, 0.25, na.rm = TRUE),
            q3 = quantile(dotval, 0.75, na.rm = TRUE)) %>% 
  ungroup() %>% 
  gather(key = stat, value = value, -iter, -loc) %>% 
  glimpse() %>% 
  filter(stat == "meanqdot") %>%
  mutate(loc = as.factor(loc)) %>% 
  ggplot(aes(x = iter, y = value, color = loc, group = loc)) +
  geom_point()
                
```


I need to investigate this further. Make sure it's behaving as intended:

- initialize ndot to be 1. 
- estimate A0, conditional on ndot
- estimate ndot


What do I expect to happen as I iterate?

- qdots get closer together

```{r}
qdots %>% 
  map(~apply(., 1, function(x) mean(log(x), na.rm = TRUE))) %>% 
  as.data.frame() %>% 
  unname() %>% 
  as.matrix() %>% 
  t()
```

OK, so it's working as expected! It just blows up. I'll probably have to regularize if I want it to work at all. For now work on prep for skype with Mike.
