---
title: "notebook20180316"
author: "Mark Hagemann"
date: "March 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Some quick-and-dirty plots that will inform how I structure the shiny vis. 

```{r}
library(swotData)

pep1cases <- names(Pepsi_v2)
pep1ncs <- sprintf("../../swotData/data/NC_files/%s.nc", pep1cases)

nclists <- map(pep1ncs, nc_list) %>% 
  setNames(pep1cases)

nclists$Platte$River_Info.gdrch <- 1:12 # get rid of 13-14.

gdrchs <- map(nclists, ~.$River_Info.gdrch)

keepvars <- c("W", "S", "A", "Q")
reachparts <- paste0("Reach_Timeseries.", keepvars)

zeroToNA <- function(mat) {
  zeros <- mat <= 0
  mat[zeros] <- NA
  mat
}

reachdata0 <- nclists %>% 
  map(~.[reachparts]) %>% 
  map(setNames, keepvars) %>% 
  map2(gdrchs, function(x, y) map(x, function(z) z[y, ])) %>% 
  map(~map(., zeroToNA))

reachdata0$Severn$W <- reachdata0$Severn$W[, -1]
reachdata0$Severn$A <- reachdata0$Severn$A[, -1]
reachdata0$Severn$S <- reachdata0$Severn$S[, -1]
reachdata0$Severn$Q <- reachdata0$Severn$Q[, -1]

reachdata0$Platte$W <- reachdata0$Platte$W[, -1]
reachdata0$Platte$A <- reachdata0$Platte$A[, -1]
reachdata0$Platte$S <- reachdata0$Platte$S[, -1]
reachdata0$Platte$Q <- reachdata0$Platte$Q[, -1]

reachncols <- map(reachdata0, ~ncol(.$A))

reachA0 <- reachdata0 %>% 
  map(~.$A) %>% 
  map(apply, 1, median) %>% 
  map2(reachncols, function(x, y) matrix(rep(x, y), ncol = y, byrow = FALSE))

reachdA <- reachdata0 %>% 
  map(~.$A) %>% 
  map2(reachA0, function(x, y) x - y)

reachdata <- map2(reachdata0, reachdA, function(x, y) {x$dA <- y; x})


Qmats <- nclists %>% 
  map(~.$Reach_Timeseries.Q[.$River_Info.gdrch, ])
```

Calculate parameters.

```{r}
manNlist <- map(reachdata, manningN_list)
constNlist <- map(manNlist, median, na.rm = TRUE)

manQlist <- map2(reachdata, constNlist, function(x, y) manningQ_list(x, y))

trueQlist <- map(reachdata, `[[`, "Q")

manQdf <- manQlist %>% 
  map(~data.frame(manQ = as.vector(.), 
                  time = rep(1:ncol(.), each = nrow(.)),
                  loc = rep(1:nrow(.), ncol(.)))) %>% 
  bind_rows(.id = "case")

trueQdf <- trueQlist %>% 
  map(~data.frame(trueQ = as.vector(.), 
                  time = rep(1:ncol(.), each = nrow(.)),
                  loc = rep(1:nrow(.), ncol(.)))) %>% 
  bind_rows(.id = "case")

Qdf <- left_join(manQdf, trueQdf, by = c("time", "loc", "case"))

Qdf %>% 
  sample_n(10000) %>% 
  ggplot(aes(x = trueQ, y = manQ)) +
  geom_point(aes(color = case), alpha = 0.4) +
  scale_x_log10() + scale_y_log10() +
  theme_minimal() +
  geom_abline(slope = 1, intercept = 0)
```

In the interactive version I'll be able to filter this and identify points. 

Now work on getting something like termplots. 

```{r}
partManResid = function(Q, A, W, S, n, Aexp = 5/3, Wexp = -2/3, Sexp = 1/2) {
  Qpart_W <- Q / A^Aexp * S^Sexp / n
  Qpart_S <- Q / A^Aexp * W^Wexp / n
  Qpart_A <- Q / W^Wexp * S^Sexp / n
  
  out <- list(wpart = Qpart_W,
              spart = Qpart_S,
              apart = Qpart_A)
  
  out
}

partManResid_list <- function(datalist) {
  A <- datalist$A
  W <- datalist$W
  S <- datalist$S
  Q <- datalist$Q
  n <- datalist$n
  
  out <- partManResid(Q = Q, A = A, W = W, S = S, n = n)
  out
}

case1 <- "Wabash"
partRes1 <- partManResid_list(datalist = c(reachdata[[case1]],
                                           list(n = constNlist[[case1]])))

plot(log(reachdata[[case1]]$S), log(partRes1$spart))
abline(0, 1/2)

```


I've outlined the major parts of the visualization in my (physical) notebook. 

Remaining figs:

- stats boxplots across all cases 
    - RRMSE, NRMSE
- case-scale hydrograph
- reach-scale data tables
- reach-scale "empirical" exponents comparison to Manning exponents
    - along with condition number?

```{r}
evalstatdf <- Qdf %>% 
  glimpse() %>% 
  group_by(case, loc) %>% 
  summarize(RRMSE = RRMSE(pred = manQ, meas = trueQ),
            NRMSE = NRMSE(pred = manQ, meas = trueQ),
            rBIAS = rBIAS(pred = manQ, meas = trueQ),
            NSE = NSE(pred = manQ, meas = trueQ)) %>% 
  ungroup()
  
allstats_gg <- evalstatdf %>%
  gather(key = stat, value = value, -case, -loc) %>% 
  ggplot(aes(x = stat, y = value)) +
  geom_boxplot() + 
  theme_bw()

allstats_gg
ggsave("../graphs/nplots/allstats_boxplot1.png")


rrmse_gg <- evalstatdf %>% 
  filter(!is.na(RRMSE)) %>% 
  ggplot(aes(x = case, y = RRMSE)) + 
  # geom_boxplot() +
  coord_flip() +
  geom_point(size = 0.5)
rrmse_gg
ggsave("../graphs/nplots/rrmse_constant_n.png")

nrmse_gg <- evalstatdf %>% 
  filter(!is.na(NRMSE)) %>% 
  ggplot(aes(x = case, y = NRMSE)) + 
  # geom_boxplot() +
  coord_flip() +
  geom_point(size = 0.5) +
  theme_bw() +
  xlab("")
nrmse_gg
ggsave("../graphs/nplots/nrmse_constant_n.png")
```


Now case-scale hydrograph. (Should be easy)

```{r}
case2 <- "Platte"
Qdf %>% 
  filter(case == case2) %>% 
  glimpse() %>% 
  ggplot(aes(x = time, group = loc)) +
  geom_line(aes(y = manQ), linetype = 1) +
  geom_line(aes(y = trueQ), linetype = 2, color = "red") +
  theme_bw() +
  scale_y_log10()
```


Fine. Now I need to make variable-n predictions. 

- claim is that there is an exponential relationship between n and A/W


```{r}
manNlist <- map(reachdata, manningN_list)
awlist <- map(reachdata, function(x) x$A / x$W)

case3 <- "Severn"
plot(awlist[[case3]], manNlist[[case3]], log = "xy")

```

This does not convince me! Certainly the relationship does not hold across reaches. In any case, I can predict and plot. 


```{r}
n_aw_dfs = map2(manNlist, awlist, 
                function(x, y) data.frame(logn = log(as.vector(x)), 
                                          logd = log(as.vector(y))))

n_aw_lms = map(n_aw_dfs, possibly(~lm(logn ~ logd, data = .), otherwise = NA))

predictVarN = function(datalist, mod) {
  d <- datalist$A / datalist$W
  logd_mat <- log(d)
  logd_vec <- as.vector(logd_mat)
  npred_vec <- predict(mod, newdata = data.frame(logd = logd_vec))
  outmat <- matrix(exp(npred_vec), nrow = nrow(d))
  outmat
}

varNlist <- map2(reachdata, n_aw_lms, possibly(predictVarN, otherwise = NA))

case4 <- "Platte"
plot(manNlist[[case4]], varNlist[[case4]])
abline(0, 1)
```



```{r}
manQlist_vn <- map2(reachdata, varNlist, function(x, y) manningQ_list(x, y))

trueQlist <- map(reachdata, `[[`, "Q")

manQdf_vn <- manQlist_vn %>% 
  map(~data.frame(manQ = as.vector(.), 
                  time = rep(1:ncol(.), each = nrow(.)),
                  loc = rep(1:nrow(.), ncol(.)))) %>% 
  bind_rows(.id = "case")


Qdf_vn <- left_join(manQdf_vn, trueQdf, by = c("time", "loc", "case"))
```

Now look at stats


```{r}
evalstatdf_vn <- Qdf_vn %>% 
  glimpse() %>% 
  group_by(case, loc) %>% 
  summarize(RRMSE = RRMSE(pred = manQ, meas = trueQ),
            NRMSE = NRMSE(pred = manQ, meas = trueQ),
            rBIAS = rBIAS(pred = manQ, meas = trueQ),
            NSE = NSE(pred = manQ, meas = trueQ)) %>% 
  ungroup()
  
allstats_vn_gg <- evalstatdf_vn %>%
  gather(key = stat, value = value, -case, -loc) %>% 
  ggplot(aes(x = stat, y = value)) +
  geom_boxplot() + 
  theme_bw()

allstats_vn_gg
ggsave("../graphs/nplots/allstats_boxplot2.png")

rrmse_vn_gg <- evalstatdf_vn %>% 
  filter(!is.na(RRMSE)) %>% 
  ggplot(aes(x = case, y = RRMSE)) + 
  # geom_boxplot() +
  coord_flip() +
  geom_point(size = 0.5)
rrmse_vn_gg
ggsave("../graphs/nplots/rrmse_variable_n_1.png")

nrmse_vn_gg <- evalstatdf_vn %>% 
  filter(!is.na(NRMSE)) %>% 
  ggplot(aes(x = case, y = NRMSE)) + 
  # geom_boxplot() +
  coord_flip() +
  geom_point(size = 0.5) +
  theme_bw() +
  xlab("")
nrmse_vn_gg
ggsave("../graphs/nplots/nrmse_variable_n_1.png")
nrmse_gg
```


Combine together for boxplots.

```{r}
evalstatdf_g <- evalstatdf %>%
  gather(key = stat, value = value, -case, -loc)
evalstatdf_vn_g <- evalstatdf_vn %>%
  gather(key = stat, value = value, -case, -loc)



allstats_both_gg <- left_join(evalstatdf_g, evalstatdf_vn_g, 
                            by = c("case", "loc", "stat")) %>%
  rename(constant = value.x, variable = value.y) %>% 
  gather(key = n_treatment, value = value, -case, -loc, -stat) %>% 
  glimpse() %>% 
  ggplot(aes(x = stat, y = value)) +
  geom_boxplot(aes(fill = n_treatment)) + 
  theme_bw()
allstats_both_gg
```

What if I also vary in space (and let lm vary in space)?



```{r}

spaManN_vecs <- map(manNlist, ~split(., f = 1:nrow(.)))
spaAW_vecs <- map(awlist, ~split(., f = 1:nrow(.)))

make_spa_lms <- function(x, y) map2(x, y, function(w, z) lm(logn ~ logd, 
                                      data = data.frame(logn = log(w), 
                                                        logd = log(z))))

spaN_AW_lms <- map2(spaManN_vecs, spaAW_vecs, 
                    possibly(make_spa_lms, otherwise = NA))

predlmsnow <- function(lmlist, veclist) {
  map2(lmlist, veclist, function(x, y) predict(x, newdata = data.frame(logd = log(y))))
}

spaN_preds <- map2(spaN_AW_lms, spaAW_vecs, 
                   possibly(predlmsnow, otherwise = NA))

toMatrix <- function(lst) {
  t(as.matrix(as.data.frame(map(lst, unname))))
}

spaN_predmats <- map(spaN_preds, possibly(toMatrix, otherwise = NA))



varNlist2 <- map(spaN_predmats, possibly(exp, otherwise = NA))

case5 <- "Po"
plot(manNlist[[case5]], varNlist2[[case5]], xlab = "true Manning n", ylab = "predicted Manning n")
abline(0, 1)
```

Now I can do the ggplots again. 

```{r}
manQlist_vn2 <- map2(reachdata, varNlist2, possibly(manningQ_list, otherwise = NA))

trueQlist <- map(reachdata, `[[`, "Q")

manQdf_vn2 <- manQlist_vn2[!is.na(manQlist_vn2)] %>% 
  map(~data.frame(manQ = as.vector(.), 
                  time = rep(1:ncol(.), each = nrow(.)),
                  loc = rep(1:nrow(.), ncol(.)))) %>% 
  bind_rows(.id = "case")


Qdf_vn2 <- left_join(manQdf_vn2, trueQdf, by = c("time", "loc", "case"))
```


```{r}
evalstatdf_vn2 <- Qdf_vn2 %>% 
  glimpse() %>% 
  group_by(case, loc) %>% 
  summarize(RRMSE = RRMSE(pred = manQ, meas = trueQ),
            NRMSE = NRMSE(pred = manQ, meas = trueQ),
            rBIAS = rBIAS(pred = manQ, meas = trueQ),
            NSE = NSE(pred = manQ, meas = trueQ)) %>% 
  ungroup()

rrmse_vn2_gg <- evalstatdf_vn2 %>% 
  filter(!is.na(RRMSE)) %>% 
  ggplot(aes(x = case, y = RRMSE)) + 
  # geom_boxplot() +
  coord_flip() +
  geom_point(size = 0.5)
rrmse_vn2_gg
ggsave("../graphs/nplots/rrmse_variable_n_2.png")


nrmse_vn2_gg <- evalstatdf_vn2 %>% 
  filter(!is.na(NRMSE)) %>% 
  ggplot(aes(x = case, y = NRMSE)) + 
  # geom_boxplot() +
  coord_flip() +
  geom_point(size = 0.5) +
  theme_bw() +
  xlab("")
nrmse_vn2_gg
ggsave("../graphs/nplots/nrmse_variable_n_2.png")
```

Finally, the boxplots.

```{r}
evalstatdf_g <- evalstatdf %>%
  gather(key = stat, value = value, -case, -loc)
evalstatdf_vn_g <- evalstatdf_vn %>%
  gather(key = stat, value = value, -case, -loc)
evalstatdf_vn2_g <- evalstatdf_vn2 %>%
  gather(key = stat, value = value, -case, -loc)


allstats_both_gg <- left_join(evalstatdf_g, evalstatdf_vn_g, 
                            by = c("case", "loc", "stat")) %>%
  rename(constant = value.x, variable1 = value.y) %>% 
  left_join(evalstatdf_vn2_g, by = c("case", "loc", "stat")) %>% 
  rename(variable2 = value) %>% 
  gather(key = n_treatment, value = value, -case, -loc, -stat) %>% 
  glimpse() %>% 
  ggplot(aes(x = stat, y = value)) +
  geom_boxplot(aes(fill = n_treatment)) + 
  theme_bw()
allstats_both_gg
ggsave("../graphs/nplots/allstats_boxplot3.png")
```

### ANOVA on logn--linear models

```{r}
anovalm1 <- reachdata$Ganges %>% 
  swot_tidy() %>% 
  glimpse() %>% 
  mutate(D = A / W, N = W^(-2/3) * A^(5/3) * S^(1/2) / Q, 
         locfac = as.factor(loc), logn = log(N), logd = log(D)) %>% 
  glimpse() %>% 
  lm(data = ., formula = logn ~ locfac * logd)

summary(anovalm1)

foo <- tidy(anova(anovalm1))
```

Whip up a function. 

```{r}
swot_n_anova_lm <- function(swotlist) {
  anovalm <- swotlist %>% 
  swot_tidy() %>% 
  mutate(D = A / W, N = W^(-2/3) * A^(5/3) * S^(1/2) / Q, 
         locfac = as.factor(loc), logn = log(N), logd = log(D)) %>% 
  lm(data = ., formula = logn ~ locfac * logd)
  anovalm
}

swot_n_anova <- function(swotlist) {
  out <- broom::tidy(anova(swot_n_anova_lm(swotlist)))
  out
}
```

Now do them all and look at meansq distributions.

```{r}

anovadf <- map(reachdata, possibly(swot_n_anova, otherwise = NA)) %>% 
  (function(x) x[!is.na(x)]) %>% 
  bind_rows(.id = "case") %>% 
  group_by(case) %>% 
  # mutate(pctTotVar = meansq / sum(meansq) * 100) %>% # Old, wrong version
  mutate(pctTotVar = sumsq / sum(sumsq) * 100) %>% # Correct version
  # mutate(pctTotVar = sumsq / length(sumsq)) %>%  # scaled to compare across locations
  ungroup() %>% 
  mutate(term = as.character(term),
         term  = plyr::mapvalues(term, 
                                from = c("locfac", "locfac:logd", "logd"),
                                to = c("location", "location:depth", "depth")))

anova_gg <- anovadf %>% 
  filter(case != "Tanana") %>% 
  ggplot(aes(x = term, color = case, y = pctTotVar)) +
  geom_point() + 
  geom_line(aes(group = case)) +
  # scale_y_log10() +
  theme_bw() +
  ylab("Percent of total variance explained")

ggplotly(anova_gg)
# ggsave("../graphs/nplots/n_anova.png")

```


There appears to be an outlier in Platte location 12. 

```{r}
platte_tidy <- swot_tidy(reachdata$Platte)
  
platte_tidy %>% 
  glimpse() %>% 
  filter(loc == 12) %>%
  mutate(Qhat = W^(-2/3) * A^(5/3) * S^(1/2), 
         logn = log(Qhat / Q), 
         logd = log(A / W)) %>% 
  # filter(logd == min(logd))
  ggplot(aes(x = Qhat, y = Q)) + 
  geom_point(aes(color = time == 1)) +
  scale_y_log10() +
  scale_x_log10()
```

OK, that's fixed. 

Also, I can't wrap my head around the ANOVA for MS DS. It says the MSE for residuals is near zero, yet the models doen't look all that good. 

```{r}
foo <- swot_n_anova_lm(swotlist = reachdata$MississippiDownstream)
summary(foo)

anova(foo)

markstats::MSE(foo)
markstats::R2(foo)
var(foo$model[, 1])



```

Right, I should be lookin at sum of squares, not mean square. 

- sum is conserved, mean is not. 

```{r}
foo <- swot_tidy(reachdata$Tanana)
glimpse(foo)
summary(foo)
```

