---
title: "notebook20180413"
author: "Mark Hagemann"
date: "April 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Manning flow imbalance (gamma)



```{r}
# casei <- uscase
# casei <- uscase %>%
#   swot_timelag(c(0, -2, -4))
casei <- reachdata$Po

gammai <- swot_gamma(casei)

plot_DAWG(gammai)
```

Now decompose gamma into constant (space-invariant) dg/dx, other components. 

$$
\gamma_{it} = \tau_{it}\bar{\gamma}'_{\cdot t}(x_i - \bar{x})
$$

```{r}
dg <- apply(gammai, 2, diff)
dx <- apply(casei$x, 2, diff)

dgdxbar0 <- (dg / dx) %>% 
  apply(2, mean) %>% 
  swot_vec2mat(dg)

dgdxbar <- dgdxbar0 %>% 
  (function(x) (x[c(1, 1:nrow(x)), ] + x[c(1:nrow(x), nrow(x)), ]) / 2) # give the right dimensions

plot_DAWG(dg/dx)
plot_DAWG(dg/dx - dgdxbar0)

var(as.vector(dg/dx - dgdxbar0)) / var(as.vector(dg/dx0)) # R2 of ~0.86
```

But how well does that explain gamma?

```{r}
deltaxi <- apply(casei$x, 2, function(x) x - mean(x))
gammahat <- deltaxi * dgdxbar + 1

plot_DAWG(gammahat)
plot(gammahat, gammai); abline(0, 1)

var(as.vector(gammai - gammahat)) / var(gammai %>% as.vector)

multerr <- gammai / gammahat

```

98% variance explained!

Now can I use this to decompose and modify the closure term? First get the notation right. 

$$
\gamma_{it} = (\bar{\gamma}'_{\cdot t}(x_i - \bar{x}) + 1) * \epsilon_{it}
$$

Here the error term has *geometric mean of 1*. Still doesn't fix awkwardness of the model. 


```{r}
decomp_gamma <- function(gammamat, xmat, additive = FALSE) {
  dg <- apply(gammamat, 2, diff)
  dx <- apply(xmat, 2, diff)
  
  dgdxbar0 <- (dg / dx) %>% 
    apply(2, mean) %>% 
    swot_vec2mat(dg)
  
  dgdxbar <- dgdxbar0 %>% 
    (function(x) (x[c(1, 1:nrow(x)), ] + x[c(1:nrow(x), nrow(x)), ]) / 2) # give the right dimensions
  
  deltax <- apply(xmat, 2, function(x) x - mean(x))
  gammahat <- deltax * dgdxbar + 1

  if (additive) {
    gammaerr <- gammamat - gammahat
  } else {
    gammaerr <- gammamat / gammahat
  }
  
  out <- list(gammahat = gammahat, gammaerr = gammaerr, dgdxbar = dgdxbar[1, ])
  out
}

foo <- decomp_gamma(swot_gamma(uscase), uscase$x)

plot_DAWG(foo$gammahat)
plot_DAWG(foo$gammaerr)
plot.ts(foo$dgdxbar)
```


```{r}

# casej <- uscase %>% 
#   swot_timelag(c(0, 2, 4))
casej <- reachdata$Po
# casej <- reachdata$Platte %>% 
#   swot_sset()




gammaj <- swot_gamma(casej)
foo <- decomp_gamma(gammaj, casej$x)

plot_DAWG(gammaj)
plot_DAWG(foo$gammahat)
plot_DAWG(foo$gammaerr)
plot.ts(foo$dgdxbar)
```


Now a function to modify flow imbalance. 

```{r}

manning_rebalance <- function(swotlist, scale_dqdx, additive = FALSE) {
  gammamat <- swot_gamma(swotlist)
  gammadecomp <- decomp_gamma(gammamat, swotlist$x, additive = additive)
  
  modghat <- gammadecomp$gammahat * scale_dqdx
  
  if (additive) {
    out <- modghat + gammadecomp$gammaerr
  } else {
    out <- modghat * gammadecomp$gammaerr
  }
} 

foo <- manning_rebalance(casej, 1)

bar <- swot_gamma(casej)

plot(foo, bar); abline(0, 1)
```

Good. Now how to use this in A0 inference? Need to look back at past weekend's work, how did I apply errors?

- Need to make error matrices, add to manning_linA_closed() 
- This requires calculating other part of closure term. But I can calculate closure term already. 

```{r}
casej <- uscase %>% #`(`
  swot_timelag(c(0, -2, -4))
foo <- manning_closure(casej)
bar <- swot_gamma(casej)

plot(bar, foo); abline(0, 1)
```

Right, manning_closure has arbitrary mean that is related to n. Order of operations:

- Decompose closure into space-invariant dq/dx component and everything else. 
- Look at relative magnitudes of these
- Make new closure with new dq/dx via manning_rebalance.

```{r}
foo <- manning_closure(uscase)
gamj <- swot_gamma(uscase)
gamdecj <- decomp_gamma(gamj, uscase$x)

plot_DAWG(foo)
plot_DAWG(foo / gamj)
plot_DAWG(foo / gamdecj$gammahat)

plot_DAWG(gamdecj$gammahat)

```

Maybe stop at this characterization. I don't think I'll have time for simulation. Try for 5 minutes to make a modification function.

```{r}
modify_closure <- function(swotlist, scale_dqdx) {
  clos1 <- manning_closure(swotlist)
  
  gam1 <- swot_gamma(swotlist)
  
  gammod <- manning_rebalance(swotlist, scale_dqdx)
  out <- clos1 / gam1 * gammod
  out
}


casek <- sscase
plot_DAWG(manning_closure(casek))
plot_DAWG(modify_closure(casek, 0.1))
plot_DAWG(modify_closure(casek, 5))
```

I think I need to do this in an additive framework for it to matter to A0 inversion. 

For now, come up with the figs I want to show Mike. 

- Show that closed forms of cases work perfectly. 
- Show that I can calculate closure term
- Show decompositions of closure term
- Do the above for:
    - 3 lisflood cases
    - Po
    - Platte
    - Severn
- Show impact of clousure magnitude on inference 


### Reading at VCU

It just occurred to me that I should characterize relationship between predictor and response errors in A0 regressions. 

```{r}
# casei <- sscase
casei <- uscase
xerr <- manning_linA_closure(casei)
xclosed <- manning_linA_closed(casei)
xopen <- manning_ws35(casei)

lhserr <- omegaProduct(xopen) - omegaProduct(xclosed)
pairs(lhserr)
rhserr <- omegaProduct_y(xopen, casei$dA) - omegaProduct_y(xclosed, casei$dA)

pairs(cbind(rhserr, lhserr))
```


For differential error, we have the following from Table 3.1 in Carroll et al.:

$$
\beta = \beta_x \Big(\frac{\sigma_{xw}}{\sigma_{w}^2}\Big) + \frac{\sigma_{\epsilon w}}{\sigma_{w}^2}
$$

Can I characterize these parts for the cases I have?

- covariance of true and observed against variance of observed

```{r}
lhsopen <- omegaProduct(xopen) %>% 
  as.data.frame()
lhsclosed <- omegaProduct(xclosed) %>% 
  as.data.frame()
term1 <- map2_dbl(lhsopen, lhsclosed, cov) / map_dbl(lhsopen, var)

term2 <- map_dbl(lhsopen, ~cov(., rhserr)) / map_dbl(lhsopen, var)

ests1 <- estA0(casei, intercept = FALSE)
ests1_adj <- (ests1 - term2) / term1

ests1
ests1_adj
```

Not so simple, explore more later. Time to switch gears

### Project documentation

- Update readme
- Explore using pkgdown
- Document functions
- Write tests



