---
title: "lisflood-latinflows"
author: "Mark Hagemann"
date: "July 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(swotr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
opts_chunk$set(echo = TRUE)

theme_set(theme_bw())
```



```{r}
source("../../lib/lisflood.R")
dirs <- paste0(list.dirs("../../../lisflood/toy_1/lat-inflows", 
                         recursive = FALSE), "/results")
dirnames <- list.dirs("../../../lisflood/toy_1/lat-inflows", 
                         recursive = FALSE, full.names = FALSE)

reslist <- map(dirs, ~lis_reaches(lis_profiles(.)))

resdf <- reslist %>% 
  setNames(dirnames) %>% 
  bind_rows(.id = "run") %>% 
  filter(loc != 4)

resdf %>% 
  ggplot(aes(x = time, y = Q, color = loc, group = loc)) +
  geom_line() + 
  facet_wrap(~run, scales = "free_y")
```

Here's a brief summary of some recent work exploring the impact of lateral inflows on MetroMan inversions. For this we used the unit-hydrograph lisflood case described in a [previous post](https://swotdawg.wordpress.com/2018/05/21/variable-slope-simple-mcfli/), modified to add a constant Q between reaches 2 and 3. The hydrograph, width, slope, and height time series for the unmodified (no lateral inflow) case are as follows:

```{r}
resdf %>% 
  filter(run == "lat-000", time > 40) %>% 
  select(-run) %>% 
  swot_untidy() %>% 
  swot_plot(vars = c("Q", "W", "S", "H"))

ggsave("fig/orig_ts.png")

```

Initially we generated 7 cases with lateral inflow of 0.05, 0.1, 0.25, 0.5, 1, 2.5, and 5 cms, corresponding to between 0.1% and 10% of the original 50 cms baseflow. Height and discharge time series for reach 3 (the only reach affected by lateral inflows) are as follows:

```{r}
oldnames <- c("lat-000", "lat-005", "lat-010", "lat-025", "lat-050", "lat-100", 
"lat-1000", "lat-10000", "lat-2000", "lat-250", "lat-500", "lat-5000")
newnames <- gsub("lat-", "", oldnames, fixed = TRUE) 

resdf %>% 
  filter(run %in% c("lat-005", "lat-010", "lat-025", 
                    "lat-050", "lat-100", "lat-250", "lat-500"),
         time > 40, loc == 3) %>% 
  mutate(Q_in = as.numeric(plyr::mapvalues(run, from = oldnames, 
                                           to = newnames)) / 100,
         Q_in = as.factor(Q_in)) %>% 
  select(time, Q_in, H, Q) %>% 
  gather(key = variable, value = value, -time, -Q_in) %>%  glimpse() %>% 
  ggplot(aes(x = time, y = value, color = Q_in, group = Q_in)) + 
  geom_line() + 
  facet_wrap(~variable, scales = "free_y")
  
ggsave("fig/cases1_hq.png")
```

MetroMan performed remarkably well on these cases, with the worst case TODO

Next we generated 4 more extreme cases with lateral inflows of 10, 20, 50, and 100 cms--that's between 20% and 200% of original baseflow. Here are the hydrographs and height time series from reach 3 for these scenarios.

```{r}
resdf %>% 
  filter(run %in% c("lat-1000", "lat-2000", "lat-5000", "lat-10000"),
         time > 40, loc == 3) %>% 
  mutate(Q_in = as.numeric(plyr::mapvalues(run, from = oldnames, 
                                           to = newnames)) / 100,
         Q_in = as.factor(Q_in)) %>% 
  select(time, Q_in, H, Q) %>% 
  gather(key = variable, value = value, -time, -Q_in) %>%  glimpse() %>% 
  ggplot(aes(x = time, y = value, color = Q_in, group = Q_in)) + 
  geom_line() + 
  facet_wrap(~variable, scales = "free_y")

ggsave("fig/cases2_hq.png")
```

For context, here is the full dataset (all 3 reaches) for the largest of these inflows.

```{r}
resdf %>% 
  filter(run == "lat-10000", time > 40) %>% 
  select(-run) %>% 
  swot_untidy() %>% 
  swot_plot(vars = c("Q", "W", "S", "H"))

ggsave("fig/biggest_ts.png")
```

Here are some profiles of what height is doing downstream.

```{r}
dirs <- paste0(list.dirs("../../../lisflood/toy_1/lat-inflows", 
                         recursive = FALSE), "/results")
dirnames <- list.dirs("../../../lisflood/toy_1/lat-inflows", 
                         recursive = FALSE, full.names = FALSE)

profdf <- map(dirs, ~lis_profiles(.)) %>% 
  setNames(dirnames) %>% 
  bind_rows(.id = "run") %>% 
  mutate(Q_in = plyr::mapvalues(run, oldnames, as.numeric(newnames) / 100),
         Q_in = factor(Q_in, levels = sort(as.numeric(newnames) / 100)))

profdf %>% 
  filter(profile %in% c("0040", "0060", "0077"), 
         ChanX <= 30000) %>% 
  mutate(time = plyr::mapvalues(profile, 
                                from = c("0040", "0060", "0077"),
                                to = c("steady", "rising", "falling")),
         time = factor(time, levels = c("steady", "rising", "falling"))) %>% 
  select(ChanX, Flow, Q_in, time) %>% 
  ggplot(aes(x = ChanX, y = Flow)) + 
  geom_line(aes(color = Q_in)) + 
  facet_wrap(~time, nrow = 3, scales = "free_y") +
  xlab("Downstream distance (m)")

ggsave("fig/q_x.png", width = 5, height = 6)  
```

