---
title: "BAM Pepsi 2"
author: "Mark Hagemann"
date: "June 28, 2018"
output: 
  ioslides_presentation:
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(bamr)
library(rstan)
library(knitr)
library(ggplot2)
library(dplyr)
library(purrr)
library(tidyr)

source("../../lib/swotr.R")
source("../../lib/estA0.R")
source("../../lib/mcman.R")

theme_set(theme_bw(base_size = 20, base_line_size = 2))
```

```{r, cache=TRUE}
load("../../data/pep2objs.RData")

names(bamsamps) <- names(pep2_qobs)
names(bamsamps_man_amhg) <- names(pep2_qobs)

pepsi2cases <- names(pep2_qobs) %>% 
  map(~sprintf("../../data/Pepsi2/%s.nc", .)) %>% 
  map(~nc_reach(., good_only = TRUE)) %>% 
  setNames(names(pep2_qobs))

vals_man <- map2(bamsamps, pep2_qobs, ~bam_validate(.x, .y)) %>% 
  setNames(names(pep2_qobs))
vals_man_amhg <- map2(bamsamps_man_amhg, pep2_qobs, ~bam_validate(.x, .y)) %>% 
  setNames(names(pep2_qobs))

predq_man <- vals_man %>% 
  map(~.$valdata) %>% 
  bind_rows(.id = "case")

predq_man_amhg <- vals_man_amhg %>% 
  map(~.$valdata) %>% 
  bind_rows(.id = "case")

statdf_man <- vals_man %>% 
  map(~as.data.frame(as.list(.$stats))) %>% 
  bind_rows(.id = "case")

statdf_man_amhg <- vals_man_amhg %>% 
  map(~as.data.frame(as.list(.$stats))) %>% 
  bind_rows(.id = "case")

estimate_logA0 <- function(Wobs) {
  lwbar <- apply(log(Wobs), 1, mean)
  lwsd <- apply(log(Wobs), 1, sd)
  
  logA0hat <- -1.782 + 1.438 * lwbar - 2.268 * lwsd

  logA0hat
}


postA0 <- map(bamsamps, ~get_posterior_mean(., pars = "A0")[, 4])
linA0 <- map(pepsi2cases, ~estA0(., zero= "minimum"))
pep2priorA0 <- map(pepsi2cases, ~exp(estimate_logA0(.$W)))

```


## Algorithm {.smaller}

- Bayesian algorithm
    - likelihood based on Manning's equation, AMHG
    - Priors from literature, a priori datasets

Mass-conserved Manning:

$$
\frac{5}{3} \log (A_{0,r} + \delta A_{rt}) - \frac{2}{3} \log W_{rt} + \frac{1}{2} \log S_{rt} =   \log n + \log Q_t + \epsilon_{rt}
$$

## Algorithm

Mass-conserved AMHG:

$$
\log W_{rt} = b_r \Big(\log Q_{t} - \log Q_c \Big) + \log W_c + \epsilon_{rt} 
$$


## Bayesian philosophy

- Before observing anything, we are in a state of incomplete ignorance
- Data are generated by a process governed by unobserved parameters
- The model conveys information from the data to the parameters
- Posterior distribution fully quantifies uncertainty structure

$$
\begin{aligned}
p(Q, A_0, n | S, W, \delta A) & \propto p(S, W, \delta A | Q, A_0, n) \\
& \times p(Q) \times p(A_0) \times p(n)
\end{aligned}
$$

## Bayesian inference

$$
\begin{aligned}
p(Q, A_0, n | S, W, \delta A) & \propto p(S, W, \delta A | Q, A_0, n) \\
& \times p(Q) \times p(A_0) \times p(n)
\end{aligned}
$$

- Priors + Data + likelihood (model) --> posterior
- Inference via Hamiltonian Monte Carlo, Stan language
- More info: Hagemann et al., 2017

## Bird's-eye view

Manning-only BAM

```{r}
ggplot(predq_man, aes(x = qpred, y = qobs)) + 
  geom_point(aes(color = case)) +
  scale_x_log10() + scale_y_log10() +
  theme(legend.position = "none") +
  geom_abline(intercept = 0, slope = 1)
```


## Bird's-eye view

Manning + AMHG BAM

```{r}

ggplot(predq_man_amhg, aes(x = qpred, y = qobs)) + 
  geom_point(aes(color = case)) +
  scale_x_log10() + scale_y_log10() +
  theme(legend.position = "none") +
  geom_abline(intercept = 0, slope = 1)

```


## Performance stats

```{r}
list(manning = statdf_man, man_amhg = statdf_man_amhg) %>% 
  bind_rows(.id = "method") %>% 
  select(RRMSE, NRMSE, NSE, rBIAS, method) %>% 
  gather(key = "Stat", value = "value", -method) %>% 
  mutate(Stat = factor(Stat, levels = c("RRMSE", "NRMSE", "rBIAS", "NSE"))) %>% 
  ggplot(aes(x = Stat, y = value, fill = method)) + 
  geom_boxplot() + 
  geom_point(position = position_jitterdodge()) +
  coord_cartesian(ylim = c(-2, 2)) +
  ylab("") + xlab("") +
  theme(legend.position = "bottom")
```


## Bias against prior bias

```{r}
qwbms <- map_dbl(pepsi2cases, ~attr(., "QWBM"))

rbiaswbm <- pep2_qobs %>% 
  map(~mean(.)) %>% 
  map2_dbl(qwbms, function(x, y) (y - x) / x)


plot(rbiaswbm, statdf_man$rBIAS, xlab = "Prior bias", 
     ylab = "Posterior bias"); abline(0, 1)

# plot(rbiaswbm, statdf_man_amhg$rBIAS, xlab = "Prior bias", 
#      ylab = "Posterior bias"); abline(0, 1)
```

```{r, eval = FALSE}
sort(rbiaswbm, decreasing = TRUE)

bam_hydrograph(setNames(bamsamps, names(pep2_qobs))$MiddleRiver, pep2_qobs$MiddleRiver)

stan_trace(setNames(bamsamps, names(pep2_qobs))$MiddleRiver, pars = c("logn", "A0"))
realA0(pepsi2cases$MiddleRiver, rezero = "minimum")
estA0(pepsi2cases$MiddleRiver, zero = "minimum")
batman_linA(pepsi2cases$MiddleRiver)
manning_closure(pepsi2cases$MiddleRiver) %>% plot_DAWG()
```

## Prior mean on A0

```{r}

# plot(unlist(linA0), unlist(map(pepsi2cases, ~realA0(., rezero = "minimum"))), log = "xy"); abline(0, 1)

plot(unlist(pep2priorA0), unlist(map(pepsi2cases, ~realA0(., rezero = "minimum"))), 
     log = "xy", xlim = c(1, 10000), ylim = c(1, 10000),
     ylab = "True A0", xlab ="Prior A0"); abline(0, 1)
```

## Posterior mean on A0

```{r}
plot(unlist(postA0), unlist(map(pepsi2cases, ~realA0(., rezero = "minimum"))), 
     log = "xy", xlim = c(1, 10000), ylim = c(1, 10000),
     ylab = "True A0", xlab = "Posterior A0"); abline(0, 1)

```

## Remaining issues

- Uncertainty (SD) provided as known
- Priors are not all correct
    - A0 too tight
    - n too loose
    - QWBM too influential
- Errors treated as white noise!


