---
title: "A0 comparison"
author: "Mark Hagemann"
date: "July 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I now have 4 ways of estimating A0. Two are Bayesian algorithms (bmar and gibbs), and two are optimization (linear model and nlm). This post will compare them using the Pepsi 1 datasets and lisflood cases. 


### datasets

```{r}
testcases <- c(reachdata[1:15], list(uscase = uscase, sscase = sscase)) %>% 
  map(~within(., {dA = rezero_dA(dA, zero = "median")}))
```


### bamr inversions

I am using the median_A0 version of bamr (still needs version bump!). 

```{r}
library(bamr)
library(rstan)
bamests <- testcases %>% 
  map(~swot_bamdata(.)) %>% 
  map(possibly(~bam_estimate(., bampriors = bam_priors(., logA0_sd = 0.5, logn_sd = 0.25),
                             variant = "manning"), otherwise = NA), iter = 2000)


```

```{r}
relsd <- function(...) {
  mat <- Reduce(cbind, list(...))
  out <- apply(mat, 1, function(x) sd(x / x[1]))
  out
}
bam_A0 <- bamests %>% 
  map(~get_posterior_mean(., pars = "A0")) %>% 
  map(as.data.frame) %>% 
  bind_rows(.id = "case") %>% 
  setNames(c("case", "chain1", "chain2", "chain3", "allchains")) %>% 
  mutate(converged = relsd(chain1, chain2, chain3) < 0.05) %>% 
  transmute(case, A0 = allchains, converged)
```

### gibbs inversions

```{r}

gsprint <- function(inputs, name) {
  print(name)
  gibbs_sample(inputs, chains = 3, iter = 4000)
}
gibbsests <- testcases %>% 
  map(~gibbs_inputs(.)) %>% 
  map2(names(.), possibly(~gsprint(.x, .y), otherwise = NA))
```

Some of them did not work.

```{r}
gibbsests$Cumberland
inps1 <- gibbs_inputs(testcases$Cumberland)
init1 <- gibbs_init_rand(inps1)
foo <- gibbs_sample(inps1, chains = 3, iter = 1000)
```


```{r}
gibbsests_sf <- map(gibbsests, gibbs_stanfit)
gibbs_A0 <- gibbsests_sf %>% 
  map(~get_posterior_mean(., pars = "A0")) %>% 
  map(as.data.frame) %>% 
  bind_rows(.id = "case") %>% 
  setNames(c("case", "chain1", "chain2", "chain3", "allchains")) %>% 
  mutate(converged = relsd(chain1, chain2, chain3) < 0.05) %>% 
  transmute(case, A0 = allchains, converged)

gibbsests_sf$Connecticut %>% stan_trace(pars = "A0")
gibbsests_sf$GaronneUpstream %>% stan_trace(pars = "A0", inc_warmup = TRUE)
gibbsests_sf$MississippiUpstream %>% stan_trace(pars = "A0") # real bad
gibbsests_sf$Seine %>% stan_trace(pars = "A0")
```


### Linear model

```{r}
lmests <- testcases %>% 
  map(~swot_purge_nas(.)) %>% 
  map(~estA0(.))

lm_A0 <- lmests %>% 
  map(~data.frame(A0 = .)) %>% 
  bind_rows(.id = "case") %>% 
  transmute(case, A0, converged = TRUE)
```


### Nonlinear model

```{r}
nlmests <- testcases %>% 
  map(~swot_purge_nas(.)) %>% 
  map(~batman_log(., iterlim = 1000))

nlm_A0 <- nlmests %>% 
  map(~data.frame(A0 = .$A0, code = .$code)) %>% 
  bind_rows(.id = "case") %>% 
  transmute(case, A0, converged = code < 4)
```

```{r}
nlm_A0 %>% 
  ggplot(aes(x = case, y = A0)) + 
  geom_point() + 
  scale_y_log10()
```


## Comparison

```{r}
A0df <- list(bam = bam_A0, gibbs = gibbs_A0, lm = lm_A0, nlm = nlm_A0) %>% 
  bind_rows(.id = "method") %>% 
  group_by(case, method) %>% 
  mutate(reach = 1:n()) %>% 
  ungroup()

realA0s <- map(testcases, ~data.frame(A0_med = realA0(., zero = "median"), 
                                      A0_min = realA0(., zero = "minimum"))) %>% 
  bind_rows(.id = "case") %>% 
  group_by(case) %>% 
  mutate(reach = 1:n(), 
         dA_shift = A0_med - A0_min)

valdf <- left_join(A0df, realA0s, by = c("case", "reach")) %>% 
  mutate(A0 = ifelse(method == "bam", A0_med - A0_min + A0, A0))

valdf %>% 
  group_by(method) %>% 
  summary()

valdf %>% 
  ggplot(aes(x = A0, y = A0_med)) + 
  geom_point(aes(color = case)) + 
  facet_wrap(~method) + 
  scale_x_log10() + scale_y_log10() + 
  geom_abline(slope = 1, intercept = 0)
  
```

