---
title: "A0 inference 20180403"
author: "Mark Hagemann"
date: "April 3, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
opts_knit$set(root.dir = "~/Documents/mcfli-swotr")
opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
library(ProjectTemplate)
load.project()
ggplot2::theme_set(theme_bw())
```


## Overview

- Lisflood model runs
    - Steady-state
    - Nonsteady flood wave
    
## Steady-state model

```{r}
is_ss <- function(...) {
  arglist <- list(...)
  ssvars <- map(arglist, function(x) c(1, diff(x))) %>% 
    map(function(x) x == 0) %>% 
    Reduce(`*`, x = .)
  out <- as.logical(ssvars)
  out
}

sscase <- lis_profiles("lisflood/toy_1/results_const_v2/") %>% 
  lis_reaches(slope_method = "s_median", agfun = median) %>% 
  filter(loc != 4) %>%
  arrange(loc, time) %>% 
  group_by(loc) %>% 
  mutate(steady = is_ss(W, S, Q, H, D)) %>% 
  group_by(time) %>% 
  filter(sum(steady) == 3) %>% 
  group_by(Q) %>% 
  filter(time == time[1]) %>% 
  mutate(A = W * D) %>% 
  select(-steady) %>% 
  swot_untidy()

swot_plot(sscase)
```

## Steady state model: inference

It appears to work perfectly!

```{r, echo = TRUE}
estA0(sscase)

apply(sscase$A, 1, min)
```

## Steady state model: Diagnostics

Mass-conserved Manning partial residuals in log space

```{r}
val_log_termplot(sscase)
```

## Steady state model: Diagnostics

Linear-A-space mass-conserved plot

```{r}
val_linA_plot(sscase) +
  stat_smooth(method = "lm", alpha = 0.2, se = FALSE)
```

## Steady state model: Diagnostics

Manning n vs depth


```{r}
val_nd_plot(sscase)
```

## How variable are steady-state estimates? 

Bootstrap resampling of model times (n = 1000)

```{r}

casei <- sscase %>% 
  swot_sset()
  # swot_sset(keeplocs = c(1, 2)) %>% 
  # swot_timelag(c(0, -2))

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE))
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = .))

boot_nunique <- map_int(sampcollist, ~length(unique(.)))
boot_lms <- map(boot_cases, estA0_lm)
boot_A0 <- map(boot_lms, coef)
boot_R2 <- map_dbl(boot_lms, A0_R2)
boot_condno <- map_dbl(boot_lms, A0_condno)

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  map2(boot_R2, ~mutate(.x, R2 = .y)) %>% 
  map2(boot_condno, ~mutate(.x, condno = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = R2)) +
  geom_point(position = position_jitterdodge()) +
  # geom_boxplot(fill = NA) +
  scale_colour_viridis_c() 

```

## Resulting variability of discharge estimates (given known n = 0.04)

```{r}
boot_Ahats <- map(boot_A0, ~swot_A(A0vec = ., dAmat = casei$dA))

boot_Qts <- map(boot_Ahats, ~manningQ(A = ., W = casei$W, S = casei$S, n = 0.04)) %>% 
  map(~apply(., 2, median)) %>% 
  map(~data.frame(Q = ., time = 1:length(.))) %>% 
  bind_rows(.id = "sampno") 

boot_Qts %>% 
  ggplot(aes(x = time, y = Q, group = sampno)) + 
  geom_line(size = 0.3, alpha = 0.04)

```

## Other observations

- Sensitive to reach subsetting (more variable, more biased with 2 reaches)
- No clear relationship between accuracy and $R^2$ or condition number

## Flood wave model

```{r}
uscase <- lis_profiles("lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20)

swot_plot(uscase)
```

## Flood wave model: inference

No bueno. 

```{r, echo = TRUE}
estA0(uscase)

apply(uscase$A, 1, min)
```

## Flood wave model: Diagnostics

Mass-conserved Manning partial residuals in log space

```{r}
uscase %>% 
  swot_sset(keeptimes = -43:-45) %>% 
  val_log_termplot()
```

## Flood wave model: Diagnostics

Linear-A-space mass-conserved plot

```{r}
uscase %>% 
  swot_sset(keeptimes = -43:-45) %>% 
  val_linA_plot() +
  stat_smooth(method = "lm", alpha = 0.2, se = FALSE)
```

## Flood wave model: Diagnostics

Manning n vs depth

```{r}
uscase %>% 
  swot_sset(keeptimes = -43:-45) %>% 
  val_nd_plot(plot = TRUE)
```

## Variability 

Bootstrap resampling of model times (n = 1000)

```{r}
casei <- uscase %>% 
  swot_sset()
  # swot_sset(keeplocs = c(1, 2)) %>% 
  # swot_timelag(c(0, -2))

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE))
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = .))

boot_nunique <- map_int(sampcollist, ~length(unique(.)))
boot_lms <- map(boot_cases, estA0_lm)
boot_A0 <- map(boot_lms, coef)
boot_R2 <- map_dbl(boot_lms, A0_R2)
boot_condno <- map_dbl(boot_lms, A0_condno)

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  map2(boot_R2, ~mutate(.x, R2 = .y)) %>% 
  map2(boot_condno, ~mutate(.x, condno = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = R2)) +
  geom_point(position = position_jitterdodge()) +
  # geom_boxplot(fill = NA) +
  scale_colour_viridis_c() 

```

## Resulting variability of discharge estimates (given known n = 0.04)

```{r}
boot_Ahats <- map(boot_A0, ~swot_A(A0vec = ., dAmat = casei$dA))

boot_Qts <- map(boot_Ahats, ~manningQ(A = ., W = casei$W, S = casei$S, n = 0.04)) %>% 
  map(~apply(., 2, median)) %>% 
  map(~data.frame(Q = ., time = 1:length(.))) %>% 
  bind_rows(.id = "sampno") 

boot_Qts %>% 
  ggplot(aes(x = time, y = Q, group = sampno)) + 
  geom_line(size = 0.3, alpha = 0.04)

```


## Flood wave model with time shift

```{r}
shiftcase <- lis_profiles("lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20) %>% 
  swot_timelag(c(0, -2, -4))

swot_plot(shiftcase)
```

## Flood wave model: inference

Better! 

```{r, echo = TRUE}
estA0(shiftcase)

apply(shiftcase$A, 1, min)
```

## Flood wave model: Diagnostics

Mass-conserved Manning partial residuals in log space

```{r}
shiftcase %>% 
  swot_sset(keeptimes = -39:-41) %>% 
  val_log_termplot()
```

## Flood wave model: Diagnostics

Linear-A-space mass-conserved plot

```{r}
shiftcase %>% 
  swot_sset(keeptimes = -39:-41) %>% 
  val_linA_plot() +
  stat_smooth(method = "lm", alpha = 0.2, se = FALSE)
```

## Flood wave model: Diagnostics

Manning n vs depth

```{r}
shiftcase %>% 
  swot_sset(keeptimes = -39:-41) %>% 
  val_nd_plot(plot = TRUE)
```

## Variability 

Bootstrap resampling of model times (n = 1000)

```{r}
casei <- shiftcase %>% 
  swot_sset()
  # swot_sset(keeplocs = c(1, 2)) %>% 
  # swot_timelag(c(0, -2))

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE))
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = .))

boot_nunique <- map_int(sampcollist, ~length(unique(.)))
boot_lms <- map(boot_cases, estA0_lm)
boot_A0 <- map(boot_lms, coef)
boot_R2 <- map_dbl(boot_lms, A0_R2)
boot_condno <- map_dbl(boot_lms, A0_condno)

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  map2(boot_R2, ~mutate(.x, R2 = .y)) %>% 
  map2(boot_condno, ~mutate(.x, condno = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = R2)) +
  geom_point(position = position_jitterdodge()) +
  # geom_boxplot(fill = NA) +
  scale_colour_viridis_c() 

```

## Resulting variability of discharge estimates (given known n = 0.04)

```{r}
boot_Ahats <- map(boot_A0, ~swot_A(A0vec = ., dAmat = casei$dA))

boot_Qts <- map(boot_Ahats, ~manningQ(A = ., W = casei$W, S = casei$S, n = 0.04)) %>% 
  map(~apply(., 2, median)) %>% 
  map(~data.frame(Q = ., time = 1:length(.))) %>% 
  bind_rows(.id = "sampno") 

boot_Qts %>% 
  ggplot(aes(x = time, y = Q, group = sampno)) + 
  geom_line(size = 0.3, alpha = 0.04)

```

## Other observations

- Highly sensitive to offset!
- cross-covariance function does not always choose optimal offset
    - may be able to restrict using sensible wave velocity and known distances
