---
title: "A' variance"
date: "May 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(dplyr)
library(purrr)
library(ggplot2)
```


The variance of A' takes the following form when slope is known:

$$
Var(A') = 4 \mu_* ^2 \sigma_* ^ 2 + 2 \sigma_*^4
$$

where $\mu_* = \sqrt{\frac{p}{2}} (H - H_1) + \frac{W_1}{\sqrt{2p}}$ and $\sigma_*^2 = \frac{p}{2} \sigma^2_H$.

This assumes normality of H error.


## Unknown slope, $p$

If slope is unknown, then further approximation is required. Depending on the magnitude of uncertainty reltive to actual slope, one of two different approximations are used. 

The general form for variance of A' is:

$$
Var(A') = 4 \mu_* ^2 \sigma_* ^ 2 + 2 \sigma_*^4 + \frac{W_1}{4} \sigma^2_{1/p}
$$

where $\sigma_{1/p}$ is the standard deviation of error in 1/p, when estimated using error-in-variables regression, and:

$$
\mu_* = \frac{H - H_1}{\sqrt{2}}\mu_1 + \frac{W_1}{\sqrt{2}}\mu_2
$$

and 

$$
\sigma_*^2 = \frac{p}{2} \sigma_H^2 + \frac{(H - H_1)^2}{2} \sigma_1^2 + \frac{W_1^2}{2} \sigma^2_2
$$

The formulae for $\mu_1, \mu_2, \sigma_1$, and $\sigma_2$ are approximations, and differ depending on relative magnitudes of $\sigma_p$ and $\sigma_{1/p}$. They are as follows:

$$
\begin{aligned}
\mu_1 &= \sqrt{p} &, \text{ if } \sigma_p < p\\
&= .8222(\sigma_p^2 + p^2)^{1/4} &, \text{ if } \sigma_p > p
\end{aligned}
$$

$$
\begin{aligned}
\mu_2 &= \sqrt{1/p}&, \text{ if } \sigma_{1/p} < \frac{1}{p}\\
&= 0.8222 (\sigma_{1/p}^2 + 1/p^2)^{1/4}&, \text{ if } \sigma_{1/p} > \frac{1}{p}
\end{aligned}
$$

$$
\begin{aligned}
\sigma^2_1 &= p\Big(\exp(\sigma^2_p / (2 p^2)) - \exp(\sigma^2_p / (4 p^2))\Big)&, \text{ if } \sigma_p < p\\
&= 0.1219 \sqrt{\sigma_p^2 + p^2}&, \text{ if } \sigma_p > p
\end{aligned}
$$


$$
\begin{aligned}
\sigma^2_2 &= \frac{1}{p}\Big(\exp(p^2\sigma^2_{1/p} / 2) - \exp(p^2\sigma^2_p / 4)\Big)&, \text{ if } \sigma_{1/p} < \frac{1}{p}\\
&= 0.1219\sqrt{\sigma_{1/p}^2 + 1/p^2}&, \text{ if } \sigma_{1/p} > \frac{1}{p}
\end{aligned}
$$

### Validation of Approximations

Here is a plot illustrating the two different approximations for $\sigma_1^2$:


```{r}
approxvar1 <- function(p, sigma_p) {
  
  expbit <- exp(sigma_p^2 / (4 * p^2))
  out <- (expbit - 1) * expbit * p
  out
}

approxvar2 <- function(p, sigma_p) {
  
  newsigma <- sqrt(sigma_p^2 + p^2)
  newsigma * 0.1219
}


p <- 2
cvs <- 10^seq(-2, 1.5, length.out = 21)
sigmas <- cvs * p
simvars <- map(sigmas, ~abs(rnorm(10000, p, .))) %>% 
  map_dbl(~var(sqrt(.)))

# simvars <- simcvs * p^2

valdf <- data.frame(p = p, sigma_p = sigmas, cv = cvs, 
                    sim_var = simvars) %>% 
  mutate(approx1 = approxvar1(p, sigma_p),
         approx2 = approxvar2(p, sigma_p))

ggplot(valdf, aes(x = cv, y = sim_var)) +
  geom_point(aes(shape = "empirical")) +
  geom_line(aes(y = approx1, color = "approx1")) +
  geom_line(aes(y = approx2, color = "approx2")) +
  # ylim(0, max(valdf$sim_var * 1.5)) +
  scale_x_log10(breaks = 10^(-2:1),
                labels = 10^(-2:1)) +
  scale_y_log10(limits = c(min(valdf$sim_var) / 2,
                           max(valdf$sim_var * 1.5))) +
  annotation_logticks() +
  xlab("Coefficient of Variation") +
  ylab("Variance") +
  theme_bw() +
  theme(legend.title = element_blank())
  
```

Here is a plot illustrating the two different approximations for $\mu_1$:

```{r}
approxmean1 <- function(p, sigma_p) {
  # expbit <- p^(1/2) * exp(sigma_p^2 / (8 * p^2))
  # out <- expbit
  sqrt(p)
  # out
}

approxmean2 <- function(p, sigma_p) {
  newsigsq <- sigma_p^2 + p^2
  # out <- sqrt(2 / pi * newsigsq)
  
  out <- 0.8222 * newsigsq^(1/4)
  out
}


p <- 5
cvs <- 10^seq(-2, 1.5, length.out = 21)
sigmas <- cvs * p
simmeans <- map(sigmas, ~abs(rnorm(10000, p, .))) %>% 
  map_dbl(~mean(sqrt(.)))

# simvars <- simcvs * p^2

valdf <- data.frame(p = p, sigma_p = sigmas, cv = cvs, 
                    sim_mean = simmeans) %>% 
  mutate(approx1 = approxmean1(p, sigma_p),
         approx2 = approxmean2(p, sigma_p))

ggplot(valdf, aes(x = cv, y = sim_mean)) +
  geom_point(aes(shape = "empirical")) +
  geom_line(aes(y = approx1, color = "approx1")) +
  geom_line(aes(y = approx2, color = "approx2")) +
  # ylim(0, max(valdf$sim_mean * 1.5)) +
  scale_x_log10(breaks = 10^(-2:1),
                labels = 10^(-2:1)) +
  scale_y_log10(limits = c(min(valdf$sim_mean) / 2,
                           max(valdf$sim_mean * 1.5))) +
  annotation_logticks() +
  xlab("Coefficient of Variation") +
  ylab("Mean") +
  theme_bw() +
  theme(legend.title = element_blank())
  
```

