---
title: "Bat McMan - Lisflood"
author: "Mark Hagemann"
date: "March 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document summarizes a series of experiments applying the Bat McMan bathymetry estimation method to output from a simple hydrodynamic model (Lisflood). The model consists of three uniform rectangular-channel reaches, each 10000 m in length. Two sets of boundary conditions were run through the model--one simulating steady-state conditions at a series of flow levels ranging from 50 to 300 cms, and the other simulating a storm hydrograph with a peak flow of 300 cms imposed on a baseflow of 50 cms. 

A set of swot-like data at the reach-average scale was composed from the lisflood output. This comprised the SWOT-measurable variables height (H), width (W), slope (S) and partial cross-sectional area ($\delta A$), as well as the unobservable variables flow (Q), and bathymetric cross-sectional area ($A_0$). 

### Method

As described in a previous report, bathymetry was inferred by solving the folloing linear model for $\mathbf{a_0}$ via ordinary least-squares:

$$
\begin{bmatrix}
\mathbf{X} \mathbf{\Omega_1} \\
\mathbf{X} \mathbf{\Omega_2} \\
\vdots \\
\mathbf{X} \mathbf{\Omega_n}
\end{bmatrix}
 \mathbf{a}_0

= 

\begin{bmatrix}
(-\delta \mathbf{A} \circ \mathbf{X}) \omega_1 \\
(-\delta \mathbf{A} \circ \mathbf{X}) \omega_2 \\
\vdots \\
(-\delta \mathbf{A} \circ \mathbf{X}) \omega_n

\end{bmatrix}
$$

where:

- $\mathbf{X}$ is a matrix whose $i,t$th component is $x_{it} = \Big(\frac{1}{n} W_{it}^{-2/3} S_{it}^{1/2} \Big)^{3/5}$
- $\mathbf{a_0}$ is a column vector of $A_0$ values
- $\mathbf{\delta A}$ is a matrix of $\delta A$ values
- $\omega_j, j = 1, \dots, n$ are unique weight vectors whose elements sum to zero (this ensures mass conservation across reaches).  
- $\Omega_j$ is the diagonal matrix composed from the vector $\omega_j$
- $\circ$ denotes the elementwise product. 

### Description of cases

As described above, two different sets of boundary conditions yield steady- and non-steady-state SWOT-like datasets. The steady-state dataset consists of 9 times, corresponding uniquely to the flow conditions 50, 75, 100, 125, 150, 175, 200, 250, and 300 cms (Figure ####). The storm dataset consists of 77 times, representing a high-temporal sampling along the rising and falling limb of a storm hydrograph (Figure ####)

```{r}

is_ss <- function(...) {
  arglist <- list(...)
  ssvars <- map(arglist, function(x) c(1, diff(x))) %>% 
    map(function(x) x == 0) %>% 
    Reduce(`*`, x = .)
  out <- as.logical(ssvars)
  out
}

sscase <- lis_profiles("../lisflood/toy_1/results_const_v2/") %>% 
  lis_reaches(slope_method = "s_median", agfun = median) %>% 
  filter(loc != 4) %>%
  arrange(loc, time) %>% 
  group_by(loc) %>% 
  mutate(steady = is_ss(W, S, Q, H, D)) %>% 
  group_by(time) %>% 
  filter(sum(steady) == 3) %>% 
  group_by(Q) %>% 
  filter(time == time[1]) %>% 
  mutate(A = W * D) %>% 
  select(-steady) %>% 
  swot_untidy()

estA0(sscase)

val_log_termplot(sscase)
# 
# foo <- swot_tidy(sscase)
# glimpse(foo)
```


```{r}
uscase <- lis_profiles("../lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20)

swot_plot(uscase, c("Q", "W", "dA", "S"))
```

Note that in the storm case the flood wave propogation results in considerable spatial variability in flows, such that applying a time shif to reaches 2 and 3 would result in something closer to uniform flow. A separate case was created by applying an optimal shift, as follows:

1. Estimate A0 using the original, unshifted data
2. Compute the cross-correlation function between reach 1 and each of the other reaches, and obtain the best lag as that which results in the largest cross-correlation for each pair of reaches.
3. Shift the reaches by the optimal lags determined in step 2 (reach 1 is not shifted)
4. Re-estimate A0 using the shifted data
5. Repeat steps 2 through 4 until the optimal lags are all zero (indicating convergence) or a maximum number of iterations is reached. 

Applied to the storm case, this resulted in a shifted dataset with lags of -1 and -4 for reaches 2 and 3, respectively. This resulted in a dataset with considerably less spatial heterogeneity in flow (Fig. ####)

```{r}
shiftcase <- uscase %>% 
  # swot_timelag(lags = c(0, -2, -4))
  ccf_lag(verbose = TRUE)

swot_plot(shiftcase, c("Q", "W", "dA", "S"))

manning_qdot(shiftcase, Avar = "A", log = TRUE) %>% 
  plot_DAWG()
```



### Results

```{r}
realA0_ss <- apply(sscase$A, 1, min)
realA0_us <- apply(uscase$A, 1, min)
A0hat_ss <- estA0(sscase)
A0hat_us <- estA0(uscase)
A0hat_us_shift <- estA0(shiftcase)

allA0 <- list(real = realA0_ss, ss = A0hat_ss, us = A0hat_us, us_shift = A0hat_us_shift) %>% 
  as.data.frame() %>% 
  mutate(reach = as.factor(1:3)) %>% 
  gather(key = case, value = A0, -reach)

allA0 %>% 
  # glimpse() %>% 
  ggplot(aes(x = case, y = A0, color = reach)) + 
  geom_point(position = position_dodge(width = 0.2)) +
  theme_bw()
  
```

While the steady-state case yielded near-perfect inference of A0, the two unsteady cases were off by roughly 100%--and in opposite directions! This indicates two things--first, that steady-state conditions are important to accurate A0 inference, and second, that A0 estimates are highly variable in unsteady-state conditions, at least with regard to the amount of time lag applied. 

### Exploration of variability

In order to investigate the degree of variability in A0 estimates, 

```{r}
# Bootstrap A0 by selecting times

casei <- uscase %>% 
  swot_timelag(c(0, -1, -4))

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE))
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = ., keeplocs = c(1, 2, 3)))
boot_A0 <- map(boot_cases, estA0)

boot_nunique <- map_int(sampcollist, ~length(unique(.)))
boot_lms <- map(boot_cases, estA0_lm)
boot_R2 <- map_dbl(boot_lms, A0_R2)
boot_condno <- map_dbl(boot_lms, A0_condno)

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  map2(boot_R2, ~mutate(.x, R2 = .y)) %>% 
  map2(boot_condno, ~mutate(.x, condno = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = R2)) +
  geom_point(position = position_jitterdodge()) +
  scale_colour_viridis_c()

boot_df %>% 
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>% 
  glimpse() %>% 
  ggplot(aes(x = R2, y = A0_estimate, color = reach)) +
  geom_point()

boot_df %>% 
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>% 
  glimpse() %>% 
  ggplot(aes(x = condno, y = A0_estimate, color = reach)) +
  geom_point()

boot_df %>% 
  gather(key = reach, value = A0_estimate, -sampno, -nunique, -R2, -condno) %>% 
  glimpse() %>% 
  ggplot(aes(x = condno, y = R2, color = reach)) +
  geom_point()
```

That is not as I expected for sscase. Is that because it's repeating hydrologic conditions?

```{r}

casei <- sscase

cols <- 1:ncol(casei$W)
sampcollist <- lapply(1:1000, function(x) sample(cols, replace = TRUE)) %>% 
  map(unique)
boot_cases <- map(sampcollist, ~swot_sset(casei, keeptimes = .))
boot_A0 <- map(boot_cases, estA0)

boot_nunique <- map_int(sampcollist, ~length(unique(.)))

boot_df <- boot_A0 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(boot_nunique, ~mutate(.x, nunique = .y)) %>% 
  bind_rows(.id = "sampno")

boot_df %>%
  gather(key = reach, value = A0_estimate, -sampno, -nunique) %>%
  # filter(nunique > 7) %>%
  ggplot(aes(x = reach, y = A0_estimate, color = as.factor(nunique))) +
  geom_point(position = position_jitterdodge()) +
  scale_colour_viridis_d()
  # geom_violin()
```

This is crazy! Just to verify, prune one flow point manually and rerun.

```{r}
sscase2 <- sscase

A0s2 <- list()

for (j in 1:9) {
  remtime <- j
  for (i in 1:length(sscase)) {
    sscase2[[i]] <- sscase[[i]][, -remtime]
  }
  A0s2[[j]] <- estA0(sscase2)
}

foogg <- A0s2 %>% 
  map(~as.data.frame(as.list(.))) %>% 
  map2(1:9, ~mutate(.x, j = .y)) %>% 
  bind_rows() %>% 
  gather(key = reach, value = A0, -j) %>% 
  ggplot(aes(x = reach, y = A0, color = j)) +
  geom_point()

plotly::ggplotly(foogg)
```

It seems like the exactitude of the steady-state A0 estimate was a happy accident afforded by the specific flows in the dataset. 

```{r}
swot_sset(sscase, keeptimes = -9) %>%
  # swot_plot()
  # A0_condno()
  # val_nd_plot() + geom_line() + xlim(0, 1.4)
  # val_linA_plot()
  manning_ws35() %>% t() %>% pairs()
  # val_log_termplot(scales = "free_x")
  # estA0_lm() %>% 
  # markstats::ggTermPlot(xvar = "x2") +
  # ylim(-2, 2)

sscase %>% 
  # swot_plot()
  # A0_condno()
  # val_nd_plot() + geom_line() + xlim(0, 1.4)
  # val_linA_plot()
  manning_ws35() %>% t() %>% pairs()
  # val_log_termplot(scales = "free_x")
  # estA0_lm() %>% 
  # markstats::ggTermPlot(xvar = "x2") +
  # ylim(-2, 2)

```

While the method is simple (linear regression), the quantities involved are abstract, and the results can be perplexing. This document is an attempt to demonstrate the method and elucidate the 

