---
title: "Untitled"
author: "Mark Hagemann"
date: "February 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Make some synthetic data to try and understand A0 calculation. 

If I have a system that varies in time and space governed by Manning's equation. Distributions shouldn't matter. 

Simulate

- Q_t
- A_0i
- delta_it
- x_it


```{r}
nt <- 3
nx <- 2
qt <- rnorm(nt, 5, 0.5)
qit <- matrix(rep(qt, nx), nrow = nt)
xit <- matrix(rnorm(nt * nx, 3, 0.3), nrow = nt)
ait <- qit / xit
# a0 <- rnorm(nx, 2, 0.2)
a0 <- c(1, 2)
a0it <- matrix(rep(a0, nt), nrow = nt, byrow = TRUE)
dait <- ait - a0it

yvec <- (xit * -dait) %*% c(1, -1)

Omega <- matrix(c(1, 0, 0, -1), nr = 2)

desgn <- xit %*% Omega


moddf <- data.frame(x1 = desgn[, 1], x2 = desgn[, 2], y = yvec)

modlm <- lm(y ~ 0 + x1 + x2, moddf)

summary(modlm)

pairs(moddf)
```

OK, that proves it! The method is sound. Now I need to write it up. 


Let's write a function to estimate A0 from data!


```{r}

estA0 <- function(wmat, smat, damat) {
  
  stopifnot(dim(wmat) == dim(smat) && dim(smat) == dim(damat))
  
  xmat <- t((wmat ^ (2/3) * smat ^ (1/2)) ^ (3/5)) # this is time-down, space-across.
  
  parinds <- t(combn(1:nrow(wmat), 2))
  
  
  nx <- ncol(xmat)
  nt <- nrow(xmat)
  nxpos <- ceiling(nx / 2)
  nxneg <- floor(nx / 2)
  
  wts <- rep(c(1/nxpos, -1 / nxneg), length.out = nx)
  
  dsgn <- xmat %*% diag(wts)
  lmout <- 
  fmla
}


estA0_2loc <- function(wmat, smat, damat, locinds) {
  
  stopifnot(dim(wmat) == dim(smat) && dim(smat) == dim(damat))
  xmat <- t((wmat ^ (2/3) * smat ^ (1/2)) ^ (3/5))[, locinds] # this is time-down, space-across.
  
  wts <- c(1, -1)
  
  dsgn <- xmat %*% diag(wts)
  rsps <- (xmat * t(-damat)[, locinds]) %*% wts 
  
  lmdf <- na.omit(data.frame(x1 = dsgn[, 1], x2 = dsgn[, 2], y = rsps))
  
  lmout <- lm(y ~ 0 + x1 + x2, data = lmdf)
  # out <- coef(lmout)
  # out
}
  

```

Make some fake data to try it on!

```{r}
nt <- 40
nx <- 2
qt <- rnorm(nt, 5, 0.5)
qit <- matrix(rep(qt, nx), nrow = nt)
xit <- matrix(rnorm(nt * nx, 3, 0.3), nrow = nt)
ait <- (qit / xit)^(3/5)
wmat <- t(matrix(rnorm(nt * nx, 1.5, 0.15), nrow = nt)) ^ (-3/2)
smat <- (t(xit) * wmat^(2/3)) ^ (2)
# a0 <- rnorm(nx, 2, 0.2)
a0 <- c(1, 2)
a0it <- matrix(rep(a0, nt), nrow = nt, byrow = TRUE)
damat <- t(ait - a0it)
locinds <- c(1, 2)

foo <- estA0_2loc(wmat, smat, damat, c(1, 2))
summary(foo)
```


Try it on some real data!

```{r}

library(swotData)

testdat <- swotData::Pepsi_v2$Ganges

wmat = testdat$w
smat = testdat$s 
damat = testdat$dA

testa0 <- estA0_2loc(wmat = testdat$w, smat = testdat$s, damat = testdat$dA, 
                     locinds = c(1, 2))
summary(testa0)
```



Here's an idea I had to add an arbitrary number of rows to the model matrix

```{r}
estA0 <- function(wmat, smat, damat) {
  
  stopifnot(dim(wmat) == dim(smat) && dim(smat) == dim(damat))
  
  # Assume these inputs come as DAWG-style time-across, space-down matrices. 
  # Transpose.
  wmat <- t(wmat)
  smat <- t(smat)
  damat <- t(damat)
  
  
  xmat <- (wmat ^ (-2/3) * smat ^ (1/2)) ^ (3/5)
  
  nx <- ncol(xmat)
  
  omegas <- lapply(1:100, function(x) rnorm(nx))
  omegas <- lapply(omegas, function(x) x - mean(x))
  
  modmats <- lapply(omegas, function(x) xmat %*% diag(x))
  
  modmat <- Reduce(rbind, modmats)
  
  rsps <- lapply(omegas, function(x) (xmat * -damat) %*% x) 
  rsp <- unlist(rsps)
  
  lmdf <- setNames(as.data.frame(cbind(rsp, modmat)), c("y", paste0("x", 1:nx)))
  
  lmout <- lm(y ~ 0 + ., data = lmdf)
  # out <- coef(lmout)
  # out
}
```

Now test *that* out!

```{r}

goodr <- foonc$River_Info.gdrch
ashw <- foonc$Reach_Timeseries.W[goodr, ]
ashs <- foonc$Reach_Timeseries.S[goodr, ]
asha <- foonc$Reach_Timeseries.A[goodr, ]
asha0 <- apply(asha, 1, median)
ashda <- asha - matrix(rep(asha0, ncol(asha)), ncol = ncol(asha))

asha0est <- estA0(wmat = ashw, smat = ashs, damat = ashda)

summary(asha0est)

asha0


plot(asha0, asha0est$coefficients)
abline(0, 1)
```


Real quick, do this for all Pepsi 2.0 cases. 

```{r}
ncfiles <- list.files("../../swotData/data/NC_files/", full.names = TRUE, pattern = "\\.nc$")

cases <- lapply(ncfiles, nc_list)
casenames <- list.files("../../swotData/data/NC_files/", pattern = "\\.nc$") %>% 
  gsub("\\.nc$", "", .)

a0list <- list()
estlist <- list()
complist <- list()

for (i in 1:length(cases)) {
  
  case <- cases[[i]]
  
  print(ncfiles[i])
  
  goodr <- case[["River_Info.gdrch"]]
  amat <- case[["Reach_Timeseries.A"]][goodr, ]
  a0vec <- apply(amat, 1, median) # median version
  # a0vec <- apply(amat, 1, min) # minimum version
  
  a0mat <- matrix(rep(a0vec, ncol(amat)), ncol = ncol(amat))
  damat <- amat - a0mat
  
  lmi <- try(estA0(wmat = case[["Reach_Timeseries.W"]][goodr, ],
               smat = case[["Reach_Timeseries.S"]][goodr, ],
               damat = damat))
  if (inherits(lmi, "try-error")) next
  
  a0list[[i]] <- a0vec
  estlist[[i]] <- coef(lmi)
  
  complist[[i]] <- data.frame(real = a0vec, est = coef(lmi))
  
}

names(a0list) <- casenames
names(estlist) <- casenames
names(complist) <- casenames

nulls <- sapply(estlist, is.null)

a0compare <- dplyr::bind_rows(complist[!nulls], .id = "case")

glimpse(a0compare)
```


See how this looks!

```{r}
library(ggplot2)

ggplot(a0compare, aes(x = real, y = est)) + 
  geom_point(aes(color = as.factor(case))) + 
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(aes(slope = 1, intercept = 0))
  
ggsave("../graphs/A0_estimate_log.png")
```



