---
title: "Lisflood cases for A0 inversion"
author: "Mark Hagemann"
date: "April 5, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
opts_knit$set(root.dir = "~/Documents/mcfli-swotr")
opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
library(ProjectTemplate)
load.project()
ggplot2::theme_set(theme_bw())
```

    
## Lisflood model

- 3 10-km reaches 
- Uniform 100-m wide rectangular channel (all reaches)
- Uniform bed slope within each reach
- Uniform Manning's n = 0.04

## Model runs

- Steady state
    - 9 Flow conditions between 50 and 300 cms
    - Model run until steady state at all reaches
    - Dataset compiled from steady-state conditions
- Nonsteady flood wave
    - NRCS unit hydrograph
    - Flow between 50 and 300 cms

## Steady-state model

```{r}
is_ss <- function(...) {
  arglist <- list(...)
  ssvars <- map(arglist, function(x) c(1, diff(x))) %>% 
    map(function(x) x == 0) %>% 
    Reduce(`*`, x = .)
  out <- as.logical(ssvars)
  out
}

sscase <- lis_profiles("lisflood/toy_1/results_const_v2/") %>% 
  lis_reaches(slope_method = "s_median", agfun = median) %>% 
  filter(loc != 4) %>%
  arrange(loc, time) %>% 
  group_by(loc) %>% 
  mutate(steady = is_ss(W, S, Q, H, D)) %>% 
  group_by(time) %>% 
  filter(sum(steady) == 3) %>% 
  group_by(Q) %>% 
  filter(time == time[1]) %>% 
  mutate(A = W * D) %>% 
  select(-steady) %>% 
  swot_untidy()

gg1 <- swot_plot(sscase[c("A", "dA", "H", "Q", "S", "W")]) + 
  theme_bw()
ggsave("../graphs/poster/sscase.png", gg1, width = 4, height = 3, dpi = 500)
```

## Flood wave model

```{r}
uscase <- lis_profiles("lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20)

gg2 <- swot_plot(uscase[c("A", "dA", "H", "Q", "S", "W")]) +
  theme_bw()
ggsave("../graphs/poster/uscase.png", gg2, width = 4, height = 3, dpi = 500)
```

## Flood wave model with time shift

```{r}
shiftcase <- lis_profiles("lisflood/toy_1/results_simple_v2/") %>% 
  lis_reaches(slope_method = "s_mean", agfun = mean) %>% 
  filter(loc != 4) %>%
  mutate(A = W * D) %>% 
  swot_untidy() %>% 
  swot_sset(keeptimes = -1:-20) %>% 
  swot_timelag(c(0, -2, -4))

swot_plot(shiftcase)
```
