---
title: "McFLI-SWOT-R ecosystem"
author: "Mark Hagemann"
date: "March 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The packages described in this document are intended to assist the inference of hydrologic quantities from SWOT-like observations. These packages include:

- swotData: read/write SWOT-like data including Pepsi test cases
- McMan
    - Mass-conserved Manning equation fucntions
    - A0 inference using Mass-conserved Manning
- Lisfloodr
    - i/o, possibly running, plotting, conversion to swotData, etc. 


### SWOT-like data

`swotdata` contain measurements of stream height, slope, width, and partial cross-sectional area. The may further contian discharge, (full) cross-sectional area, and other data or metadata. 

### Data import functions

- nc_list (swotData)
- nc_reach (swotData)
- nc_xs (swotData)


### Data prep functions

- calcdA_vec (swotData, improved in lib/estA0.R)
- calcdA_mat (swotData, improved in lib/estA0.R)
- calcslope (swotData)

- swot_tidy (manningEval/lib/utils.R)
- swot_untidy (manningEval/lib/utils.R)
- swot_timelag (manningEval/lib/utils.R)
- smart_timelag (TODO)
- rezero_dA (manningEval/lib/utils.R)
- swot_A (manningEval/lib/utils.R)


### Inference functions

- estA0, estA0_list (lib/estA0.R)
- manning_qdot (manningEval/lib/utils.R)

### Diagnostics functions

- swot_n_anova_lm (manningEval/lib/anova.R)
- swot_n_anova (manningEval/lib/anova.R)

### Data plotting functions

- plot_DAWG (swotData)
- plot_swot (swotData)


### Lisflood functions

- read_lisProfile (lib/lisflood.R)


## Example Analysis 1

### Read in data 

### Calculate SWOT variables

dA

### Plot, EDA

### Deal with non-positive and/or missing values

### Tidy and untidy

### Add a time lag

```{r}
foo <- reachdata$Platte
foolag <- ccf_lag(foo)
```


## Example analysis 2: doing some inference

```{r}
foolag$dA <- rezero_dA(foolag$dA, zero = "median")
fooA0 <- estA0_list(foolag)
summary(fooA0)
```


## Example analysis 3: validation using known discharge and area

### Diagnose Manning equation performance

SWOT-like data

- Flow imbalance estimate

```{r}
# swot_dqdx(foolag)
# swot_dqdx(foo)
dqs_lagged <- reachdata %>% 
  map(possibly(ccf_lag, otherwise = NA)) %>% 
  map_dbl(possibly(swot_dqdx, otherwise = NA_real_))
dqs_nolag <- reachdata %>% 
  map_dbl(possibly(swot_dqdx, otherwise = NA_real_))

plot(dqs_nolag - dqs_lagged)
abline(h = 0)
```



- 3/5-space condition number
    - stepwise regression
- 3/5-space R^2
    - stepwise regression

SWOT+ data

- Exact flow imbalance
- Reach-scale R^2 in 3/5 space
    - steady-state mass conserved and not
- Manning partial residual plots
- Manning equation R2 in log space
- ANOVA on log n


### Validate A0 inference

- Straight-up A0 comparison

```{r}

```

- Marginal influence on Q estimation
