---
title: "Closure decomposition"
author: "Mark Hagemann"
date: "May 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Not including measurement error, the model for Mass-conserved Manning uncertainty is as follows:

$$
\begin{aligned}
-\frac{2}{3}\log W_{it} + \frac{5}{3} \log A_{it} + \frac{1}{2} \log S_{it} - \log ( \bar{Q_t}\bar{n}) &= \gamma_{it} + \nu_{it} \\
&=  \gamma'_t(x_i - \bar{x})) + \nu_{i \cdot} + \epsilon_{it} \\
\end{aligned} \\
\gamma'_t \sim N(0, \sigma_{\gamma'}) \\
\sum_{i = 1}^{N_r}\nu_{i\cdot} = 0 \\
\epsilon_{it} ~ N(0, \sigma_\epsilon)
$$

Given a swot-like dataset, there are several functions for analyzing the closure term on the right-hand side of equation ####.

First, load a dataset. We'll work with GaronneUpstream from the Pepsi 1 cases. 



```{r}
datai <- reachdata$GaronneUpstream

swot_plot(datai)
```

The full closure term, $n_it$ is given by $n_{it} = \frac{1}{\bar{Q}_t} W_{it}^{-2/3} A_{it}^{5/3} S_{it}^{1/2}$. Toggling off the the `mc` argument gets rid of steady-state mass conservation, replacing $\bar{Q}_t$ with $Q_{it}$ in the previous equation. 

```{r}
plot_DAWG(manning_closure(datai))
plot_DAWG(manning_closure(datai, mc = FALSE))
```

Various functions show the decomposition of the closure term. The first-level decomposition is into variability from flow-imbalance ($\gamma_{it}$) and variability from non-constant flow resistance ($\nu_{it}$).

```{r}
plot_DAWG(manning_gamma(datai))
plot_DAWG(manning_nu(datai))

```

We can further decompose these. First $\gamma_{it} = \hat{\gamma}_{it} + \epsilon_{\gamma} = \gamma'_{t}(x_{i} - \bar{x}) + \epsilon_\gamma$

```{r}
gamma_decomp <- datai %>% 
  manning_gamma() %>% 
  decomp_gamma(xmat = datai$x)

plot_DAWG(gamma_decomp$gammahat)
plot_DAWG(gamma_decomp$gammaerr)
plot(gamma_decomp$dgdx, type = "l")
sd(gamma_decomp$dgdx) * sd(datai$x[, 1])
```

And similarly, but more simply, $\nu = \hat{\nu}_{it} + \epsilon_\nu = \bar{\nu}_{i\cdot} + \epsilon_{\nu}$

```{r}
nu_decomp <- datai %>% 
  manning_nu() %>% 
  decomp_nu()

plot_DAWG(nu_decomp$nuhat)
plot_DAWG(nu_decomp$nuerr)
plot(nu_decomp$nubar)

```

Finally we can do the full closure decomposition using `characterize_closure`. 

```{r}
clos_decomp <- characterize_closure(datai)
str(clos_decomp)

clos_decomp$dgdx * clos_decomp$dx
clos_decomp$alpha
clos_decomp$err
```


This leads to a universally applicable metric for flow imbalance. A rough appoximation of the amount of flow variability between two reaches of a given river segment is given by the product of the standard deviation of `dgdx` and that of `x`. 


### ANOVA on closure. 

We can inspect how much of the total closure error comes from the various terms in equation #### in a couple of ways.

```{r}

```


```{r}
dQ_df <- c(reachdata, list(sscase = sscase, uscase = uscase)) %>% 
  map(swot_purge_nas) %>% 
  map(possibly(~characterize_closure(., method = "anova"), 
               otherwise = list(dQ_pct = NA_real_))) %>% 
  # `[`(!is.na(.)) %>%
  map_dbl(~.[["dQ_pct"]]) %>% 
  data.frame(case = names(.), dQ_pct = ., row.names = NULL)

dQ_df %>% 
  glimpse() %>% 
  ggplot(aes(x = case, y = dQ_pct)) + 
  geom_point() +
  theme(axis.text = element_text(angle = 90))

plot_DAWG(manning_closure(reachdata$Cumberland, log = TRUE))

manning_closure(reachdata$Cumberland, log = TRUE) %>% 
  list(clos = .) %>% 
  swot_tidy() %>% 
  filter(time > 125, time < 145) %>% 
  glimpse()
```

Let's look at a ribbon for anovas.

```{r}
anova1df <- c(reachdata, list(sscase = sscase, uscase = uscase)) %>% 
  map(swot_purge_nas) %>% 
  map(possibly(closure_anova, otherwise = data.frame())) %>% 
  bind_rows(.id = "case")

anova1df %>% 
  ggplot(aes(x = case, y = pctTotVar, fill = piece)) +
  geom_bar(stat = "identity", position = "stack") +
  theme(axis.text = element_text(angle = 90)) +
  xlab("")

```

